### 2009-02-26 honors_thesis/paper.tex
Started on honors thesis, added some other things.

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{Honors Thesis}
\author{Matthew Gardner}
\date{\today}

\begin{document}
\maketitle

\section{Statement of Intent}

\section{Background and Significance}

\section{Methodology/Procedures}

\section{Preliminary Prospectus of Finished Thesis}

\section{Preliminary Research}

\section{Qualifications of Investigator}

\section{Qualifications of Faculty Advisor}

\section{Schedule}

\section{Expenses/Budget}

\section{Closure}

\section{References}

\section{Appendices}

\end{document}