### 2009-03-09 honors_thesis/paper.tex
Some revisions.

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{amsmath}

\DeclareMathOperator{\URand}{U}
\providecommand{\ppos}{\ensuremath{\Vec{x}}}
\providecommand{\pvel}{\ensuremath{\Vec{v}}}
\providecommand{\gbest}{\ensuremath{\Vec{g}}}
\providecommand{\pbest}{\ensuremath{\Vec{p}}}
\providecommand{\constriction}{\ensuremath{\chi}}
\providecommand{\coeff}{\ensuremath{\phi}}
\newcommand{\bib}[1]{../../bib/#1}

\topmargin .5in

\title{Honors Thesis Proposal}
\author{Matthew Gardner}
\date{\today}

\begin{document}
\maketitle

\section{Background and Significance}

Optimization problems are ubiquitous in our modern world.  Businesses such as
Google need to decide where to place ads, scientists need to fit models to
data, and airlines need to schedule their flights.  All of those problems are
optimization problems, and optimization techniques can be used to find
solutions.  Optimization algorithms are derived from all sorts of natural
phenomena.  Some methods look at the landscape of the function as they are
searching and follow hills to the best values, as in the gradient ascent
method.  Other methods take ideas from biology or metallurgy, like genetic
algorithms and simulated annealing.  Particle swarm optimization is a recently
developed optimization technique that draws on ideas from the sociology of
flocking birds.  All of these methods, however, are fundamentally sequential in
nature---they need to be run in a specific order, and because of that, they are
typically only run on one machine.  The problems people want to solve are
getting bigger, and larger problems need to run on multiple machines in
parallel.  Parallel implementations of these algorithms have been attempted,
but they do not often perform as well as sequential versions for the amount of
computation performed.  The work I propose focuses on one of those algorithms,
particle swarm optimization, and improving its performance in a large-scale,
parallel world.

\section{Statement of Intent}

I will study the feasibility and potential performance gain of speculative
execution in particle swarm optimization.  By speculative execution I mean the
use of additional computational resources, presumably extra processors running
in parallel, to perform two or more iterations of particle swarm optimization
at a time, speeding up the algorithm significantly.

\section{Procedures}

The bulk of the work for this project will be writing code and running
experiments.  I have already read many conference papers on particle swarm
optimization to know the current state of the field, so I will know what the
results need to be compared to.  In implementing speculative execution, I will
take care that the algorithm behaves exactly like the original, standard
particle swarm optimization (PSO) algorithm in every aspect except the
speed-up.  The same constants will be used as are commonly published in the
literature.

In running experiments, many trials will be run and averaged, as there is a
random component in this algorithm.  I will use statistical tests for all
comparisons to the original PSO algorithm.  The success criterion will be a
better performance of the algorithm as measured by the lowest function
evaluation seen over a number of iterations for several benchmark and
real-world problems.  Speculative execution takes more computation than regular
PSO with the same number of particles, so the two algorithms will be compared
at points where they use equal amounts of processing power.

\section{Preliminary Prospectus of Finished Thesis}

The finished thesis will be presented as a conference paper, including the
following sections:

\begin{enumerate}
\item Introduction
\item Related Work
\item Particle Swarm Optimization
\item Speculative Execution
\item Speculative Execution in PSO
\item Experiments
\item Results and Conclusions
\item Future Work
\item References
\end{enumerate}

\section{Preliminary Research}

\subsection{Particle Swarm Optimization}

Particle swarm optimization was proposed in 1995 by James Kennedy and Russell
Eberhart.  It tries to intelligently search a multi-dimensional space by
mimicking the swarming and flocking behavior of birds and other
animals.~\cite{kennedy-icnn95} It is a sociological algorithm that depends on
interaction between particles to quickly and consistently find the optimal
solution to a problem.  The algorithm keeps track of a number of potential
solutions, called particles, which move somewhat randomly through the search
space.  Particles remember the best place they have been, or solution they have
evaluated, and are attracted back to that place, as well as to the best
solution other particles have seen.  Specifically, the formulas for updating
the position $\ppos_t$ and velocity $\pvel_t$ of a particle at iteration $t$ are
as follows:
\begin{align}
\label{eq:velupdate}
	\pvel_{t+1} &=
		\constriction \left[ \pvel_t +
			\coeff_1\URand()\otimes(\pbest - \ppos_t) +
			\coeff_2\URand()\otimes(\gbest - \ppos_t)
		\right] \\
\label{eq:posupdate}
	\ppos_{t+1} &= \ppos_t + \pvel_{t+1}
\end{align}
where \( \URand() \) is a vector of random numbers drawn from a uniform
distribution, the \( \otimes \) operator is an element-wise vector
multiplication, $\pbest$ is the best position the current particle has seen,
and $\gbest$ is the best position any of the other particles have seen.  \(
\coeff_1 \), \( \coeff_2 \), and \( \constriction \) are parameters with
prescribed values required to ensure convergence (2.05, 2.05, and .73,
respectively).~\cite{clerc-tec02}~\cite{poli-aea08}

The sociology in this algorithm defines which other particles form the
``neighborhood'' of a given particle---the other particles whose best solution
it sees.  There are many ways to define a particle's neighborhood, varying from
the entire rest of the swarm to just one other particle.

The way a neighborhood is defined can have drastic effects on the performance
of the algorithm; the more neighbors each particle has, the faster information
spreads throughout the particles, and the quicker the algorithm converges.  For
some problems it is possible that the algorithm converges too quickly and gets
caught in a local optimum; it stops on a hill when there is a mountain next to
it.  For those kinds of problems, less communication, or smaller neighborhoods,
is often preferable.

\subsection{Speculative Execution}

Speculative execution in a program is the execution of code that may or may not
end up actually being needed.  Modern processors routinely do this when a
conditional branch is encountered; they try to predict which branch will
actually be needed, and continue their execution.  If it turns out that the
prediction was incorrect, the work is discarded.  But if the prediction was
right, the program can execute much faster than if it had waited on the branch.

In particle swarm optimization, speculative execution can be performed because
the next position a particle moves to doesn't depend on the value of the
function at its current position.  That means you can just compute all possible
next positions (depending on whether or not the gbest or pbest were updated---\gbest 
and \pbest in equation \ref{eq:velupdate} got updated), and evaluate each position 
in parallel.  That computes two iterations at the same time.

\section{Qualifications of Investigator}

I am a Computer Science major.  My main focuses in my undergraduate career have
been optimization, natural language processing, and machine learning.  The
optimization emphasis has come mainly in my work with Dr. Kevin Seppi, for whom
I am a research assistant.  I have spent the last year working with him,
working mainly on particle swarm optimization topics.  We worked together with
Dr. Branton Campbell, a physics professor, to solve a crystallography problem
for him using a parallel implementation of PSO.  Those results will be
submitted to Nature Materials.  Recently we, along with a graduate student who
works in the lab, submitted a paper to the Congress on Evolutionary
Computation, a major venue for optimization techniques.  The paper was on
topologies and communication in large particle swarms, and formed the
beginnings of the work I propose.

\section{Qualifications of Faculty Advisor}

Dr. Kevin Seppi is the head of the Applied Machine Learning Laboratory at BYU.
Three of his last five graduate students have done their graduate work in
particle swarm optimization, and he has authored or co-authored over 10 papers
on the topic.  I took CS 236 from him in the fall of 2007 and decided I wanted
to do research in the area he was researching, so I joined his lab.  I too
became interested in PSO and have worked in that area with him for the past
year.

\section{Schedule}

\begin{tabular}{|l|l|}
\hline
Run preliminary tests to ensure feasibility & February 28, 2009 \\
\hline
Finish writing and debugging code & March 30, 2009 \\
\hline
Finish gathering data and statistics & June 30, 2009 \\
\hline
Complete analysis of data and begin writing thesis & September 1, 2009 \\
\hline
Submit completed copy of thesis to advisor & November 15, 2009 \\
\hline
Submit paper detailing results to CEC 2010 & December 20, 2009 \\
\hline
Final thesis and portfolio submitted & February 1, 2010 \\
\hline
Thesis defense completed & March 1, 2010 \\
\hline
Final four thesis copies submitted & March 8, 2010 \\
\hline
\end{tabular}

\section{Expenses/Budget}

This project will not require any outside funding.  The only expenditure
will be my own time researching, conducting experiments, and writing.

\section{Closure}

Thank you for reviewing my proposal.  I look forward to exploring speculative
execution in particle swarm optimization.

\bibliographystyle{plain}
\bibliography{%
\bib{pso/kennedy-icnn95},%
\bib{pso/clerc-tec02},%
\bib{pso/poli-aea08}}

\end{document}