### 2009-10-29 orca/finalreport2008.tex
Revisions to my honors thesis, a bunch of other random stuff

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{cite}

\newcommand{\bib}[1]{/aml/home/mjg82/bib/#1}

\title{Speculative Evaluation in Particle Swarm Optimization}
\author{Matthew Gardner\\Mentor: Kevin Seppi, Computer Science}
\date{}

\begin{document}
\maketitle

Optimization problems are ubiquitous in our modern world.  Businesses such as
Google need to decide where to place ads, scientists need to fit models to
data, and airlines need to schedule their flights.  All of these problems are
optimization problems, and optimization techniques can be used to find
solutions.  Particle swarm optimization (PSO) is a relatively new optimization
technique that draws on ideas from the sociology of flocking birds.  It was
originally developed for use on a single machine, and attempts at modifying the
algorithm to run in parallel are still quite preliminary.  However, the
problems people want to solve are getting bigger, and larger problems require
more computation power than is available on any single computer.  To
effectively solve those problems, optimization algorithms need to be improved
to efficiently use parallel clusters of machines.  This project addressed that
issue.

Particle swarm optimization was proposed in 1995 by James Kennedy and Russell
Eberhart~\cite{kennedy-icnn95}.  The algorithm is used to intelligently search
a multi-dimensional space by mimicking the swarming and flocking behavior of
birds and other animals. It is a social algorithm that depends on interaction
between particles to quickly and consistently find the optimal solution to some
objective function.

The motion of particles through the search space has three components: an
inertial component that gives particles momentum as they move, a cognitive
component where particles remember the best solution they have found and are
attracted back to that place, and a social component by which particles are
attracted to the best solution that any of their neighbors have found.

Within the last few years, researchers have begun to recognize the need to
develop parallel implementations of PSO.  The methods they have used include
various synchronous algorithms, asynchronous algorithms, and formulating PSO in
Google's MapReduce framework.  Parallelizing the evaluation of the objective
function can also be done in some cases, though that is not an adaption of the
PSO algorithm and thus is not the focus of this paper.

All of these parallel techniques use additional processors to increase the
number of particles in the swarm, either by adding individual particles or by
adding entire new subswarms.  The number of iterations of PSO that the
algorithms can perform is thus inherently limited by the time it takes to
evaluate the objective function--additional processors add more particles, not
more iterations.  However, there are many functions for which performing more
iterations with fewer particles is a better use of resources than performing
fewer iterations with more particles.

To solve this problem, we take inspiration from speculative execution
techniques commonly used in processors.  Modern processors, when faced with a
branch on which they must wait, guess which way the branch will go and start
executing.  If the processor guessed right, execution is much farther ahead
than if it had just idly waited on the branch.  If it guessed wrong, execution
restarts right where it would have been anyway.  This is called speculative
execution or branch prediction.

In PSO, the motion of particles depends only on the \emph{positions} of the
particle and its neighbors and which neighbor's position is best, not the
\emph{value} of the objective function at those positions.   We do not have to
wait for a particle's position to be evaluated to know where it might move to
next, so we can compute the set of all possible next positions and send them to
be evaluated along with the particle's current position.  Once the current
position of each particle has been evaluated, we can determine which of the
speculatively evaluated positions matches the branch actually taken by the
algorithm.  Because we already evaluated that position, we have performed two
iterations at once.

We implemented this parallel adaptation of the PSO algorithm and found that in
many cases it performed better than a traditional parallelization of PSO.  This
work was my honors thesis and was submitted to BYU for that purpose.  We also
submitted our work to \emph{The Genetic and Evolutionary Computation
Conference} (GECCO), one of the most prestigious conferences for PSO-related
research, and we will find out if it was accepted by March 10, 2010.

This work left us with a lot of open questions.  The most burning question is
how to make the speculative algorithm more usable.  The algorithm we developed
for speculative evaluation was somewhat limited in that it required a large
number of extra processors to perform speculative evaluations.  For example,
for a swarm of size 50 it required 400 processors.  If you want a swarm of 50
particles, but only have 200 processors, with the algorithm we developed you
can not use the extra 150 processors to perform some speculative evaluation;
you have to have 400, or decrease your swarm size.  A much more desirable trait
of the algorithm would be to specify a swarm size, then use any additional
processors, however many there are, to perform what speculative evaluation you
can.  This would require some adaptation to the algorithm, not requiring that
particles are at the same iteration at the same time, as there are only enough
processors for some of the particles to speculate about their future positions.

We plan to pursue this area of research, and are submitting another ORCA grant
proposal to answer that question.  We would then take the conference paper we
submitted to GECCO and expand it with the aim of submitting it to
\emph{Transactions on Evolutionary Computation}, a journal that deals with
PSO-related research, along with other evolutionary computation topics.

\bibliographystyle{plain}
\bibliography{%
\bib{wolpert-tec97},%
\bib{merris-01},%
\bib{pso/clerc-tec02},%
\bib{pso/kennedy-icnn95},%
\bib{pso/bratton-sis07},%
\bib{pso/largeswarms/perez-emc05},%
\bib{pso/largeswarms/montes-gecco08.bib},%
\bib{pso/topology/liang-sis05.bib},%
\bib{pso/parallel/belal-ijicis04},%
\bib{pso/parallel/chu-sci06},%
\bib{pso/parallel/jin-aps05},%
\bib{pso/parallel/koh-ijnme06},%
\bib{pso/parallel/mostaghim-report06},%
\bib{pso/parallel/parsopoulos-aia04},%
\bib{pso/parallel/schutte-ijnme04},%
\bib{pso/parallel/venter-wcsmo05},%
\bib{ga/poli-ai06},%
\bib{sim-anneal/parallel/witte-tpds91.bib},%
\bib{pso/topology/kennedy-cec02.bib},%
\bib{pso/topology/mohais-cec04.bib},%
\bib{pso/topology/jordan-gecco08.bib},%
\bib{pso/topology/clerc-oep03.bib},%
\bib{pso/topology/miranda-ijcir08.bib},%
\bib{pso/topology/mendes-phd04.bib},%
\bib{aml/mcnabb-cec07},%
\bib{aml/mcnabb-cec09}}

\end{document}