### 2009-11-19 great_works/music_brahms.tex
Some initial work on writing more responses

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{Mendelssohn's Italian Symphony: A Music Response}
\author{Matthew Gardner}
\date{}

\begin{document}

\textbf{Music Response}

\begin{tabular}{ll}
  Name:&Matthew Gardner \\
  Work:&Mendelssohn's Symphony No. 4 (The Italian Symphony) \\
  Artist:&Felix Mendelssohn \\
  Date:&1833 \\
\end{tabular}

\section*{Historical Context}

\section*{Critical Analysis}

\section*{Personal Reflection}

\end{document}