### 2009-11-19 great_works/theater_eden.tex

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{}
\author{Matthew Gardner}
\date{}

\begin{document}

\textbf{____ Response}

\begin{tabular}{ll}
  Name:&Matthew Gardner \\
  Work:& \\
  Artist:& \\
  Date:& \\
\end{tabular}

\section*{Historical Context}

\section*{Critical Analysis}

\section*{Personal Reflection}

\end{document}