### 2009-12-03 great_works/science_lda.tex
Finished another Great Works response, a science one on LDA

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{setspace}
\doublespacing

\title{}
\author{Matthew Gardner}
\date{}

\begin{document}

\textbf{Physical Science Response}

\begin{tabular}{ll}
  Name:&Matthew Gardner \\
  Work:&Computer Science Colloquium---Latent Dirichlet Allocation \\
  Speaker:&Charles Elkan \\
  Date:&October 1, 2009 \\
\end{tabular}

\section*{Historical Context}

Though still largely in its infancy, the field of natural language processing
in computer science has flowered in recent years.  Natural language processing
is using a computer to understand, analyze, or otherwise process natural spoken
or written language.  Companies like Google use natural language processing
when trying to understand a search query or when trying to decide whether any
particular email is spam or not.  Machine translation systems also fall under
the heading of natural language processing.  As little as 10 years ago,
attempts to build search engines or machine translation systems were faltering
at best.  Today, the ability of computers to analyze large collections of
documents and perform important tasks with the results of the analysis is very
impressive.  However, a lot of work still needs to be done before we can claim
that computers completely understand natural language.

\section*{In-Depth Analysis}

There are some common tasks that are often done with large collections of
documents.  The first is called classification.  When performing
classification, each document has some kind of label, such as ``spam'' or ``not
spam'' if we are dealing with emails.  The task of classification is to use the
labeled data, often called ``training data,'' to learn some way of
distinguishing between the various labels.  With this method, often a
statistical model, the computer can then decide what label to give to a new
document that was not part of the training documents and does not have a label.
In this way, Google can analyze a lot of emails that have been labeled as being
spam and figure out common characteristics, and use those characteristics to
implement a spam filter.

A slightly more complicated task is that of clustering.  The clustering task
is very similar to the classification task, except it is done without labels.
The computer analyzes all of the documents and groups them into some number of
``clusters,'' where each cluster represents what would have been a label in the
classification task.  Thus without any information except that contained in the
documents themselves, the computer can decide which documents are similar to
each other and which are different.  For example, a researcher could take a
large collection of news articles on different topics and use this method to 
determine which articles talk about the same or similar themes.

A limitation of clustering is that it assumes that each document fundamentally
is about one topic, and tries to figure out which topic each document belongs
to.  What about documents that are about more than one topic, such as faith and
repentance?  If there is a lot of overlap between topics that the documents
talk about, the performance of clustering methods degrades rather quickly.  To
get around this issue, a technique called Latent Dirichlet Allocation, or LDA,
was introduced in 2002.  LDA breaks the assumption that each document comes
from a single topic and instead assumes that each word in each document comes
from a particular topic, allowing documents to contain many topics.  Performing
LDA on a collection of documents will produce a set of topics that contain
similar words and information about how often each document uses words from
each topic.  A good example again is a collection of news articles.  Suppose
the articles talked about Iraq, Vietnam, Health Care, the President of the
United States, and Congress.  Likely, there would be a topic containing words
used in politics, one with words used when talking about war, one about
doctors, and so on.  Each of the articles would have some proportion of words
from each topic, and someone looking for articles that talked about both
politics and health care would have an easy time finding them.

Natural language processing has shown to be very useful when trying to analyze
and perform tasks with large collections of documents.  LDA is the most
advanced and most recently developed of these techniques, and it is still being
improved.  This lecture talked about one particular issue with LDA and how it 
can be addressed.  The problem was that of ``burstiness'' in natural languages,
the fact that if a rare word is used once in a document, the probability of
seeing it again is very high.  Charles Elkan solved the problem by using a
different probability distribution when training his LDA model.  He showed
results that seemed to be promising.

\section*{Personal Reflection}

I first learned about LDA in one of my computer science classes.  This lecture
introduced me to some techniques that I hadn't considered before, and I quite
enjoyed the lecture.  I did a project for a class using LDA where I used it to
analyze the set of talks from LDS General Conferences from 1900 to the present.
The results I saw were fascinating---there was a topic about tithing, one about
welfare, one about Relief Society, one about Sunday School, and many more.  And
the proportions of those topics in the talks over time that LDA found matched
very well with real-world events.  For example, the tithing topic spiked in
1900, when Lorenzo Snow was president of the church and focused heavily on
paying a full tithe, and has since leveled off.  The Relief Society topic was
almost non-existent in the talks until the 1970s, when women started speaking
in General Conference.  I gained an appreciation for the technique through the
experiments I did, and I'm planning on doing graduate work in natural language
processing because of how much I enjoyed using LDA and other algorithms.

\end{document}