### 2010-01-14 great_works/music_brahms.tex
Started work on the last two responses

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{setspace}
\doublespacing

\title{}
\author{Matthew Gardner}
\date{}

\begin{document}

\textbf{Music Response}

\begin{tabular}{ll}
  Name:&Matthew Gardner \\
  Work:&Brahm's First Piano Concerto \\
  Artist:&Johannes Brahms \\
  Date:&1858 \\
\end{tabular}

\section*{Historical Context}

\section*{Critical Analysis}

\section*{Personal Reflection}

\end{document}