### 2010-01-14 great_works/music_brahms.tex
Working on the last response

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{setspace}
\doublespacing

\title{}
\author{Matthew Gardner}
\date{}

\begin{document}

\textbf{Music Response}

\begin{tabular}{ll}
  Name:&Matthew Gardner \\
  Work:&Piano Concerto No. 1 in D Minor, Op. 15 \\
  Artist:&Johannes Brahms \\
  Date:&1858 \\
\end{tabular}

\section*{Historical Context}

Johannes Brahms lived from 1833 to 1897 and was one of the central figures in
the Romantic period of music.  He was heavily influenced by earlier periods,
especially fellow German composers Johann Sebastian Bach and Ludwig van
Beethoven.

\section*{Critical Analysis}

\section*{Personal Reflection}

\end{document}