### 2010-01-25 great_works/faculty_references.tex
A "faculty references" document for my honors portfolio

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{Faculty References}
\author{}
\date{}
\setlength{\parskip}{20pt}

\begin{document}
\maketitle

\begin{tabular}{ll}
  \textbf{Name:}&\textbf{Kevin Seppi} \\
  \emph{Phone:}&\emph{422-4619} \\
  \emph{Office:}&\emph{3330 TMCB} \\
  Classes taken:&CS 236 \\
\end{tabular}

Dr. Seppi is my honors thesis advisor, and I have worked with him as a research
assistant for the past two years.

\begin{tabular}{ll}
  \textbf{Name:}&\textbf{Mike Goodrich} \\
  \emph{Phone:}&\emph{422-6468} \\
  \emph{Office:}&\emph{2220 TMCB} \\
  Classes taken:&CS 252, CS 470, CS 670 \\
\end{tabular}

Dr. Goodrich is my honors thesis referee.

\begin{tabular}{ll}
  \textbf{Name:}&\textbf{Eric Ringger} \\
  \emph{Phone:}&\emph{422-7615} \\
  \emph{Office:}&\emph{3368 TMCB} \\
  Classes taken:&CS 479, CS 679 \\
\end{tabular}

I submitted some work I did with Dr. Ringger to the BYU Studies Symposium, and
some of my work with Dr. Seppi has been in collaboration with Dr. Ringger.

\end{document}