### 2010-06-24 orca/finalreport2009.tex
Added my final report for ORCA for the 2009 grant I was awarded

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{cite}

\title{A Speculative Approach to Parallelization in Particle Swarm Optimization}
\author{Matthew Gardner\\Mentor: Kevin Seppi, Computer Science}
\date{}

\begin{document}
\maketitle

Optimization problems are ubiquitous in our modern world.  Businesses such as
Google need to decide where to place ads, scientists need to fit models to
data, and airlines need to schedule their flights.  All of these problems are
optimization problems, and optimization techniques can be used to find
solutions.  Particle swarm optimization (PSO) is a relatively new optimization
technique that draws on ideas from the sociology of flocking birds.  It was
originally developed for use on a single machine, and attempts at modifying the
algorithm to run in parallel are still quite preliminary.  However, the
problems people want to solve are getting bigger, and larger problems require
more computation power than is available on any single computer.  To
effectively solve those problems, optimization algorithms need to be improved
to better use parallel clusters of machines.  This project addressed that
issue.

Particle swarm optimization was proposed in 1995 by James Kennedy and Russell
Eberhart~\cite{kennedy-1995-particle-swarm-optimization}.  The algorithm is
used to intelligently search a multi-dimensional space by mimicking the
swarming and flocking behavior of birds and other animals. It is a social
algorithm that depends on interaction between particles to quickly and
consistently find the optimal solution to some objective function.

The motion of particles through the search space has three components: an
inertial component that gives particles momentum as they move, a cognitive
component where particles remember the best solution they have found and are
attracted back to that place, and a social component by which particles are
attracted to the best solution that any of their neighbors have found.

Within the last few years, researchers have begun to recognize the need to
develop parallel implementations of PSO.  The methods they have used include
various synchronous algorithms and asynchronous algorithms.  Parallelizing the
evaluation of the objective function can also be done in some cases, though
that is not an adaption of the PSO algorithm and thus was not the focus of this
work.

These previous parallel techniques distribute the computation needed by the
particles in the swarm over the available processors.  If more processors are
available, these techniques increase the number of particles in the swarm,
either by adding individual particles or by adding entire new sub-swarms.  In
almost all cases, adding additional particles produces better results in the
same amount of time.  However, indefinitely increasing the swarm size produces
only incremental benefit; performance improves, but not by very much.  The
reason for this is that there comes a point of diminishing returns with respect
to adding particles.

In this work we considered PSO parallelization strategies for clusters of
hundreds or thousands of processors and functions for which a single evaluation
will take at least tens of seconds, but probably minutes or perhaps hours.  Our
purpose is to explore the question of what to do with a thousand processors
when 50 or 100 particles is the most efficient swarm size, and simply adding
particles results in only incremental improvement.

In previous work (my honors thesis, which was also sponsored by an ORCA grant)
we developed an algorithm we called Speculative Evaluation in PSO.  The
algorithm works by finding all possible positions of particles at future
iterations, then evaluating them in parallel.  That work resulted in a
publication in the proceedings of \emph{Parallel Problem Solving in Nature,} a
prestigious conference in parallelizing nature-based algorithms.

However, that work left us many open questions, and the results we presented
were decent, but not spectacular.  In this work we expanded greatly on the
results of our previous paper, exploring in much more depth the possibilities
of speculative computation in PSO.

Our original algorithm reproduced the PSO algorithm exactly, two iterations at
a time, in parallel, so that the algorithm could be run more quickly on
functions that took a long time to evaluate.  But exactly reproducing PSO
required us to use eight times as many processors as standard PSO, while those
processors could have been used to increase the swarm size of the standard
algorithm.  Our algorithm thus required too much computation to be competitive
in very many instances.

To solve this problem, we relaxed the requirement that our algorithm exactly
reproduce the behavior of standard PSO.  Thus we were able to save some
computation costs and increase the swarm size of our algorithm, and our results
greatly improved.  After introducing some of these cost-saving measures, we
were able to also speculate several iterations ahead, instead of just one, and
that led to further performance improvements in many instances.

The result of this work was an expanded version of our original paper that was
twice as long.  We submitted this more thorough treatment of speculative
evaluation in PSO to the journal \emph{Transactions on Evolutionary
Computation}, one of the premier journals treating research on evolution- and 
nature-inspired algorithms, such as PSO.  We hope to hear a response from the
journal soon.

As is typical of research projects, we were left again with many open
questions.  One of the most intriguing is how to most effectively allocate
speculative evaluations to available processors.  Our approaches treat every
particle equally, giving the same number of extra evaluations to both poorly
performing particles and those that are performing very well.  Perhaps a better
approach would be to draw on ideas from the multi-robot task allocation
literature and dynamically allocate extra processors to particles that could
best make use of them.

\bibliographystyle{plain}
\bibliography{../../bib/pso/general/bib}

\end{document}