### 2010-08-02 textual_criticism/paper/paper.tex
Significant work on the textual criticism paper

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}

\title{An Inquiry into Automated Methods for\\Textual Criticism}
\author{Matthew Gardner}
\date{\today}

\begin{document}
\maketitle

\section{Introduction}
\label{sec:intro}

Textual criticism can be loosely described as the analysis of a series of
manuscripts of a document (generally an ancient document) in an attempt to
reconstruct the original text produced by the author(s) of the document.
Because of the importance of religious documents and the proliferation of their
manuscripts, textual criticism has grown into a very large field, though its
use has not been confined solely to religious documents.  Both the New and Old
Testaments have been widely studied from a text critical perspective, leading
to great revisions in the proposed original texts for those books.  But other,
secular, works, such as the Canterbury Tales, the writings of William
Shakespeare, and a great number of classical texts such as the Illiad and the
Odyssey, have also received significant attention from textual critics.

The process of textual criticism involves laboring over differing manuscripts
searching for discrepancies, recording them by hand and sifting through the
various manuscripts in order to come to some conclusion about the manuscripts'
origins.  After deciding which manuscripts are most trustworthy, the textual
critic produces a text that he or she believes most closely matches the
original text, along with a ``critical apparatus'' containing variant readings
and justification for the editorial desicions made.

For hundreds of years, the methodology used by the textual critic was
painstaking and tedious, often involving travel to several countries to examine
manuscripts that were held by institutions spread throughout the world.  Bruce
Metzger, a prominent New Testament textual critic, described part of the
process as ``such an enormous amount of labor, out of all proportion to the
importance of the results, that scholars are content'' to make textual
decisions considering only a subset of the data available to
them~\cite[179]{metzger-1992-text-of-the-new-testament}.

With the advent of the computer, the work of textual criticism has started to
change.  A computer can now catalogue discrepancies in manuscripts, and even
try to come to a conclusion about which manuscripts were derived from other
manuscripts.  But while some work has been done in automating the process of
textual criticism, there are many areas in which computing resources and ideas
from computer science seem completely untapped.

This work attempts to document the current extent of the use automated methods 
in the field of textual criticism.  We start in \secref{methods} with an
overview of the basic methods of the textual critic, for without a basic
understanding of the methods one cannot proceed to automate them.  We then move
in \secref{previous-work} to a discussion of previous work and the state of the
field in automating textual criticism.  In \secref{proposed-ideas} we discuss
our own ideas for how to automate parts of the process that have not yet been
automated.  \secref{experiments} then presents some preliminary results of
experiments testing some of our ideas, and in \secref{conclusion} we conclude.

\section{Methods of Textual Criticism}
\label{sec:methods}

Much has been written over the years on the methodology involved in the process
of textual criticism.  Two of the foundational works on the subject were
written by Bruce M. Metzger and by Kurt and Barbara
Aland~\cite{metzger-1992-text-of-the-new-testament,
aland-text-of-the-new-testament}.  A more accessible introduction to the
lay-person was written recently by Bart D.
Ehrman~\cite{ehrman-2005-misquoting-jesus}.  His book, however, as it is a
popular and not a scholarly work, is lacking the rigor and depth found in the
books by Metzger and Aland, and he frequently dips into the philosophical
implications of the uncertainty in the textual tradition of the New Testament.

\subsection{Rules for Textual Criticism}

Aland, in his book, gives ``twelve basic rules for textual
criticism''~\cite[280--282]{aland-text-of-the-new-testament}.  Not all of his
rules are relevant to our present discussion, so we will summarize only eight
of them here, giving them new numbers:

\begin{enumerate}
  \item Only one reading can be original (i.e., the author did not produce two
	versions of the same text; an assumption which may in some cases be
	erroneous).
  \item In determining the original text, external evidence should be
	considered before internal evidence (we will discuss each type of evidence
	shortly).
  \item Texts in the original language are the primary authority; translations
	and quotations in other works are of less importance, especially when the
	translation back to the original language is ambiguous.
  \item Manuscripts should be weighed, not counted; the preponderance of a
	particular reading does not necessarily imply originality.
  \item It is unlikely that any one manuscript contains the original text in
	its entirety; more likely is that all manuscripts contain a mix of correct
	and incorrect readings, and only by using all of the evidence can the
	original text be obtained.
  \item The reading which can most easily explain the derivation of other forms
	is most likely the original, so stemmatology (the construction of a family
	tree for the manuscripts) is very important.
  \item The more difficult reading is the more probable reading, as a scribe is
	more likely to make the text easier, not more difficult.
  \item The shorter reading is the more probable reading.
\end{enumerate}

Outside of his rules, Aland also makes the point that, especially in the case
of the New Testament, it is incredibly likely that the original reading for any
variant passage is still intact in some manuscript, because when scribes were
confronted with differing manuscripts they would often just conflate both
readings.  Aland says, 
\begin{quotation}
the element of tenacity in the New Testament textual tradition not only permits
but demands that we proceed on the premise that in every instance of textual
variation it is possible to determine the form of the original text \ldots
assuming a proper understanding of the textual tradition.  The vast number of
witnesses to the text is not simply a burden---it is also a positive
aid.~\cite[294]{aland-text-of-the-new-testament}
\end{quotation}
Thus according to Aland, to obtain the original text we merely
need to list every place of variation and decide which of the extant variants
is the original.  No postulation of missing readings is necessary.

The rules make reference to external and internal evidence.  External evidence
is evidence that is outside of the text of the document itself.  For example,
which manuscripts attest a given reading, the reliability of those manuscripts,
and the age and location of the manuscript are all external criteria.  Internal
evidence includes considerations based on the text of the document, such as
what the author was likely to have written (based on stylistic and theological
considerations), and the probability that a scribe changed the reading from one
variant to another.

\subsection{Kinds of Variation in the Text}

Another important thing to understand in deciding between textual variants is
the manner in which variations creep into the manuscript tradition.  Ehrman
lists several different types of variants, categorized broadly into intentional
and non-intentional variations~\cite[90--98]{ehrman-2005-misquoting-jesus}.
Among the non-intentional kind of variations he lists the following:

\begin{itemize}
  \item spelling errors
  \item periblepsis (skipping some text because of similar line endings or
	beginnings)
  \item misunderstanding of abbreviations used in the text (the various names
	of God, among other words, were often abbreviated in the text)
  \item substitution of homophones (because texts were often copied by
	dictation, so that multiple scribes could copy a manuscript at the same
	time)
  \item changes in word order.  
\end{itemize}

His list of intentional variations includes:

\begin{itemize}
  \item ``fixing'' factual errors in the text (we put ``fixing'' in quotation
	marks, because the scribe may think he is correcting the document, but he
	is in fact changing the original text; the textual critic is interested in
	the original text, whether or not there were factual errors in it)
  \item ``fixing'' doctrinal or interpretive errors
  \item ``harmonizing'' similar passages in the text (this is especially
	applicable to the Gospels in the New Testament textual tradition)
\end{itemize}

\subsection{Knowledge Sources Needed}

In order to obtain the evidence required for determining the original reading
of a passage, there are several sources of knowledge that must be consulted.
If we are to automate the process of textual criticism, those knowledge sources
must be available to and interpretable by a computer.  Thus we discuss here all
of the kinds of knowledge that are necessary to the work of the textual critic.

The most basic knowledge source is the manuscript itself.  Each manuscript has
a date, a location, and text.  The location of the manuscript is trivially
determined upon its discovery, and the date is also fairly easily obtained by
archaeologists.  The text of the manuscript may be in dispute, if the
manuscript is incredibly corrupted or is incomplete, or if the text has been
written over, as in some of our oldest manuscripts (the uncial manuscript known
as 04 or C is called the ``Codex Ephraemi Syri Rescriptus,'' where the word
``rescriptus'' is added for that very
reason~\cite[109]{aland-text-of-the-new-testament}).  As far as computers
performing the work of textual criticism, it seems reasonable to assume a
faithful transcription for each manuscript, with the date and location as
metadata.  This knowledge source, then, contains almost all of the information
listed as ``external evidence.''

The one part of the external evidence not immediately apparent from the above
description of the manuscript is the reliability of the manuscript.  A
manuscript's reliability is determined by how often it contains correct
readings.  But at the same time, correct readings are often established in part
by the reliability of the manuscripts that contain them.  Clearly the process
is circular, or, as the world of computer science would put it, iterative.
Other knowledge sources must be used in order to establish the reliability of
manuscripts, and establishing reliability in an automated fashion must be an
iterative process.

Another item that should be mentioned is that manuscripts were often corrected
by a hand not belonging to the original scribe.  Those corrections provide more
evidence that can be used in textual criticism.  One must decide how to encode
that evidence.  The approach taken by all critical apparati that we observed
was to superscript the symbol for the corrected manuscript with a ``c'' for the
corrected version, and an asterisk for the original.  The two versions of the
manuscript could then be used as two separate manuscripts in an automated
system, though care would have to be taken to not double count their evidence
in passages where there was no correction.

The knowledge sources needed to judge the ``internal'' criteria are much more
complicated.  Determining what the author was likely to have written is not a
trivial task.  One must have a grasp of the author's writing style, vocabulary,
and theology, none of which can be ascertained perfectly, though various
approximations can and have been made.  A computer would have to make far more
drastic and unreasonable approximations than those a human makes, though some
parts of the knowledge source seem at least approachable in an automated
fashion, such as style and vocabulary.

Determining the probability with which a scribe introduces a variant reading
requires further knowledge, including an understanding of the way scribes
operated (which is why we included a section on the kinds of variations that
exist in the text).  Much of this information could be encoded into an
automated system without too much difficulty by simply specifying a prior
probability of each type of error occurring.  However, when considering whether
a scribe introduced a variant, the concept of an ``easier'' reading is referred
to, for scribes are more likely to make the reading easier than more difficult.
Those types of considerations seem at present beyond the capacity of automated
methods.

There are also other kinds of knowledge sources involved in the process of
textual criticism that are more analogous to algorithms than to data in the
world of computer science.  For example, knowing the history of the manuscript
tradition leads to decisions to model the manuscripts as a family tree, similar
to genetic evolution.  Such knowledge sources are encoded into an automated
system in the way the system is built.

Lastly, if one wants to use quotations of the text in writings of other
authors, or translations of the work into other languages, still more knowledge
sources are required.  For the simple case of a quotation, the extract text as
quoted could be considered its own ``manuscript.''  To use other-language
texts, translation back into the original is necessary, or some other way to
use the foreign-language text to aid in the decision between variant readings.
Perhaps the use of a bilingual dictionary would be useful.  It seems more
difficult, however, to propose a new variant reading from a translated version
of the text.

\section{Previous Work in Automating Textual Criticism}
\label{sec:previous-work}

An early survey of the history of computers in textual criticism was published
in 1994 by Robert
Kraft~\cite{kraft-1994-computers-in-new-testament-textual-criticism}.  For the
reader interested in the history, his is a good survey to read.  However, in
our current discussion we are interested in the present use of computers in
textual criticism, so we will defer further description of the history of the
field to that survey.  One line, however, seems germaine to the topic at hand:
``We have, in short, not come very far in realizing the promise offered by
computer-assisted research for NT textual criticism, although much of the
groundwork has been laid.''

Since 1994 significant progress has been made, though the vast majority of that
work has been in one small part of the process of textual criticism, that of
trying to produce a hierarchical relationship between each of the manuscripts
of a given text.  There is a very broad literature on the topic of constructing
trees of manuscripts, normally called either stemmatology or cladistics.  For
an example of how broad the field is, see
\cite{spencer-2002-exploring-textual-genealogy} and
\cite{van-reenen-2004-studies-in-stemmatology-ii}, including the articles they
cite.

The purpose of research into stemmatology seems to be to establish the
reliability of each manuscript.  If one can establish which manuscripts are
earliest in the tree, one can come to some conclusion about which manuscripts
are the most reliable.  A survey of the current state of the art in manuscript
stemmatology is beyond the scope of this paper; we suffice ourselves in noting
that significant research has been done in that area and to which aspects of
textual criticism the field applies.

There has also been some investigation into using clustering methods to group
manuscripts by similar readings~\cite{willker-pca-of-manuscripts-of-john}.  One
article we found even uses hierachical clustering to form trees of documents,
instead of the more common techniques taken from computational
biology~\cite{thorpe-2002-statistical-manuscript-classification}.  These
methods seem to reproduce rather well the ``families'' of manuscripts created
by textual critics to describe sections of the New Testament manuscript
tradition (for a description of the manuscript families, see
\cite[212--218]{metzger-1992-text-of-the-new-testament}).

A related, though somewhat distinct, use of computers in New Testament studies
is the attempt to determine authorship of documents using automated stylometry.
Such methods build some kind of model for the style of each particular author,
then try to attribute disputed documents to one of the authors in question.
\cite{alviar-2008-computational-linguistics-and-biblical-studies} is an example
of such a study.  The relation to textual criticism comes because stylistic
considerations are often involved in the determination of an original reading.
One would simply take the models that these methods produce and judge between
two variant readings, instead of entire books.

As far as we could tell, all previous work started with the digitized text of a
\emph{critical apparatus}, none of which are completely exhaustive, not the
text of the manuscripts themselves.  Thus all previous studies have used only a
subset of the possible data that they could have used; however, it is arguable
that the data that was not included would not have a significant effect on the
results of the algorithm.

The reason that previous work has not used the actual transcriptions of the
manuscripts is that such transcriptions are not widely available, and in many
cases do not exist in digital form at all.  There have been several calls for
the digitization of manuscripts, at least in image form, though none of them
have as yet been successful~\cite{trobisch-2000-central-electronic-database}.
One particularly eloquent summary comes from a 1999 paper by L. W. Hurtado.  He
says,
\begin{quotation}
To date, major obstacles remain in the way of these aims [the use of computers
in textual criticism], especially the necessity to transcribe the witnesses
into machine-readable form (a highly labour-intensive process and fraught with
the danger of introducing additional variants inherent in any copying process!)
and the need for fully adequate software able to collate and in some level
classify variants.

\ldots

Any significant uses of computer technology for the work of textual criticism
of the New Testament, uses that would materially simplify, expedite and improve
upon more traditional measures, are still unfortunately the stuff of
dreams.~\cite{hurtado-1999-beyond-the-interlude}
\end{quotation}

In his paper, Hurtado mentions the ``Electronic New Testament Manuscript
Project,'' an effort to collect digital images and transcriptions of all New
Testament manuscripts, as particularly hopeful.  It was in its infancy in 1999
when Hurtado published his paper, and it did not survive past infancy; hardly a
trace of the effort can be found today.

However, in their book, \emph{Jesus Christ and the World of the New Testament},
Holzapfel, Wayment and Huntsman mention that the majority of New Testament
manuscripts would be transcribed and available online by the end of
2010~\cite[?]{holzapfel-2006-jesus-christ-new-testament}.  Our extensive
searching produced only the uncial manuscripts for the Gospel of John, in a
format that was searchable, though not readily
downloadable~\cite{igntp-gospel-of-john-online}.  The project behind that
website, the International Greek New Testament Project (www.igntp.org), aims to
create a critical edition of the New Testament, and the publishing of those
manuscripts online seems to be merely incidental to their work.

Upon contacting Thomas Wayment in reference to his statement
in~\cite{holzapfel-2006-jesus-christ-new-testament}, we were informed that
BibleWorks, a company that has long produced software for those interested in
biblical studies, is in the process of digitizing the manuscripts.  A thread on
their user forums was of particular interest to the present
discussion~\cite{bible-works-forum}.  In the thread, Michael Bushell, the owner
of and lead programmer for the company, confirmed Wayment's statement that they
are in the process of digitizing the manuscripts, and that they hope to have
some progress to show by the end of 2010.  However, he says that the
transcriptions will only be made available with the purchase of BibleWorks
software initially, then ``eventually'' to ``organizations'' who will not
profit from the transcriptions.

Also in that thread we found the closest thing to what we are currently
investigating.  Twice Bushell mentions software that could ``generate apparati
on the fly,'' though in both instances he is doubtful that he will live to see
such software.  It would appear that at least one other person has the idea of
completely automating the process of textual criticism.  We found that idea to
be rather unique among everything we read; other authors spoke of automating
some part of the process, but nowhere else did we find someone even considering
generating an entire critical apparatus automatically.  Most authors seemed
skeptical of automated methods at best.  Future work in this area should
certainly involve contact with Michael Bushell.

\section{Proposed Ideas for the Automation of Textual Criticism}
\label{sec:proposed-ideas}

\section{Experiments}
\label{sec:experiments}

Our experiments involved sections of manuscripts of the Gospel of John
downloaded from the internet, at this site~\cite{igntp-gospel-of-john-online}
mentioned previously.  We aligned the text using a piece of software written
for aligning DNA and protein sequences in computational
biology~\cite{lee-2002-sequence-alignment-poa}.  This aligner performs a
character-level alignment, showing insertions, substitutions, and deletions for
each of the manuscripts.

An important consideration in the use of the aligner is to keep track of the
difference between an actual shorter reading in a text (perhaps because of a
later scribal addition) and a deficient manuscript that is simply missing
entire passages (because only part of the manuscript was found, for instance).
In the latter circumstance the manuscript is said to have lacunae (sing.
lacuna), or gaps.  One does not want to consider a lacuna to be evidence for a
shorter reading, nor ignore a legitimately shorter reading as if it were a
lacuna.

Given an alignment of the text, we can find locations where not all of the
manuscripts agree.  The aligner only gives character-level differences, but we
can easily extend the differences to word boundaries, and connect adjacent
differing blocks.  Upon performing this post-processing, we discovered 181
passages in the first chapter of John alone that contained variant readings,
consisting of 40\% of the text, with just nine manuscripts as evidence.  The
vast majority of these differences were a spelling error or word order change
introduced into one of the nine manuscripts, though some were more significant.

\section{Conclusion}
\label{sec:conclusion}

\bibliographystyle{plain}
\bibliography{../../../bib/textual-criticism/bib}

\end{document}