### 2010-09-27 textual_criticism/paper/paper.tex
Adding the final revision to the textual criticism paper that I never committed

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}

\title{An Inquiry into Automated Methods for\\Textual Criticism}
\author{Matthew Gardner}
\date{\today}

\begin{document}
\maketitle

\section{Introduction}
\label{sec:intro}

Textual criticism can be loosely described as the analysis of a series of
manuscripts of a document (generally an ancient document) in an attempt to
reconstruct the original text produced by the author(s) of the document.
Because of the importance of religious documents and the proliferation of their
manuscripts, textual criticism has grown into a very large field, though its
use has not been confined solely to religious documents.  Both the New and Old
Testaments have been widely studied from a text critical perspective, leading
to great revisions in the proposed original texts for those books.  But other,
secular, works, such as the Canterbury Tales, the writings of William
Shakespeare, and a great number of classical texts such as the Illiad and the
Odyssey, have also received significant attention from textual critics.

The process of textual criticism involves laboring over differing manuscripts
searching for discrepancies, recording them by hand and sifting through the
various manuscripts in order to come to some conclusion about the manuscripts'
origins.  After deciding which manuscripts are most trustworthy, the textual
critic produces a text that he or she believes most closely matches the
original text, along with a ``critical apparatus'' containing variant readings
and justification for the editorial decisions made.

For hundreds of years, the methodology used by the textual critic was
painstaking and tedious, often involving travel to several countries to examine
manuscripts that were held by institutions spread throughout the world.  Bruce
Metzger, a prominent New Testament textual critic, described part of the
process as ``such an enormous amount of labor, out of all proportion to the
importance of the results, that scholars are content'' to make textual
decisions considering only a subset of the data available to
them~\cite[p. 179]{metzger-1992-text-of-the-new-testament}.

With the advent of the computer, the work of textual criticism has started to
change.  A computer can now catalogue discrepancies in manuscripts, and even
try to come to a conclusion about which manuscripts were derived from other
manuscripts.  But while some work has been done in automating the process of
textual criticism, there are many areas in which computing resources and ideas
from computer science seem completely untapped.

This work attempts to document the current extent of the use automated methods
in the field of textual criticism.  I start in \secref{methods} with an
overview of the basic methods of the textual critic, for without a basic
understanding of the methods one cannot proceed to automate them.  I then move
in \secref{previous-work} to a discussion of previous work and the state of the
field in automating textual criticism.  In \secref{proposed-ideas} I discuss my
own ideas for how to automate parts of the process that have not yet been
automated, and some preliminary results of experiments testing some of my
ideas.  In \secref{conclusion} I conclude.

\section{Methods of Textual Criticism}
\label{sec:methods}

Much has been written over the years on the methodology involved in the process
of textual criticism.  Two of the foundational works on the subject were
written by Bruce M. Metzger and by Kurt and Barbara
Aland~\cite{metzger-1992-text-of-the-new-testament,
aland-text-of-the-new-testament}.  A more accessible introduction to the
lay-person was written recently by Bart D.
Ehrman~\cite{ehrman-2005-misquoting-jesus}.  His book, however, as it is a
popular and not a scholarly work, is lacking the rigor and depth found in the
books by Metzger and Aland, and he frequently dips into the philosophical
implications of the uncertainty in the textual tradition of the New Testament.

\subsection{Rules for Textual Criticism}

Aland, in his book, gives ``twelve basic rules for textual
criticism''~\cite[pp. 280--282]{aland-text-of-the-new-testament}.  Not all of
his rules are relevant to the present discussion, so I will summarize only
eight of them here, giving them new numbers:

\begin{enumerate}
  \item Only one reading can be original (i.e., the author did not produce two
	versions of the same text; an assumption which may in some cases be
	erroneous).
  \item In determining the original text, external evidence should be
	considered before internal evidence (I will discuss each type of evidence
	shortly).
  \item Texts in the original language are the primary authority; translations
	and quotations in other works are of less importance, especially when the
	translation back to the original language is ambiguous.
  \item Manuscripts should be weighed, not counted; the preponderance of a
	particular reading does not necessarily imply originality.
  \item It is unlikely that any one manuscript contains the original text in
	its entirety; more likely is that all manuscripts contain a mix of correct
	and incorrect readings, and only by using all of the evidence can the
	original text be obtained.
  \item The reading which can most easily explain the derivation of other forms
	is most likely the original, so stemmatology (the construction of a family
	tree for the manuscripts) is very important.
  \item The more difficult reading is the more probable reading, as a scribe is
	more likely to make the text easier, not more difficult.
  \item The shorter reading is the more probable reading.
\end{enumerate}

Outside of his rules, Aland also makes the point that, especially in the case
of the New Testament, it is incredibly likely that the original reading for any
variant passage is still intact in some manuscript, because when scribes were
confronted with differing manuscripts they would often just conflate both
readings.  Aland says, 
\begin{quotation}
the element of tenacity in the New Testament textual tradition not only permits
but demands that we proceed on the premise that in every instance of textual
variation it is possible to determine the form of the original text \ldots
assuming a proper understanding of the textual tradition.  The vast number of
witnesses to the text is not simply a burden---it is also a positive
aid.~\cite[p. 294]{aland-text-of-the-new-testament}
\end{quotation}
Thus according to Aland, to obtain the original text one merely
needs to list every place of variation and decide which of the extant variants
is the original.  No postulation of missing readings is necessary.

The rules make reference to external and internal evidence.  External evidence
is evidence that is outside of the text of the manuscript itself.  For example,
which manuscripts attest a given reading, the reliability of those manuscripts,
and the age and location of the manuscript are all external criteria.  Internal
evidence includes considerations based on the text of the document (both the
manuscript in question and the likely original text), such as what the author
was likely to have written (based on stylistic and theological considerations),
and the probability that a scribe changed the reading from one variant to
another.

It should also be noted that textual critics often use the word ``probability''
when describing their rules, such as in saying that ``the shorter reading is
more probable,'' or ``more likely.''  When they use those terms, however, they
do not refer to actual probabilities in the strict, mathematical sense of the
word.  In building an automated system, it might be fruitful to build a
statistical model using the rules mentioned here as guidelines, and then the
nebulous use of the word ``probability'' would be translated into something
more concrete.

\subsection{Kinds of Variation in the Text}

Another important thing to understand in deciding between textual variants is
the manner in which variations creep into the manuscript tradition.  Ehrman
lists several different types of variants, categorized broadly into intentional
and non-intentional variations~\cite[pp. 90--98]{ehrman-2005-misquoting-jesus}.
Among the non-intentional kind of variations he lists the following:

\begin{itemize}
  \item spelling errors
  \item periblepsis (skipping some text because of similar line endings or
	beginnings)
  \item misunderstanding of abbreviations used in the text (the various names
	of God, among other words, were often abbreviated in the text)
  \item substitution of homophones (because texts were often copied by
	dictation, so that multiple scribes could copy a manuscript at the same
	time)
  \item changes in word order.  
\end{itemize}

His list of intentional variations includes:

\begin{itemize}
  \item ``fixing'' factual errors in the text (I put ``fixing'' in quotation
	marks because the scribe may think he is correcting or improving the text
	of the manuscript he is copying, but he is in fact changing the original
	text; the textual critic is interested in the original text, whether or not
	there were factual errors in it)
  \item ``fixing'' doctrinal or interpretive errors
  \item ``harmonizing'' similar passages in the text (this is especially
	applicable to the Gospels in the New Testament textual tradition)
\end{itemize}

\subsection{Knowledge Sources Needed}

In order to obtain the evidence required for determining the original reading
of a passage, there are several sources of knowledge that must be consulted.
If one is to automate the process of textual criticism, those knowledge sources
must be available to and interpretable by a computer.  Thus I discuss here all
of the kinds of knowledge that are necessary to the work of the textual critic.

\subsubsection{Sources of ``External'' Evidence}

The most basic knowledge source is the manuscript itself.  Each manuscript has
a date, a location, and text.  The location of the manuscript is trivially
determined upon its discovery, and the date is also fairly easily obtained by
archaeologists.  The text of the manuscript may be in dispute, if the
manuscript is incredibly corrupted or is incomplete, or if the text has been
written over, as in some of our oldest manuscripts (the uncial manuscript known
as 04 or C is called the ``Codex Ephraemi Syri Rescriptus,'' where the word
``rescriptus'' is added for that very reason~\cite[p.
109]{aland-text-of-the-new-testament}).  As far as computers performing the
work of textual criticism, it seems reasonable to assume a faithful
transcription for each manuscript, with the date and location as meta data.
This knowledge source, then, contains almost all of the information listed as
``external evidence.''

The one part of the external evidence not immediately apparent from the above
description of the manuscript is the reliability of the manuscript.  A
manuscript's reliability is determined by how often it contains correct
readings.  But at the same time, correct readings are often established in part
by the reliability of the manuscripts that contain them.  Clearly the process
is circular, or, as the world of computer science would put it, iterative.
Other knowledge sources must be used in order to establish the reliability of
manuscripts, and establishing reliability in an automated fashion must be an
iterative process.

Another item that should be mentioned is that manuscripts were often corrected
by a hand not belonging to the original scribe.  Those corrections provide more
evidence that can be used in textual criticism.  One must decide how to encode
that evidence.  The approach taken by all critical apparati that I observed
was to superscript the symbol for the corrected manuscript with a ``c'' for the
corrected version, and an asterisk for the original.  The two versions of the
manuscript could then be used as two separate manuscripts in an automated
system, though care would have to be taken to not double count their evidence
in passages where there was no correction.

\subsubsection{Sources of ``Internal'' Evidence}

The knowledge sources needed to judge the ``internal'' criteria are much more
complicated.  Determining what the author was likely to have written is not a
trivial task.  One must have a grasp of the author's writing style, vocabulary,
and theology, none of which can be ascertained perfectly, though various
approximations can and have been made.  A computer would have to make far more
drastic and unreasonable approximations than those a human makes, though some
parts of the knowledge source seem at least approachable in an automated
fashion, such as style and vocabulary.

A possibility for using computers to make theological decisions is to create
extensive markup about the theology contained in a text, so the computer would
not have to automatically detect the theology.  Then the computer could decide
if any particular variant agreed theologically with the rest of the text in
question.  However, determining what kind of markup is sufficient and creating
the markup for every text (and possibly every manuscript of the text) seems
like far more work than simply having the human make the theological decision
himself.

Determining the probability with which a scribe introduces a variant reading
requires further knowledge, including an understanding of the way scribes
operated (which is why I included a section on the kinds of variations that
exist in the text).  Much of this information could be encoded for processing
by an automated system without too much difficulty by simply specifying a prior
probability of each type of error occurring.  However, when considering whether
a scribe introduced a variant, the concept of an ``easier'' reading is referred
to, for scribes are more likely to make the reading easier than more difficult.
Those types of considerations seem at present beyond the capacity of automated
methods.

For instance, a prime example of a scribe making the reading easier is found in
Mark 1:41~\cite[pp. 133--139]{ehrman-2005-misquoting-jesus}.  In the passage, a
leper comes to Jesus seeking to be healed.  The majority of manuscripts
available to us today say that Jesus had compassion on the man and healed him.
However, there are some significant, very old manuscripts that instead say that
Jesus became angry with the man, then healed him.  It is very easy to
understand that Jesus had compassion, but much more difficult to explain anger.
It is much more likely, then, the textual critic would say, that a scribe
changed Jesus from being angry to showing compassion, not the other way around.
How to get a computer to understand and use the knowledge required to make
these judgments is beyond the current understanding of natural language
processing.

\subsubsection{Other Knowledge Sources}

There are also other kinds of knowledge sources involved in the process of
textual criticism that are more analogous to algorithms than to data in the
world of computer science.  For example, knowing the history of the manuscript
tradition leads to decisions to model the manuscripts as a family tree, similar
to genetic evolution.  Such knowledge sources are encoded into an automated
system in the way the system is built.

Lastly, if one wants to use quotations of the text in writings of other
authors, or translations of the work into other languages, still more knowledge
sources are required.  For the simple case of a quotation, the extracted text
as quoted could be considered its own ``manuscript,'' though certainly it
should hold less weight than a real manuscript.  To use other-language texts,
translation back into the original is necessary, or some other way to use the
foreign-language text to aid in the decision between variant readings.  Perhaps
the use of a bilingual dictionary would be useful.  It seems more difficult,
however, to propose a new variant reading from a translated version of the
text.

\section{Previous Work in Automating Textual Criticism}
\label{sec:previous-work}

An early survey of the history of computers in textual criticism was published
in 1994 by Robert
Kraft~\cite{kraft-1994-computers-in-new-testament-textual-criticism}.  For the
reader interested in the history, his is a good survey to read.  However, in
the current discussion I am interested in the present use of computers in
textual criticism, so I will defer further description of the history of the
field to that survey.  One line, however, seems germane to the topic at hand:
``We have, in short, not come very far in realizing the promise offered by
computer-assisted research for NT textual criticism, although much of the
groundwork has been laid.''

Since 1994 significant progress has been made, though the vast majority of that
work has been in one small part of the process of textual criticism, that of
trying to produce a hierarchical relationship among the manuscripts of a given
text.  There is a very broad literature on the topic of constructing trees of
manuscripts, normally called either stemmatology or cladistics.  For an example
of how broad the field is, see \cite{spencer-2002-exploring-textual-genealogy}
and \cite{van-reenen-2004-studies-in-stemmatology-ii}, including the articles
they cite.

The purpose of research into stemmatology seems to be to establish the
reliability of each manuscript.  If one can establish which manuscripts are
earliest in the tree, one can come to some conclusion about which manuscripts
are the most reliable.  It seems that the method at the forefront of New
Testament stemmatology is called the ``Coherence-Based Genealogical Method,''
presented in detail in an essay called ``Problems of a highly contaminated
tradition: the New Testament: Stemmata of variants as a source of a genealogy
for witnesses''~\cite[pp. 13--87]{van-reenen-2004-studies-in-stemmatology-ii}.
This method is currently being used by the Institute for New Testament Textual
Research at the University of M\"{u}nster to aid in the production of the text
for the Editio Critica Maior (ECM), an important critical edition of the New
Testament.

The general idea of the method is that in a single manuscript the variant
readings could have different ages of origin.  Thus individual variants are
treated separately, and stemmatological decisions are made on the basis of an
understanding of how the variants came about.  The paper also recognizes the
circular nature of the reliability of a manuscript, and uses an iterative
method to establish that reliability~\cite[p.
25]{van-reenen-2004-studies-in-stemmatology-ii}.  One of the key insights that
can be gained from this method is that a manuscript perhaps should not have a
single measure of reliability; different sections of the same manuscript can
have different levels of reliability.

There has been much more work in the field of New Testament stemmatology than
we have room or time to present here.  A broad and detailed survey of the field
could easily be longer than the current study.  It seems sufficient to
mention one of the most prominent methods in the field and move on to other
areas of work in the automation of textual criticism.

There has also been some investigation into using clustering methods to group
manuscripts by similar readings~\cite{willker-pca-of-manuscripts-of-john}.  One
article I found even uses hierarchical clustering to form trees of documents,
instead of the more common techniques taken from computational
biology~\cite{thorpe-2002-statistical-manuscript-classification}.  These
methods seem to reproduce rather well the ``families'' of manuscripts created
by textual critics to describe sections of the New Testament manuscript
tradition (for a description of the manuscript families, see \cite[pp.
212--218]{metzger-1992-text-of-the-new-testament}).

A related, though somewhat distinct, use of computers in New Testament studies
is the attempt to determine authorship of documents using automated stylometry.
Such methods build some kind of model for the style of each particular author,
then try to attribute disputed documents to one of the authors in question.
\cite{alviar-2008-computational-linguistics-and-biblical-studies} is an example
of such a study.  The relation to textual criticism comes because stylistic
considerations are often involved in the determination of an original reading.
One would simply take the models that these methods produce and judge between
two variant readings, instead of entire books.

As far as I could tell, all previous work started with the digitized text of a
\emph{critical apparatus}, none of which are completely exhaustive, not the
text of the manuscripts themselves.  Thus all previous studies have used only a
subset of the possible data that they could have used; however, it is arguable
that the data that was not included would not have a significant effect on the
results of the algorithm.

The reason that previous work has not used the actual transcriptions of the
manuscripts is that such transcriptions are not widely available, and in many
cases do not exist in digital form at all.  There have been several calls for
the digitization of manuscripts, at least in image form, though none of them
have as yet been successful~\cite{trobisch-2000-central-electronic-database}.
One particularly eloquent summary comes from a 1999 paper by L. W. Hurtado.  He
says,
\begin{quotation}
To date, major obstacles remain in the way of these aims [the use of computers
in textual criticism], especially the necessity to transcribe the witnesses
into machine-readable form (a highly labour-intensive process and fraught with
the danger of introducing additional variants inherent in any copying process!)
and the need for fully adequate software able to collate and in some level
classify variants.

\ldots

Any significant uses of computer technology for the work of textual criticism
of the New Testament, uses that would materially simplify, expedite and improve
upon more traditional measures, are still unfortunately the stuff of
dreams.~\cite{hurtado-1999-beyond-the-interlude}
\end{quotation}

In his paper, Hurtado mentions the ``Electronic New Testament Manuscript
Project,'' an effort to collect digital images and transcriptions of all New
Testament manuscripts, as particularly hopeful.  It was in its infancy in 1999
when Hurtado published his paper, and it did not survive past infancy; hardly a
trace of the effort can be found today.

However, in their book, \emph{Jesus Christ and the World of the New Testament},
Holzapfel, Wayment and Huntsman mention that the majority of New Testament
manuscripts would be transcribed and available online by the end of
2010~\cite[p. 55]{holzapfel-2006-jesus-christ-new-testament}.  My extensive
searching produced only two websites, both of which were difficult to use.  One
contained the uncial manuscripts for the Gospel of John, in a format that was
searchable, though not readily downloadable~\cite{igntp-gospel-of-john-online}.
The project behind that website, the International Greek New Testament Project
(www.igntp.org), aims to create a critical edition of the New Testament, and
the publishing of those manuscripts online seems to be merely incidental to
their work.  The other website was that of the The Institute for New Testament
Research at the University of M\"{u}nster~\cite{munster-website}.   The site
contains a large number of manuscripts, though not nearly all of them, and it
is similarly difficult to download anything in bulk.  The website also states
the intention of the institute to make available transcriptions of all of the
manuscripts for which they have photographs, which is nearly all in existence.

Upon contacting Thomas Wayment in reference to his statement
in~\cite{holzapfel-2006-jesus-christ-new-testament}, I was informed that
BibleWorks, a company that has long produced software for those interested in
biblical studies, is in the process of digitizing the manuscripts.  A thread on
their user forums was of particular interest to the present
discussion~\cite{bible-works-forum}.  In the thread, Michael Bushell, the owner
of and lead programmer for the company, confirmed Wayment's statement that they
are in the process of digitizing the manuscripts, and that they hope to have
some progress to show by the end of 2010.  However, he says that the
transcriptions will only be made available with the purchase of BibleWorks
software initially, then ``eventually'' to ``organizations'' who will not
profit from the transcriptions.

Also in that thread I found the closest thing to what I am currently
investigating.  Twice Bushell mentions software that could ``generate apparati
on the fly,'' though in both instances he is doubtful that he will live to see
such software.  It would appear that at least one other person has the idea of
completely automating the process of textual criticism.  I found that idea to
be rather unique among everything I read; other authors spoke of automating
some part of the process, but nowhere else did I find someone even considering
generating an entire critical apparatus automatically.  Most authors seemed
skeptical of automated methods at best.  Future work in this area should
certainly involve contact with Michael Bushell.

\section{Proposed Ideas and Experiments}
\label{sec:proposed-ideas}

\subsection{Generating Critical Apparati}

Michael Bushell mentioned a software tool that would generate a critical
apparatus given some user-specified criteria.  Through some experimentation, I
have concluded that this software is trivial to write if the ``user-specified
criteria'' includes a critical text.  Using out-of-the-box alignment
software\footnote{The aligner I used was written for DNA and protein sequences,
but worked reasonably enough with Greek text represented using Beta Code.  The
aligner was the result of a paper published in \emph{Bioinformatics}, and code
is freely available on the internet~\cite{lee-2002-sequence-alignment-poa}.  I
am indebted to Scott Lloyd, a PhD student in computer science at BYU, for
pointing me to this software.  The aligner performs a character-level
alignment, showing insertions, substitutions and deletions for each of the
manuscripts.}, in a matter of days I was able to produce code that is the basis
for generating a critical apparatus.  From an alignment of the manuscripts, one
can easily find passages that differ across the manuscripts, and if one of them
is labeled as the critical text to use, all that remains to do is format the
output in some recognizable form (i.e., output the critical text with footnotes
that describe which manuscripts contain variant readings).

An important consideration in the use of the aligner is to keep track of the
difference between an actual shorter reading in a text (perhaps because of a
later scribal addition) and a deficient manuscript that is simply missing
entire passages (because only part of the manuscript was found, for instance).
In the latter circumstance the manuscript is said to have lacunae (sing.
lacuna), or gaps.  One does not want to consider a lacuna to be evidence for a
shorter reading, nor ignore a legitimately shorter reading as if it were a
lacuna.

The output of the alignment software looked as follows\footnote{In using the
software I substituted underscores for spaces, as the software did not allow
spaces in the sequence.  I also removed all punctuation and diacritical marks,
except those indicating abbreviations.  The spaces are restored here for easier
interpretation of the output}:

\begin{quotation}
{\tt
john/005.txt        ---------------------------------------------------

john/003.txt        o pemyas me baptizein en ---udati ekeinos moi eipe-

john/001.txt        o pemyas me baptizein en tw udati ek-inos moi eipen

john/009.txt        o pemyas me baptizein en ---udati ekeinos moi eipen

john/007.txt        o pemyas me baptizein en ---udati ekeinos moi eipen

john/004.txt        o pemyas me baptizein en ---udati ekeinos moi eipen

john/013.txt        o pemyas me baptizein en ---udati ekeinos moi eipen

john/011.txt        o pemyas me baptizein en ---udati ekeinos moi eipen

john/002.txt        o pemyas me baptizein en ---udati ekeinos moi eipen
}
\end{quotation}

This is part of the text for John 1:33, translating roughly to ``he who sent me
to baptize in water, that man said to me.''  The first column in each line is
the name of the sequence that was given to the aligner; I just used the
filename, where the filename was the number commonly given to the manuscript.
Hyphens indicate a deletion in a manuscript, or an insertion in one of the
other manuscripts.  In this section of text, we can see that manuscript 5 had a
lacuna (the marker that I used for lacunae is found prior to this section,
however), manuscript 1 had a word addition (``the water'' instead of ``water'')
and misspelled \emph{ekeinos} (``that man''), manuscript 3 left out a nu on
\emph{eipen} (``he said''), and all of the other manuscripts agreed on the
text.

Given this output, I wrote code that parsed it and presented lists of variant
passages along with which manuscripts attested which readings.  An example
passage follows.

\begin{quotation}
{\tt
Location: (4731, 4737)

uios: [`john/003.txt', `john/002.txt']

us$\sim$: [`john/001.txt', `john/009.txt', `john/007.txt', `john/013.txt']

xs$\sim$: [`john/011.txt'] 
}
\end{quotation}

The location shows the character positions this reading spanned (it includes
the spaces surrounding the word; thus there are six characters in the reading
instead of four).  The $\sim$ symbol indicates that an abbreviation mark was
found in the manuscript.  \emph{uios}, meaning ``son,'' was abbreviated as
\emph{us} with an over-bar.  The abbreviation \emph{xs} means ``Christ.''  Thus
manuscript 11 replaced the word son with the word Christ, and all other
manuscripts that do not have a lacuna in this area contain the same reading,
though some are abbreviated.  This particular substitution is not very
surprising, because in capital Greek characters it is the change from a
character much like our ``Y'' to a character much like our ``X''---a very small
difference in appearance and a similarly small change in meaning.

If one of the manuscripts were marked as the critical text to use, from this
list of variants it is a simple task to output the critical text and mark all
other readings as footnotes in a critical apparatus.

It seems, then, that Michael Bushell's hope of automatically generated critical
apparati is not very difficult, as I cannot imagine that he also intends to
generate the critical text on the fly, unless there is some very serious markup
added to the variant readings (including information on theology).  The main
obstacle is the transcription of the manuscripts.  The main benefit of such an
automatically generated apparatus is that it can be completely exhaustive.

\subsection{Generating a Critical Text}

If one could generate a critical text in a completely automated fashion, I have
already shown that the critical text can be used to easily construct an
apparatus.  Thus the work of textual criticism would be complete, with no more
need for textual critics.  However, given the incredibly complicated nature of
some of the knowledge sources needed for the work of textual criticism, it is
extremely unlikely that a reliable critical text will be automatically
generated any time in the foreseeable future.  As Dr. Eric Ringger, an
assistant professor of computer science at BYU, put it, determining whether a
passage matches the theology of an author is ``AI complete''---solving that
problem amounts to solving all of artificial intelligence.  But we can make
some faltering attempts.

My experiments involved sections of manuscripts of the Gospel of John
downloaded from the internet, at the site previously
mentioned~\cite{igntp-gospel-of-john-online}.  I limited my exploration to the
first chapter of John and the nine earliest manuscripts available on the site
(this was because the website provided no bulk download of the data, so each
chapter of each manuscript had to be copied and pasted and pre-processed by
hand; another evidence of the sparsity of data available in this field).  I
used the alignment software previously mentioned to discover where variant
readings existed in the manuscripts.

As mentioned previously, according to Aland, all one must do in order to
establish the critical text of the New Testament is to decide between each of
the extant readings for every passage in which there are variants.  Upon
aligning the manuscripts and finding variant passages, I discovered 187
passages in the first chapter of John alone that contained variant readings,
consisting of 40\% of the text, with just nine manuscripts as evidence.  That
gave me some idea of the enormity of this task, which I had not previously
appreciated.  The vast majority of these differences were a spelling error or
word order change introduced into one of the nine manuscripts, though some were
more significant.  Having obtained a list of variants, there were several
methods I experimented with to shorten the list.

\subsubsection{Abbreviations, Spelling Errors and Shortest Readings}

A simple place to start in identifying easy-to-correct variant passages is to
look for words that are abbreviated.  In addition to the word for ``Son'' seen
previously, there are many other places where abbreviations were commonly used,
including the words for ``God,'' ``man,'' ``Jesus,'' and ``Israel.''  I made a
list of those words and their abbreviations, and upon finding a variant whose
only readings were abbreviations of the longer form of the same word, I marked
it as ignorable.  However, such efforts only shortened the list of variants by
6, to 181.

A slightly more complicated type of variant to check for, but still relatively
simple, is spelling errors.  I built a dictionary by using a parsed version of
the Perseus project~\cite{perseus-project}.  I downloaded all of the text for
some 120 classical Greek works, including such authors as Sophocles, Xenophon,
and Homer.  To build the dictionary, I simply made a set containing all of the
words those authors used.  I then checked every word in each reading against
the dictionary and marked spelling errors.  If it was clear that there was only
one variant that had correct spelling, I marked it as the correct text.  That
effort took the number of variants to around 150.

Grammatical checking is the next logical step, for frequently there are errors
where the article in front of a word does not match the word's case in one of
the manuscripts, among other simple errors.  However, I did not have access to
a Greek grammatical parser or checker and am not proficient enough in ancient
Greek to write one, so I was not able to try this method.  I am confident that
it would greatly reduce the number of variant passages left to decide.

Moving to methods less solidly grounded in knowledge sources available to a
computer, one can also look for areas where words were added, applying the rule
that the shorter reading is the more likely rule.  However, Aland is quick to
say after stating the rule that it cannot be mechanically applied without
consideration of other evidence; thus it gets very difficult for a computer to
decide which reading is most likely to be correct.  Finding out which reading
is shortest is trivial on a computer; deciding when to accept the shorter
reading and when to accept the longer reading is much harder.  In my
experiments I detected which variants looked like they had word additions,
though I was unable to come to any automatic decision about which reading to
keep.

\subsubsection{Combining Manuscript Evidence}

The biggest issue I encountered in trying to apply the ``shorter reading'' rule
was that of combining the evidence from various manuscripts in some reasonable
way.  As Aland's fourth rule listed above says, manuscripts should be weighed,
not counted.  Thus some measure of reliability is required, along with a way to
combine ``votes'' from various manuscripts and their reliabilities in order to
come to a conclusion about the original reading.  As mentioned in the
discussion of knowledge sources, the concept of reliability is somewhat
circular, because manuscripts are reliable when they contain correct readings,
and correct readings are in turn determined by the reliability of the
manuscripts that attest them.  This lends itself naturally to some iterative
process in an automated algorithm.

However, even given a set of reliability scores, I was unable to determine a
reasonable way to combine the evidence of the various manuscripts.  A simple
majority vote weighted by reliability does not seem reasonable, as a poor
manuscript could have hundreds of copies; even if those poor copies have very
small weight, they could easily outweigh a single, much better manuscript.  A
simple average also does not seem reasonable, for sometimes even the most
reliable manuscripts are wrong; this method would fail miserably when the most
reliable manuscript is the only manuscript attesting a particular incorrect
reading.  Clearly there needs to be some more complicated way of combining the
evidence, though exact methods for such a procedure are elusive to me.
Perhaps a professional textual critic could give some insight that would be
helpful in this area.

\subsubsection{Establishing Reliability of Manuscripts}

Even without a good method for combining manuscript evidence, one can still
compute some initial reliability scores based on the critical decisions that
have been made.  I calculated a simple score based on the spelling errors and
word insertions I observed.  The values of the score were arbitrary; I started
each manuscript with a base score of 100 and subtracted a small number of
points for each spelling error, and a much larger number of points for each
passage that had word insertions.  I weighted the number of points subtracted
by the length of the manuscript, because some manuscripts contained only
fractions of the text, and thus had less opportunity to lose reliability,
leading to scores that were disproportionately high.

The results of the initial reliability scores were somewhat interesting.  The
manuscript commonly referred to as 5 had the highest score, largely because it
only contained one fourth of the first chapter of John (I was not able to
completely erase the higher score given to shorter manuscripts).  However, the
next highest manuscript in reliability was 1 (Codex Sinaiticus), followed by 3
(Codex Vaticanus); those two manuscripts are commonly considered by textual
critics to be the most reliable manuscripts that we have, in that order.  So a
simple reliability measure affected almost entirely by addition of words in
just the first chapter of John reproduced to some extent the ordering
determined by professional textual critics.

The results I saw could very possibly be entirely accidental.  However, what
one can observe is that given a set of critical decisions about passages in the
text, a reliability score can easily be assigned to each manuscript.  There is
some question about how best to calculate the score, but a textual critic could
readily give some basic guidelines as to which kinds of variants affect
reliability the most, or even mark for each individual variant how much a
correct reading in that variant should matter in the reliability score.  That
score could then be used in other parts of an automated system for textual
criticism, including an iterative update to the reliability if a way of
combining evidence is decided upon.

\subsection{An Aid to the Textual Critic}
\label{sec:aid-to-critic}

Through my experiments in trying to produce a critical text, I came to the
conclusion that at least at present the place for computers in the work of
textual criticism is as an aid to, not a replacement of, the textual critic.
There are several ways in which computers can aid in the process.

The first and most obvious benefit is the automatic detection of variants in
the manuscripts.  This is clearly feasible, given transcriptions of the
manuscripts, as I demonstrated in my code.  The automatic detection of variants
allows for completely exhaustive critical apparati, which as of yet do not
exist.

The second benefit that computers can bring to textual criticism is the
filtering of variants, catching and marking easily decidable variants such as
spelling errors.  With much of the noise removed, the textual critic could
focus his attention on the important passages.  I was only able to filter
spelling errors, but I imagine a grammatical check could detect other easily
decidable errors, and a textual critic could probably list a few more
possibilities.

The third benefit that I see is the automated calculation of a reliability
score.  Given some criteria specified by the textual critic, such as how to
weight different types of variants, a computer could return a reliability for
each manuscript which could be used in later decisions.  I can envision a
system where a list of variants for a passage is presented on the screen,
ordered by the reliability of the manuscripts that contain the variant (and
perhaps some other criteria).  The critic would then select a variant as the
correct one, and the computer would recompute reliability scores and show
another variant.  Something related to active learning could even be used in
the process, with the computer selecting which variant to show next.

In short, I am convinced that while the automated generation of a critical text
is currently far beyond our reach, automated methods could greatly reduce the
workload on the textual critic.  What Metzger called ``an enormous amount of
labor'' would all be done by the computer, leaving the text critic himself to
focus on more important things.

\section{Conclusion}
\label{sec:conclusion}

The purpose of this project was to explore the current extent of automated
methods in textual criticism, and what room there is for the increased use of
automated methods.  In doing this project, I have learned a great deal about
textual criticism itself, reading books and portions of books on the subject.
I also have had an enlightening experience attempting to automate myself some
of the processes involved in textual criticism.

I have reviewed the literature extensively, and inquired of BYU professors
involved in textual criticism, and come to the conclusion, along with many
others in the field, that the field of textual criticism could be greatly
enhanced by the more effective use of computing resources.  It has come a long
way in recent years, particularly in the application of methods from
computational biology to reconstructing manuscript trees.  But much more work
can yet be done.

Lastly, I have experimented with specific methods to aid in the work of the
textual critic, and proposed some future work that I think would be very
beneficial to textual criticism.  Computers can remove much of the mundane from
current methodologies, allowing researchers to focus their efforts on critical
tasks.

\section{Future Work}

One obvious next step in this work is to produce a system like that described
in \secref{aid-to-critic}.  The code I produced could be a starting place, but
significant work would need to be done on top of my code, including a mechanism
for accepting input from a user, improvements to the reliability score
calculation, the addition of some kind of grammatical checker to filter the
variant list, and some kind of active-learning decision about which variants to
show the user.  The output from a stemmatological analyzer could also likely be
incorporated into the code, probably in the calculation of the reliability
score.  Such work would best be done in conjunction with a professional textual
critic, though who exactly would be most interested in such a project is
unknown to me.  As mentioned previously, Michael Bushell of BibleWorks would
likely be a good person to contact.

The helpful input that could be had from a textual critic has been mentioned
throughout this paper; in brief summary, it includes specific advice about how
to determine how reliable a manuscript is, how to combine evidence from many
manuscripts, and what other kinds of variants are easily decidable.  Certainly
it would also be good to know if a tool such as the one described would be
found useful.  In order to answer these questions, there are several professors
at BYU who may be good to talk to.  I found it somewhat difficult to interview
the faculty, as many of them are away for the summer.  I did receive responses
to my emails from two faculty members (Thomas Wayment and Stephen Bay).  While
both were helpful in answering my questions, neither seemed particularly
excited about my project.  Perhaps seeing this report containing some concrete
ideas for work in automating textual criticism may prove more effective in
gaining interest.

Future work could also more fully explore the ideas used in the
``Coherence-Based Genealogical Method'' for stemmatology.  While the work
focuses on producing a stemma of the text, they also are interested in
producing a critical text in a semi-automated fashion, and some of their ideas
are applicable to this work.  I came across the work too late in my project to
allow for very much exploration.

Another idea that could be explored in automating the process is that of
creating some kind of statistical model for the copying process, then running
inference in order to determine the original text.  Such a model would be very
complicated, and I do not have many ideas on how to proceed in that area.  I
suspect that the model would have severe deficiencies because of the difficulty
of modeling the ``easiness'' of a reading.

\bibliographystyle{plain}
\bibliography{../../../bib/textual-criticism/bib}

\end{document}