### 2010-10-08 masters_thesis/entropy_nips_workshop/paper.tex
Moved the original nips workshop to entropy_nips_workshop, because we are thinking of doing two papers

\documentclass{article}
\usepackage{nips10submit_e,times}
\usepackage{graphicx}

\title{Improving the Visualization of Topic Models}
\author{
Matthew J. Gardner \\
Department of Computer Science \\
Brigham Young University \\
Provo, UT 84604 \\
\texttt{mjg82@byu.edu} \\
\And
Other authers\ldots \\
Affiliation \\
Address \\
\texttt{email} \\
}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

\begin{abstract}
\end{abstract}

\section{Introdcution}

Topic modeling has shown to be an effective means of extracting semantic
meaning from a collection of
documents~\cite{blei-2003-latent-dirichlet-allocation}.  Topic models have been
used as an inputs to other kinds of algorithms, such as classification and
sentiment tasks (cite some papers here), but the usefulness of topic models as
visualization tools in and of themselves has not been well established, and is
even in question (cite more papers).  Some recent work has been done in an
attempt to make corpus browsers using the results of LDA, and to make the
results of LDA more easily interpretable, but there is much work yet to be done
(still more citations).

Common approaches to using topic models to facilitate corpus browsing and
visualization are limited mainly to showing the similarity between documents in
the corpus in various ways.  Such methods are useful in their own right when
one is individually scanning documents, but they are insufficient to answer
broader questions about the corpus, like, ``What topics show up most in these
documents?'' and, ``What does a particular author in the document talk about
most frequently that is unique to that author?''  Those questions turn one away
from analyzing individual documents to trying to get information out of the
topics themselves.

Upon looking at the topics one is confronted with tens or hundreds of lists of
words, sorted by a number given to each list by the algorithm.  Sifting through
those lists and trying to understand what they mean, and whether or not a
particular topic is important, can be a daunting task, particularly to one not
very familiar with the algorithm.  The common practice of cherry-picking good
topics to present in a paper hides the fact that a large number of topics
returned by topic models are very hard to interpret or even completely
meaningless to humans.

In order for topic models to be useful for corpus browsing, some sense has to
be made out of the topics.  One recently proposed method is to assign to each
topic a coherence value, so that the person can ignore topics that are likely
to be meaningless~\cite{newman-2010-automatic-evaluation-of-topic-coherence}.
While topic coherence is able to eliminate meaningless topics from
consideration, it is not able to address many broader questions that a person
browsing the corpus might have.

In this paper we present the idea of using the distribution of documents for a
given topic, $p(d|z)$, in order to answer questions about a corpus that
previous visualization techniques are not able to answer.  The distribution of
topics in a document, $p(z|d)$, has been widely used in all applications of
topic modeling, from dimensionality reduction of document representations to
showing similar documents in a corpus browser.  This is understandable, given
that estimating the distribution $p(z|d)$ is one of the primary purposes of
topic modeling.  However, $p(d|z)$ also contains interesting information that
can be very useful for visualizing and understanding the output of the topic
model.

The rest of this paper is outlined as follows.  We first describe a corpus of
campaign speeches from the 2008 presidential primaries and general election.
Then we present questions that a person browsing through these campaign
speeches may want to answer, and how one can use $p(d|z)$ to answer those
questions where other methods fail.  Finally, we conclude.

\section{Campaign Speeches}

TODO

\section{What topics were most prevelant during the 2008 campaign cycle?}

TODO

\section{What topics were most universal across candidates?}

TODO

\section{What did Barack Obama talk about most frequently?}

TODO

\section{Conclusion}

TODO

\bibliographystyle{unsrt}
\bibliography{../../../bib/lda/bib}

\end{document}