### 2010-11-13 applications/nsf_fellowship/personal_statement.tex
Finished a draft of my research statement, started on my personal statement.

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{natbib}

\setlength{\textheight}{9in}

\begin{document}

\pagestyle{empty}

\begin{center}
  \textbf{Personal Statement}\\
  Matthew Gardner
\end{center}
\vspace{-4mm}

My father, grandfather, and great-grandfather all earned PhDs in physics.  I
grew up wanting to be a physicist, to win the Nobel Prize by discovering a
Unified Field Theory.  So of course, when I started college, I chose physics as
my major.  During my freshman year I was actively involved in the Society of
Physics Students, encouraging other students to study physics, helping to tutor
lower-level students, and even going to give physics and chemistry
demonstrations at a local middle school, hoping to excite the children to learn
about science.

I also took a couple of computer science classes my freshman year, as the field
was interesting to me, though physics and mathematics still dominated my
thoughts.  It wasn't until I took a break from school for two years and had
some time away from classes that I really started to be fascinated by computer
science.

After my freshman year of college, I spent two years as a volunteer missionary
in Alabama, paying my way to serve the hispanic community in that state.  I
learned Spanish in order to communicate with them, and I came to understand
much better the circumstances that immigrants to our country live in.  I did a
significant amount of Spanish-English interpretation, and it was in that
capacity that I first started thinking about the potential for computer science
to improve the lives of ordinary people.  What if a computer could interpret
live speech from one language to another, accurately and effectively?  Such a
system would ease the burdens on so many immigrants who have much difficulty
communicating with other people for their daily needs, and not just those in
our country.

I came back from my experience in Alabama with a new-found desire to study
computer science.  I changed my major to computer science, with minors in
physics and mathematics, as those subjects were still interesting to me and
they seemed like relevant background.  I found a computer science professor who
was doing research that I found intriguing; I wanted to learn about machine
translation, which I knew was part of machine learning, so I started working in
the Applied Machine Learning Lab with Dr. Kevin Seppi.

My time working in the Applied Machine Learning lab has been the defining
experience of my undergraduate education.  I had very close interaction with
several professors and many graduate students who taught me what research was
like.  I also learned a great deal more than was taught in my classes in order
to understand the research that was going on in the lab.  Several times in my
classes the professor had me give part of a lecture that dealt with things I
was actively researching.

While I have tutored other students, mostly informally, throughout all of my
college career, my first formal teaching experience came my first semester as a
masters student.  I was a teaching assistant for two classes that Dr. Seppi
taught: Introduction to Articial Intelligence, and Bayesian Inference.  I had
always enjoyed helping people understand concepts, but I didn't think I would
enjoy the teaching aspect of being a professor until that experience.  I found
that figuring out how to explain complicated concepts simply and seeing the
light of understanding in the students' faces gave me great satisfaction.

I envision my future career as one of research, likely in the domain of natural
language processing, though possibly also in computer vision or other areas of
machine learning.  I have not yet decided between research in industry or
academia; perhaps my career will take me first to industry, ending up as a
professor later in life.

I have spent some time as a teaching assistant, and I
feel that I would enjoy teaching students of all the interesting problems there
are in computer science, getting them excited to do research in the field.  But
what I most want from my career is to find new problems to solve that will
better society, then solve them.  Things like machine translation and
interpretation, or extracting knowledge from large collections of text and
making that knowledge accessible to people, are intriguing problems that I
would love to spend my life's work on.

\end{document}