### 2010-11-13 applications/nsf_fellowship/previous_research.tex

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{natbib}

\setlength{\textheight}{9in}

\begin{document}

\pagestyle{empty}

\begin{center}
  \textbf{Previous Research Experience}\\
  Matthew Gardner
\end{center}
\vspace{-4mm}

Brigham Young University is somewhat unique among universities in that it gives
undergraduates superb opportunities to participate in research, even working
very closely with professors.  During my second year of college, I began
working in the Applied Machine Learning lab at BYU, headed by Dr. Kevin Seppi,
and I was one of only three students that he advised at the time.  For the next
two and a half years, I worked on several aspects of essentially the same
problem---how to effectively solve long-running optimization problems with a
large cluster of hundreds or thousands of machines.

This project started when a physics professor approached my advisor seeking
help solving a crystallography problem.  Having just begun working with my
advisor, I accepted the problem mostly in an auxiliary role, writing some code
and running experiments for Dr. Campbell, the physics professor.  We used
particle swarm optimization, a bio-inspired optimization algorithm similar to
evolutionary algorithms, to try to fit to data his model of what was happening
in a particular kind of crystal.  After spending a lot of time seeing the
output of the algorithm, I also helped to interpret the results.  I discovered
that we had too many free parameters in our model and helped Dr. Campbell to
make progress in solving his problem.  Because of our help in solving his
problem, Dr. Campbell has made what he considers a breakthrough in
understanding these crystals and plans to submit his work for publication soon.

The crystallography problem we were solving required the evaluation of a
function that took several minutes to run.  As each particle in the algorithm
must evaluate the function at every iteration, we used several hundred
processors on BYU's supercomputer in our optimization of the function.  This
led us to ask some interesting questions about how we could best use those
processors to solve the problem.  We first experimented with the topology of
the algorithm, or how the particles in the swarm communicate with each other.
I did much of the experimentation, though I was directed by a graduate student
in the lab, who eventually wrote the paper that we
published~\cite{mcnabb-2009-large-particle-swarms}.  We saw some improvements
in the performance of the algorithm with our new topologies, though we also
noticed that increasing the number of particles with additional processors had
diminishing returns.

To solve this problem, my advisor proposed applying the idea of speculative
decomposition in the parallelization literature to particle swarm optimization
(PSO).  Instead of increasing the number of particles in the swarm, additional
processors could be used to evaluate possible next-iteration positions of each
particle in parallel.  I thought it was an interesting idea, and I made it my
honors thesis.  I took the idea and discovered how it could actually be done,
performing the evaluations needed and piecing the results back together from
messages sent between processors so that speculative evaulation could be done
in a distributed PSO environment with only a single round of message passing
per iteration.  After implementing the technique, I ran many experiments to
compare our methods to other means of parallelizing the algorithm.  Once I was
satisfied that my results were significant, I wrote, submitted, and
successfully defended my honors thesis.  I also modified the paper to fit in
the aims and page requirements of Parallel Problem Solving from Nature, a
well-respected parallel optimization conference.  I submitted my thesis
there and it was accepted~\cite{gardner-2010-speculative-evaluation-in-pso}.

In implementing speculative evaluation, however, I noticed a number of ways the
method could potentially be improved.  Even though my honors thesis was
completed, I continued my investigation.  I found that a large number of the
speculative evaluations we performed were wasted, and that the algorithm could
be improved by replacing useless evaluations with better evaluations, such as
speculating several iterations ahead on paths that particles were likely to
take.  I significantly expanded the work of my honors thesis, preparing to
submit it to a journal on optimization.  After a rejection of a smaller version
from a conference, I got quite discouraged, wanting to give up and move on to
something new.  My advisor insisted I persevere, saying it was good preparation
for a PhD, and so I continued.  After running more experiments to strengthen
the paper, we submitted it to the journal Swarm Intelligence, where it is
currently under
review~\cite{gardner-2010-speculative-approach-to-parallelization-pso}.

While I have since move on to other interests, this project was my first
experience performing research and dealing with the submission process at
conferences.  All told, I worked on particle swarm optimization for about two
and a half years.  I learned a lot about what research is, the perseverance
required in getting a PhD, and how much I love exploring new topics and
extending the limits of what we can do, as well as helping researchers in other
areas to solve their problems.  This project set the stage for my masters
thesis and for my future work as a PhD student.

Along with the research I performed as an undergraduate research assistant,
there are two other research projects worth noting here.  One of them was my
internship at Google in the summer of 2009.  I was presented with the problem
of how to determine the effectiveness of TV advertisements given only a list of
at what times and on what channels the ads played, along with a list of phone
calls received in response to advertisements.  This was complicated because
often ads play at the same time on different channels, and they use the same
phone number.  I developed and implemented a probabilistic model that assigned
phone calls to ads that were played.  The model learned a response rate for
each channel the ads played on and helped advertisers know where to most
effectively spend their money.

The other research area that I have pursued is topic modeling, a method for
capturing the semantics of a document collection, and that is the area of my
masters thesis.  As an undergraduate, I implemented a topic model and used it
to analyze a novel dataset of religious discourses, which I presented at an
international conference held at
BYU~\cite{gardner-2010-general-conference-topics}.  As a graduate student, I
have been investigating how to use topic models to get specific, corpus-wide
information out of large collections of documents.  My initial endeavors have
involved building a tool to browse the output of a topic model in novel
ways~\cite{gardner-2010-topic-browser}.  As other departments at BYU have seen
the work that I and others have done with our topic browser, they have been
very interested in using our tool.  Three separate groups at BYU have asked to
use our browser to analyze data with purposes ranging from finding public
health information in Twitter feeds to finding problem areas in teaching from
free-form student ratings of classes and professors.  My future plans with
topic modeling are presented in the research proposal submitted with this
application.

\footnotesize
\bibliographystyle{plain}
\renewcommand\bibsection{\subsubsection*{Publications}}
\bibliography{../../../bib/aml/bib}

\end{document}