### 2010-12-14 resume_cv/cv.tex
Updated my cv a little bit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% Simple LaTeX CV Template %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Document Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't like 10pt? Try 11pt or 12pt
\documentclass[10pt]{article}

% This is a helpful package that puts math inside length specifications
\usepackage{calc}

% Simpler bibsection for CV sections
% (thanks to natbib for inspiration)
\makeatletter
\newlength{\bibhang}
\setlength{\bibhang}{1em}
\newlength{\bibsep}
 {\@listi \global\bibsep\itemsep \global\advance\bibsep by\parsep}
\newenvironment{bibsection}
    {\minipage[t]{\linewidth}\list{}{%
        \setlength{\leftmargin}{\bibhang}%
        \setlength{\itemindent}{-\leftmargin}%
        \setlength{\itemsep}{\bibsep}%
        \setlength{\parsep}{\z@}%
        }}
    {\endlist\endminipage}
\makeatother

% Layout: Puts the section titles on left side of page
\reversemarginpar

%
%         PAPER SIZE, PAGE NUMBER, AND DOCUMENT LAYOUT NOTES:
%
% The next \usepackage line changes the layout for CV style section
% headings as marginal notes. It also sets up the paper size as either
% letter or A4. By default, letter was used. If A4 paper is desired,
% comment out the letterpaper lines and uncomment the a4paper lines.
%
% As you can see, the margin widths and section title widths can be
% easily adjusted.
%
% ALSO: Notice that the includefoot option can be commented OUT in order
% to put the PAGE NUMBER *IN* the bottom margin. This will make the
% effective text area larger.
%
% IF YOU WISH TO REMOVE THE ``of LASTPAGE'' next to each page number,
% see the note about the +LP and -LP lines below. Comment out the +LP
% and uncomment the -LP.
%
% IF YOU WISH TO REMOVE PAGE NUMBERS, be sure that the includefoot line
% is uncommented and ALSO uncomment the \pagestyle{empty} a few lines
% below.
%

%% Use these lines for letter-sized paper
\usepackage[paper=letterpaper,
            %includefoot, % Uncomment to put page number above margin
            marginparwidth=1.2in,     % Length of section titles
            marginparsep=.05in,       % Space between titles and text
            margin=1in,               % 1 inch margins
            includemp]{geometry}

%% Use these lines for A4-sized paper
%\usepackage[paper=a4paper,
%            %includefoot, % Uncomment to put page number above margin
%            marginparwidth=30.5mm,    % Length of section titles
%            marginparsep=1.5mm,       % Space between titles and text
%            margin=25mm,              % 25mm margins
%            includemp]{geometry}

%% More layout: Get rid of indenting throughout entire document
\setlength{\parindent}{0in}

%% This gives us fun enumeration environments. compactitem will be nice.
\usepackage{paralist}

%% Reference the last page in the page number
%
% NOTE: comment the +LP line and uncomment the -LP line to have page
%       numbers without the ``of ##'' last page reference)
%
% NOTE: uncomment the \pagestyle{empty} line to get rid of all page
%       numbers (make sure includefoot is commented out above)
%
\usepackage{fancyhdr,lastpage}
\pagestyle{fancy}
%\pagestyle{empty}      % Uncomment this to get rid of page numbers
\fancyhf{}\renewcommand{\headrulewidth}{0pt}
\fancyfootoffset{\marginparsep+\marginparwidth}
\newlength{\footpageshift}
\setlength{\footpageshift}
          {0.5\textwidth+0.5\marginparsep+0.5\marginparwidth-2in}
\lfoot{\hspace{\footpageshift}%
       \parbox{4in}{\, \hfill %
                    \arabic{page} of \protect\pageref*{LastPage} % +LP
%                    \arabic{page}                               % -LP
                    \hfill \,}}

% Finally, give us PDF bookmarks
\usepackage{color,hyperref}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}

%%%%%%%%%%%%%%%%%%%%%%%% End Document Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%% Helper Commands %%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The title (name) with a horizontal rule under it
%
% Usage: \makeheading{name}
%
% Place at top of document. It should be the first thing.
\newcommand{\makeheading}[1]%
        {\hspace*{-\marginparsep minus \marginparwidth}%
         \begin{minipage}[t]{\textwidth+\marginparwidth+\marginparsep}%
                {\large \bfseries #1}\\[-0.15\baselineskip]%
                 \rule{\columnwidth}{1pt}%
         \end{minipage}}

% The section headings
%
% Usage: \section{section name}
%
% Follow this section IMMEDIATELY with the first line of the section
% text. Do not put whitespace in between. That is, do this:
%
%       \section{My Information}
%       Here is my information.
%
% and NOT this:
%
%       \section{My Information}
%
%       Here is my information.
%
% Otherwise the top of the section header will not line up with the top
% of the section. Of course, using a single comment character (%) on
% empty lines allows for the function of the first example with the
% readability of the second example.
\renewcommand{\section}[2]%
        {\pagebreak[2]\vspace{1.3\baselineskip}%
         \phantomsection\addcontentsline{toc}{section}{#1}%
         \hspace{0in}%
         \marginpar{
         \raggedright \scshape #1}#2}

% An itemize-style list with lots of space between items
\newenvironment{outerlist}[1][\enskip\textbullet]%
        {\begin{itemize}[#1]}{\end{itemize}%
         \vspace{-.6\baselineskip}}

% An environment IDENTICAL to outerlist that has better pre-list spacing
% when used as the first thing in a \section
\newenvironment{lonelist}[1][\enskip\textbullet]%
        {\vspace{-\baselineskip}\begin{list}{#1}{%
        \setlength{\partopsep}{0pt}%
        \setlength{\topsep}{0pt}}}
        {\end{list}\vspace{-.6\baselineskip}}

% An itemize-style list with little space between items
\newenvironment{innerlist}[1][\enskip\textbullet]%
		{\begin{compactitem}[#1]}{\end{compactitem}}

% To add some paragraph space between lines.
% This also tells LaTeX to preferably break a page on one of these gaps
% if there is a needed pagebreak nearby.
\newcommand{\blankline}{\quad\pagebreak[2]}

%

%%%%%%%%%%%%%%%%%%%%%%%% End Helper Commands %%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% Begin CV Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\makeheading{Matthew J. Gardner}

\section{Contact Information}
%
% NOTE: Mind where the & separators and \\ breaks are in the following
%       table.
%
% ALSO: \rcollength is the width of the right column of the table
%       (adjust it to your liking; default is 1.85in).
%
\newlength{\rcollength}\setlength{\rcollength}{1.85in}%
%
\begin{tabular}[t]{@{}p{\textwidth-\rcollength}p{\rcollength}}
795 Walnut Avenue & \textit{Cell:} (801) 995-9460 \\
Provo, UT 84604 USA    & \textit{E-mail:} mjg82@byu.edu \\
\end{tabular}

\section{Research Interests}
%
Natural language processing, latent Dirichlet allocation, topic modeling,
Bayesian machine learning, particle swarm optimization

\section{Education}
%
{\textbf{Brigham Young University}},
Provo, Utah USA
\begin{outerlist}
\item[] M.S.,
		Computer Science, April 2011 (anticipated)
        \begin{innerlist}
		\item Thesis Topic: \emph{Information discovery in large corpora
		  through topic modeling}
        \item Advisor: Professor Kevin Seppi
        \end{innerlist}

\item[] B.S.,
		Computer Science, April 2010
        \begin{innerlist}
        \item Graduated summa cum laude (GPA 4.0) and with University honors
        \item Minors in Physics and Mathematics
        \item Honors Thesis Topic: \emph{Speculative Evaluation in Particle
		  Swarm Optimization}
        \item Advisor: Professor Kevin Seppi
        \end{innerlist}

\end{outerlist}

\section{Awards}
%
Brigham Young University
\begin{innerlist}
\item National Merit Scholarship, 2004--2010
\item Awarded a research grant from the Office of Research and Creative
  Activities in both 2008 and 2009
\item Inducted into the Phi Kappa Phi Honors Society, 2009
\item Barry M. Goldwater Scholarship Nominee, 2009
\item Claricode Medical Essay Scholarship Finalist, 2009
\end{innerlist}

\section{Academic Experience}
{\textbf{Johns Hopkins University}}, Baltimore, Maryland USA
\begin{outerlist}

\item[] \textit{Human Language Technologies Summer School Attendee}%
        \hfill \textbf{June 2010}
\begin{innerlist}
\item One of seven recipients of a NAACL scholarship to attend the summer school
  at JHU.
\item Studied with top researchers in HLT at JHU and the University of Maryland.
\end{innerlist}
\end{outerlist}

\blankline

{\textbf{Brigham Young University}}, Provo, Utah USA
\begin{outerlist}
\item[] \textit{Teaching Assistant}%
	\hfill \textbf{May 2010 to June 2010}
\begin{innerlist}
\item CS 470: Introduction to Artificial Intelligence.  Helped create new
  curriculum for the class, tutored students, and graded homework and reports.
\item CS 677: Bayesian Inference.  Tutored students and graded homework and
  reports.
\end{innerlist}

\item[] \textit{Undergraduate Researcher}%
        \hfill \textbf{December 2007 to April 2010}
\begin{innerlist}
\item Worked in the Applied Machine Learning Lab in the Computer Science
  Department under Dr. Kevin Seppi.
\item Research topics included particle swarm optimization and Markov decision
  processes.
\item Reviewed papers for the International Conference on Machine Learning 2009.
\end{innerlist}
\end{outerlist}

\section{Professional Experience}
%
\textbf{Google, Inc.},
Irvine, California USA
\begin{outerlist}

\item[] \textit{Software Engineering Intern for the TV Ads Team}%
        \hfill \textbf{May 2009 to August 2009}
\begin{innerlist}

\item Used machine learning techniques to attribute calls received by
  advertisers to specific ads played on TV networks.  Official blog post about
  the project can be found at:
  http://google-tvads.blogspot.com/2010/01/track-cost-per-call-data-through-google.html.
\item Gave several presentations to the machine learning lunch group.
\end{innerlist}

\end{outerlist}

\section{Journal Publications}
  \begin{bibsection}
	\item Matthew Gardner, Andrew McNabb, and Kevin Seppi.  A Speculative
	  Approach to Parallelization in Particle Swarm Optimization.
	  Submitted to: \emph{Swarm Intelligence}.
  \end{bibsection}

\section{Conference Publications}
  \begin{bibsection}
	\item Matthew Gardner, Andrew McNabb, and Kevin Seppi.  Speculative
	  Evaluation in Particle Swarm Optimization.  In: \emph{Parallel Problem
	  Solving in Nature, 2010}.

    \item Andrew McNabb, Matthew Gardner, and Kevin Seppi. An Exploration of
	  Topologies and Communication in Large Particle Swarms.  In:
	  \emph{Congress on Evolutionary Computation, 2009}.
  \end{bibsection}

\section{Workshop and other refereed publications}
  \begin{bibsection}
	\item Matthew Gardner, J. Lutes, J. Lund, J. Hansen, D. Walker, E. Ringger,
	  K.  Seppi. The Topic Browser: An Interactive Tool for Browsing Topic
	  Models.  In the Challenges of Data Visualization workshop at \emph{Neural
	  Information Processing Systems, 2010}.

	\item Matthew Gardner and Eric Ringger.  Automatic Topic Discovery in 100
	  Years of General Conference Talks.  Presented at the \emph{50th
	  Anniversary BYU Studies Symposium, 2010}.

  \end{bibsection}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%% End CV Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%