### 2011-01-25 masters_thesis/proposal/proposal.tex
Finished an initial draft of my proposal

\documentclass[ms]{byuprop}
% Options for this class include the following (* indicates default):
%
%   10pt -- 10 point font size
%   11pt -- 11 point font size
%   12pt (*) -- 12 point font size
%
%   ms -- produce a thesis proposal (off)
%   areaexam -- produce a research area overview (off)
%   phd -- produce a dissertation proposal (off)
%   
%   layout -- show layout lines on the pages, helps with overfull boxes (off)
%   grid -- show a half-inch grid on every page, helps with printing (off)

% This command fixes my particular printer, which starts 0.03 inches too low,
% shifting the whole page down by that amount.  This shifts the document
% content up so that it comes out right when printed.
%
% Discovering this sort of behavior is best done by specifying the ``grid''
% option in the class parameters above.  It prints a 1/2 inch grid on every
% page.  You can then use a ruler to determine exactly what the printer is
% doing.
%
% Uncomment to shift content up (accounting for printer problems)
%\setlength{\voffset}{-.03in}

% Here we set things up for invisible hyperlinks in the document.  This makes
% the electronic version clickable without changing the way that the document
% prints.  It's useful, but optional.
\usepackage[
    ps2pdf,
    bookmarks=true,
    breaklinks=true,
    raiselinks=true,
    pdfborder={0 0 0},
    colorlinks=false,
    ]{hyperref}

% Rewrite the itemize, description, and enumerate environments to have more
% reasonable spacing:
\newcommand{\ItemSep}{\itemsep 0pt}
\let\oldenum=\enumerate
\renewcommand{\enumerate}{\oldenum \ItemSep}
\let\olditem=\itemize
\renewcommand{\itemize}{\olditem \ItemSep}
\let\olddesc=\description
\renewcommand{\description}{\olddesc \ItemSep}

% Get a little less fussy about word spacing on a line.  Sometimes produces
% ugly results, so keep your eyes peeled.
\sloppy

% Important settings for the byuprop class. %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Because I use these things in more than one place, I created new commands for
% them.  I did not use \providecommand because I absolutely want LaTeX to error
% out if these already exist.
\newcommand{\Title}{Improved Corpus Exploration with Topic Modeling}
\newcommand{\Author}{Matthew Gardner}
\newcommand{\SubmissionMonth}{January}
\newcommand{\SubmissionYear}{2011}

% Take these from the commands defined above
\title{\Title}
\author{\Author}
\monthsubmitted{\SubmissionMonth}
\yearsubmitted{\SubmissionYear}

% Committee members
\committeechair{Kevin Seppi}
\committeemembera{Eric Ringger}
\committeememberb{David Embley}
%\committeememberc{}
%\committeememberd{}

% Department graduate coordinator
\graduatecoordinator{Kent Seamons}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Set up the internal PDF information so that it becomes part of the document
% metadata.  The pdfinfo command will display this. Be sure to set the document
% type and add your own keywords.
\hypersetup{%
    pdftitle=\Title,%
    pdfauthor=\Author,%
    pdfsubject={Masters thesis proposal, BYU CS Department: %
                Submitted \SubmissionMonth~\SubmissionYear, Created \today},%
    pdfkeywords={topic modeling, corpus browsing, corpus discovery, %
                               text mining},%
}

% These packages allow the bibliography to be sorted alphabetically and allow references to more than one paper to be sorted and compressed (i.e. instead of [5,2,4,6] you get [2,4-6])
\usepackage{url}
\usepackage[numbers,sort&compress]{natbib}
\bibliographystyle{annotnat}
\usepackage{hypernat}
% Note, use \citet to name the authors (text mode) and \citep to do just give
% the number (parenthetical mode).  Also, \citep* will give all of the authors
% rather than an abbreviated list.

% Additional packages required for your specific thesis go here. I've left some I use as examples.
\usepackage{graphicx}
\usepackage{psfrag}
%\usepackage{pdfsync}
\usepackage{amsmath}

% Better references, I think
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}

\begin{document}

% Produce the preamble
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Abstract}
% 1 to 2 paragraphs

Topic modeling, including latent Dirichlet allocation and its variants, has
been shown to be an effective means of modeling a collection of documents for a
wide variety of purposes.  Topic models have been successfully used in
dimensionality reduction for classification tasks, in associating text with
components in an image, and in aligning words in bilingual corpora.  Despite
the wide-spread use of topic models in many areas of machine learning, however,
their usefulness as an aid to corpus browsing, or trend discovery in large
collections of documents, has not been well established.

I propose to investigate the application of topic models to the task of corpus
browsing.  That is, I seek to show that a topic model is an effective aid to a
human looking through a previously unseen collection of documents.  I will show
this in two parts.  The first part involves the building of a tool that allows
one to easily browse through a corpus in the context of a topic model.  The
second part of this work will establish that topic models do in fact capture
themes that a human would have found in the data and that those topics can be
automatically presented without the user having to dig through the topic model
to find them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
% 1 to 4 pages

In our modern world, large document collections are becoming more prevalent and
more important.  Companies want to look at online reviews of their products and
intelligence analysts have thousands of documents to sift through looking for
suspicious activity.  The recent controversies involving the public release of
hundreds of thousands of documents by WikiLeaks about the wars in Afghanistan
and Iraq highlight the enormity of this problem.  How does a person even begin
to look through these massive corpora?

Typically, the interface between a person and a document collection has been a
search engine, using a Google-like tool to find documents that contain specific
information.  While such tools can be very useful for some tasks, they cannot
give broad pictures about the corpus as a whole, and they are limited in the
kinds of queries they can answer.  A person has to have specific information in
mind in order to form a query; current search tools are not very helpful when a
person has no idea what to even look for.

Recent methods have been proposed that attempt to address part of this issue.
Topic models, including latent Dirichlet
allocation~\cite{blei-2003-latent-dirichlet-allocation}, gather information
about topics used in a corpus.  These topics are modeled as collections of
words that are used together across documents.  Topic models could feasibly be
used to give a broad overview of a corpus and even answer queries about topics
in the documents, instead of just queries about the documents.

As far as I am aware, however, topic models have not as yet been used for this
purpose.  They have successfully been used as input into document
classification and many other tasks
(e.g.,~\cite{blei-2008-supervised-topic-models}), but attempts to aid humans in
corpus analysis through topic models are very preliminary, mostly giving the
user better access to the documents in the corpus, not to the topics
themselves~\cite{newman-2010-visualizing-with-topic-maps}.  It is generally
assumed that topic models do a good job at capturing thematic elements in large
bodies of text, but while this has been anecdotally asserted on many
occasions~\cite{griffiths-2004-finding-scientific-topics}, to my knowledge it
has never been rigorously established.  Thus topic models have not been widely
used in the qualitative analysis of large corpora.

\subsection{Topic Modeling}

Topic modeling seeks to find the set of topics which are discussed in a corpus
and to what proportion each document contains each of those topics.  Several
methods of topic modeling have been proposed.  Among those are latent semantic
analysis or indexing (LSA or LSI)~\cite{dumais-1988-latent-semantic-analysis},
a probabilistic version of LSA, called pLSA or pLSI~\cite{hofmann-1999-plsi},
and latent Dirichlet allocation, or
LDA~\cite{blei-2003-latent-dirichlet-allocation}.  In this proposed work I
focus on LDA.

LDA is a generative probabilistic model of a collection of documents.  The
documents are modeled as deriving from a collection of topics, where each
document comes from some combination of those topics.  Each word in a
document is drawn from one of the topics in that document, where the topic is
simply a multinomial distribution over words.  The graphical model of LDA is
shown in \figref{lda}.

\begin{figure}
  \centering
  \psfrag{b}[C][C]{\Large$\beta$}
  \psfrag{a}[C][C]{\Large$\alpha$}
  \psfrag{t}[C][C]{\Large$\theta$}
  \psfrag{f}[C][C]{\Large$\phi$}
  \psfrag{w}[C][C]{\Large$w$}
  \psfrag{z}[C][C]{\Large$z$}
  \psfrag{N}[C][C]{\large$N$}
  \psfrag{M}[C][C]{\large$M$}
  \psfrag{K}[C][C]{\large$K$}
  \includegraphics[width=.5\textwidth]{../../figures/lda_model}
  \caption{The graphical model of LDA.}
  \label{fig:lda}
\end{figure}

In the figure, $\alpha$ is a Dirichlet prior over topic distributions in a
document, $\theta$ is a per-document multinomial topic distribution, $z$ is a
topic label for each word in the document, and $w$ is an observed word in the
document.  $\beta$ is a Dirichlet prior over the word distributions in a topic
and $\phi$ is a per-topic multinomial word distribution.  The distributions can
be written formally as follows:

\begin{align*} 
  w_i|z_i,\phi^{(z_i)} &\sim \mathrm{Categorical}(\phi^{(z_i)}) \\
  \phi &\sim \mathrm{Dirichlet}(\beta) \\
  z_i|\theta^{(d_i)} &\sim \mathrm{Categorical}(\theta^{(d_i)}) \\ 
  \theta &\sim \mathrm{Dirichlet}(\alpha) 
\end{align*}

Given a collection of documents, the parameters $\theta$ and $\phi$ can be
inferred using standard statistical techniques.  The two most common methods
for performing inference are Gibbs
sampling~\cite{griffiths-2004-finding-scientific-topics} and variational
methods~\cite{blei-2003-latent-dirichlet-allocation}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}
% 1 to 2 pages

Topic modeling has become incredibly popular in recent years, saturating the
field with all kinds of applications and variations of the method.  Of
particular interest to this proposal are two small areas that have not received
much attention in the literature: visualizing topic models and validating topic
models.

\subsection{Visualizing Topic Models}

Typically, topic models have been visualized by simply listing the 10 words
with the highest probability of occurring in that
topic~\cite{blei-2003-latent-dirichlet-allocation,
griffiths-2004-finding-scientific-topics}.  When presenting topics in a paper,
researchers often sift through the results by hand in order to show topics that
are most meaningful.  This is understandable as showing all of the resultant
topics in a paper is infeasible.  However, such presentations hide the fact
that it can be time consuming and difficult to go through the output of a topic
model by hand.

While displaying an entire topic model analysis is impossible in a research
paper, several static visualizations of whole topic models exist as
websites~\cite{blei-arxiv-corpus-browser, mimno-nyt-browser}.  These typically
present long lists of topics that are very difficult to use for finding
specific information.

Apart from visualizing all of the topics from a topic model as a whole, work
has been done in how to present a topic in more interesting ways than simple
lists of words.  Turbo Topics is a method for extracting meaningful phrases
used in a particular topic that can be presented along with individual
words~\cite{blei-2009-turbo-topics}.  Several people have presented trends of
the use of a topic over time in a
corpus~\cite{griffiths-2004-finding-scientific-topics, mimno-nyt-browser}, and
some effort has been made in trying to automatically give topics a meaningful
name, instead of characterizing the topic only by the most frequently appearing
words~\cite{mei-2007-automatic-labeling-of-topic-models,
lau-2010-best-topic-word-selection}.

The output from a topic model has also been used to inform a more general
corpus browsing experience.  Newmann et al. make what they call ``topic maps,''
which are layouts of the documents in a corpus according to a topic-based
similarity metric~\cite{newman-2010-visualizing-with-topic-maps}.  Stoica et
al. use topic models to create facets to use in a browsing interface, though
they conclude that other methods are superior to topic models in this
task~\cite{stoica-2007-faceted-metadata-structures}.

\subsection{Validating Topic Models}

The question of how to evaluate topic models has been addressed in a number of
different ways.  These include both intrinsic and extrinsic evaluations.

Intrinsic evaluations of topic models use some metric on held-out data to judge
the quality of the topic model itself.  These are typically perplexity or the
likelihood of the held-out data according to the model.  Early papers on LDA
used these metrics~\cite{blei-2003-latent-dirichlet-allocation,
griffiths-2004-finding-scientific-topics}, and recent papers have given better
methods for evaluating these
metrics~\cite{wallach-2009-evaluation-for-topic-models,
buntine-2009-estimating-likelihoods-for-topic-models}.

A new intrinsic evaluation method presented at NAACL this summer is evaluating
the coherence of the words in a topic, or how well the words fit together as
judged by humans~\cite{newman-2010-automatic-evaluation-of-topic-coherence}.
The authors found that using co-occurrence counts gathered from Wikipedia is a
good surrogate for human judgments in this regard.  However, as the paper
focused on finding an automated metric that matched human responses, the
authors did not actually use their metric to compare topic models in any way.

Apartment from intrinsic evaluation methods, which may have limited
applicability in any real task, are extrinsic evaluation methods.  These
methods use the topic model, or some part of it, as input into a secondary
task, such as document classification.  The original paper introducing LDA used
one of these metrics~\cite{blei-2003-latent-dirichlet-allocation}, and
virtually all papers describing an application of topic models have also had
some kind of extrinsic metric.  Other examples of extrinsic evaluations
including predicting the citation count of a scientific
paper~\cite{gerrish-2010-measuring-scholarly-impact} and classifying and
annotating images~\cite{wang-2009-image-classification-annotation}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Thesis Statement}
% 1 to 2 sentences

Topic models, when properly visualized, can be an effective tool in the
qualitative analysis of large text corpora.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Project Description}
% 2 to 5 pages

This project will consist of two main parts.  The first part will explore good
visualizations of topic models and will result in a tool that incorporates
those visualizations in an interactive browser.  The second part will consider
the usefulness of topic models for thematic analysis, including the ability of
a topic model to capture human-specified topics and the validation of automated
methods for finding topics of interest to a person browsing the corpus.

\subsection{The Topic Browser}

Proper visualizations are essential to extracting information from and
identifying trends in data, especially large sets of high dimensional data.
Large text corpora are particularly difficult to visualize, as they may include
many thousands of documents and millions of words.  Topic
modeling~\cite{blei-2003-latent-dirichlet-allocation} is a method of reducing
the dimensionality of a corpus into a set of meaningful topics.  Topic models
have been used with some success as an aid to corpus
visualization~\cite{blei-2009-topic-models,
newman-2010-visualizing-with-topic-maps}.  However, in many papers the
presentation of topic models has been limited to hand-selected, coherent
topics, when often there are many meaningless topics to sift through.  These
presentations leave the viewer wondering what else the topic model discovered
and provide little help in gaining a deep understanding of the model.

To aid in the visualization of topic models and in pattern discovery in
document collections, this project will build a system we call The Topic
Browser, a web-based tool for interactively exploring both the output of a
topic modeling algorithm and its attendant corpus.  This topic browser will
incorporate many visualizations of topic models previously published as well
as some innovative ideas of our own.  The Topic Browser is an aid both to those
who wish to browse through a corpus and for those who wish to analyze the topic
model itself.  We believe that this interactive approach to visualization in
browsing reveals topics and topical trends in large document collections in a
way that is not possible with static visualizations.

One of the innovative ideas incorporated into the topic browser is the use of
topic metrics that can be used to separate topics of interest from
not-as-interesting topics.  The sidebar navigation tool in the topic browser
allows the user to sort and filter topics according to these metrics.  This
allows a person navigating the corpus to easily find the topics he is looking
for, instead of having to sift through all of the topics by hand.

\subsection{LDA for Thematic Analysis}

The second part of this project is motivated by the first---given a tool to
browse through the output of a topic model, and topic metrics that are supposed
to show topics relevant to certain user questions, can we quantitatively
demonstrate that it actually works?

Topic models have been shown to be useful in many tasks, and it is generally
assumed that they do a good job at capturing thematic elements in large bodies
of text.  While this has been anecdotally asserted on many occasions, to our
knowledge it has never been rigorously established.  Thus topic modeling has
not been widely used in the qualitative analysis of large corpora.

This part of the project will quantitatively show that topic models can
effectively capture qualitative or thematic elements in a collection of
documents as judged by humans, and that automated methods can select topics
from a topic model that match human thematic judgments in a variety of
situations.

In order to establish this, I will use two sources of data.  The first comes
from student ratings of courses and professors at Brigham Young University.
After answering many questions which require selecting one of several options,
students are given the opportunity to write in free text whatever comments they
have about the class.  We will use those free-form comments as one source of
text in which we will find thematic elements.  The second source of data we
will use is a set of restaurant reviews from New York CitySearch.

We will choose four subsets of the data to annotate, two in each data source,
each subset consisting of about 40 documents (where each document is
approximately one paragraph long).  We will have 10 people annotate each subset
of the data with topics that they judge to be relevant, so that we can have
some idea of the consistency in human-judged topics.

We will then develop methods for comparing a human-annotated topic to a topic
produced by a topic model.  One of the main parts of this comparison will be
finding the percent of words in the topic annotated by humans that were labeled
with the topic by the topic model.  We will elicit from the annotators words
and phrases that are representative of each topic annotated, along with a
phrasal description of the topic, both of which will be useful in matching the
human topics with machine-learned topics.

Once we have methods to compare the human annotations with the machine output,
we will see to what extent topic models capture human topics.  This will
provide a new extrinsic evaluation method for topic models, as some models will
better capture human topics than others.  We will then also validate the use of
topic metrics to automatically sort topics.  In particular, we will see if
these methods can find the topics the humans thought we prevalent in the
annotated subset of the data when the topic model is run on the entire dataset.
Because we will have two annotated subsets for each data source, we will also
be able to test the ability of automated methods to determine which topics are
common between the two subsets, and which are unique.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Validation}
% 1/2 to 2 pages

This project is expected to produce an easily usable, extensible, and helpful
tool for browsing the output of a topic model, and to validate the use of topic
modeling for the thematic analysis of large text corpora.

\subsection{The Topic Browser}

The first stage of this project is successful if it produces a tool that is
useful to the topic modeling community (including those who do research in
topic modeling and those who use topic models) to visualize their results.
While I will not seek to quantitatively verify with user studies that the tool
produced is a better way to browse topic models than existing tools, its
usefulness should be intuitively visible to those using the browser.  This can
be at least anecdotally verified by other groups at BYU and other universities
using our tool, and the acceptance of a paper describing the browser at a
peer-reviewed conference workshop.

While the project will not verify the usability of the tool as a whole, some
parts of the tool can be individually verified.  Some of the previous research
included in the topic browser has already been independently validated, such as
the Turbo Topics method of extracting phrases from
topics~\cite{blei-2009-turbo-topics}.  In particular, the second stage of this
project will validate the novel contributions in the sidebar of the browser.

\subsection{LDA for Qualitative Analysis}

The second stage of this project investigates the use of topic models as an aid
in the qualitative analysis of large copora.  This stage is successful if it
identifies in what ways topic models successfully assist a human in analyzing a
large collection of text.  We take successfully assisting a human to mean
saving the human work by automatically reproducing themes that a person would
have spent hours or days producing.  We will determine a topic model's ability
to reproduce human themes as outlined in the project description.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Thesis Schedule}
% 1/4 to 1/2 page

\begin{itemize}

\item Completed: build the Topic Browser.

\item February 4, 2011: write code to compare human topics with LDA topics.

\item February 11, 2011: finish collecting human annotations.

\item March 4, 2011: submit a paper to CoNLL detailing the work on validating
  LDA for thematic analysis.

\item March 11, 2011: rework individual papers into a coherent thesis and
submit draft to committee chair.

\item March 20, 2011: submit thesis to second committee member for review.

\item March 30, 2011: submit thesis to third committee member for review.

\item April 13, 2011: defend thesis.

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand\bibsection{\section{Annotated Bibliography}}
%\newcommand\bibpreamble{Material to be added between heading and entries.}
% 2 to 5 pages

\bibliography{../../../bib/lda/bib}

\end{document}