### 2011-02-10 topic_validation/evaluation_proposal/paper.tex
Moved some files because I'm changing my thesis

\documentclass[onecolumn, 11pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{Proposed validation of topic models and topic ranking}
\author{Matthew Gardner and Mike Thompson}
\date{\today}

\begin{document}
\maketitle

\section{Introduction}

Topic models have been shown to be useful in many tasks, and it is generally
assumed that they do a good job at capturing thematic elements in large bodies
of text.  While this has been anecdotally asserted on many occasions, to our
knowledge it has never been rigorously established.  Thus topic modeling has
not been widely used in the qualitative analysis of large corpora.

It is our intent to quantitatively show that topic models can effectively
capture qualitative or thematic elements in a collection of documents as judged
by humans, and that automated methods can select topics from a topic model that
match human thematic judgments in a variety of situations.

\section{Data}

In order to establish this, we will use two sources of data.  The first comes
from student ratings of courses and professors at Brigham Young University.
After answering many questions which require selecting one of several options,
students are given the opportunity to write in free text whatever comments they
have about the class.  We will use those free-form comments as one source of
text in which we will find thematic elements.

The second source of data we will use is a set of product reviews from
Amazon.com, or some other review site that has publicly available data.  We
have yet to select a particular website from which to retrieve data or which
products to use.  However, we will select products which are common enough that
an average person would be qualified to find themes through a set of comments.

\section{Annotation}

As there are hundreds of thousands of comments available from the student
ratings data that we will use, and similarly large numbers of product reviews,
having annotators read and label every document with thematic elements is an
impossible task.  Instead, will we select a subset of the data that is large
enough to exhibit themes but small enough that an average college student can
annotate the data in approximately half an hour.  We imagine that somewhere
between 50 to 100 comments (where each comment is about a paragraph long) is an
appropriate size.  We will judge the amount of time required by annotating the
data ourselves first.  We expect that between 10 and 20 annotators for each
data set will be sufficient to establish a solid set of test data.  Previous
work on evaluating other aspects of topic models successfully used 9
annotators.

An important aspect of the subset of the data that we annotate will be that it
will consist of a set of documents that all have the same value for some
attribute, such as ratings for the same teacher, or reviews about the same
product.  This will facilitate both the finding of consistent thematic elements
and more extensive kinds of validation, which will be explained later.

The kinds of information we will elicit from the annotators will include the
top overall themes (probably ordered) in the whole collection of documents they
are present, as well as where those themes show up in the documents.  It would
likely be helpful for the annotator to select for each theme a single document
that best exhibits that theme, as well as list words that are commonly used in
the theme.  All of this will help us to judge whether a particular topic from
the topic model actually matches the theme given by the annotators.

\section{Validation Methodology}

There are two parts to validating the use of topic models in the qualitative
analysis of text.  The first is showing that topic models actually capture the
thematic elements in a set of documents, as judged by humans.  The second is
showing that methods for ranking the topics in a topic model accurately rank
topics according to the ranking given by humans.  Each part requires separate
validation.

\subsection{Validating the topic model's topics}

To demonstrate that topic models capture thematic elements, a mechanism is
necessary to compare a topic model's topic to the themes recorded by humans.
Devising this similarity metric will require some research, but as a starting
point we will compare the words given by the annotators to the words in the
topic, as well as the top documents given by the annotators to the documents
with the highest proportion of each topic.  We will also try to use some kind
of mutual information between the words in the topic and the title or
description of each theme as given by the annotators.

\subsection{Validating the topic rankings}

In order to show that automatic rankings can faithfully reproduce human
rankings, we will use standard metrics from information retrieval (IR), such as
mean average precision, to compare the two rankings.  This validation depends
on the first, as we must be able to determine if the topics ranked by the
computer match the themes ranked by humans.  This can be problematic if there
is no topic in the topic model that successfully captures a human theme, or if
there are two topics that each capture the theme equally well.  We may have to
modify the standard IR metrics to be better suited to our task.

\end{document}