### 2011-02-10 topic_validation/pre_proposal/paper.tex

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{General Outline of Proposed Research}
\author{Matthew Gardner}
\date{Thursday, August 19, 2010}

\begin{document}
\maketitle

\section{Introduction}

There are often occasions when people are confronted with tens or hundreds of
thousands of documents to analyze, either to find something specific or to
simply figure out what the documents contain.  Intelligence analysts routinely
have 50,000 documents or more to go through in a
year~\cite{washingtonpostarticle}.  Several recent controversies involving
documents posted on WikiLeaks presented people with tens of thousands of
documents to sift through.  Even with a specific purpose in mind, exploring
such a large collection of documents can be entirely unmanageable.  An
intelligence analyst simply looking for interesting subjects in the documents,
without a particular query in mind, would have an even harder time.

A journalist browsing through the Afghan war documents that were recently
posted on WikiLeaks said he started with a query, then went where it led him,
following his curiosity to still more queries, using search as his primary
corpus-browsing tool.  Such methods are clearly very useful in many cases, but
in the case of discovering the interesting aspects of a set of documents, they
cannot be completely exhaustive.  One cannot query for something he has not
thought of.

Topic modeling, including latent Dirichlet allocation and its variations, is a
tool that is well suited to capturing the semantic content in a set of
documents~\cite{blei-2003-latent-dirichlet-allocation}.  Topic modeling has
been used in many applications, including being coupled with sentiment
classifiers, as a substitute for part of speech labels, and as an aid in
information retrieval tasks.  However, its usefulness as a general tool for
corpus browsing is not well established, and efforts to make it so are still
very preliminary, though there has been good work in the last few years.

This work seeks to make LDA more applicable to the task of discovering
information in a corpus.  In particular, there are three points of interest in
this work: how to effectively give a label to a topic, how to automatically
discover which topics have important or interesting content and which contain
merely fluff or function words, and how LDA can or should be modified to handle
specific user queries.

\section{Topic Labeling}

Several recent papers have addressed the issue of how to automatically assign a
coherent label to a topic~\cite{mei-2007-automatic-labeling-of-topic-models,
lau-2010-best-topic-word-selection}.  These papers present ideas that are
successful in some instances, but largely ignore hard cases.  It seems that by
looking at the context of how the topic is used in the documents, a decent
label should be found.

\section{Labeling ``Interesting'' Topics}

There are many topics in LDA that capture semantics whose relevance is
immediately apparent to a person browsing the results.  There are other topics
that are still quite relevant, but the relevance is not immediately apparent
for poor visualization.  There are still other topics that simply contain
function words, or formatting, or other uninteresting elements in the documents
(that possibly should have been removed in a preprocessing step).

Some work has been done in the automatic detection of the relevancy of a topic,
including two recent papers on judging topic
coherence~\cite{chang-2009-reading-tea-leaves,
newman-2010-automatic-evaluation-of-topic-coherence}.  However, the lack of
``coherence'' as judged from the top ten words of a topic and evaluated over
Wikipedia does not necessarily imply lack of relevance in the corpus itself.
There must be some way of judging the relevance of a particular topic, and such
a result would make topic modeling a much more accessible tool for corpus
browsing.

\section{LDA with Queries}

In some instances, a person does not have a particular query in mind when
searching through a set of documents.  In many other cases, however, he does.
There are a few published or obvious methods for using LDA to answer search
queries, but to my knowledge they all involve a post-hoc analysis of the
resultant topic model, not a modification to the algorithm in order to answer
queries (granted, I have not done much searching yet in this area, so ``to my
knowledge'' does not mean a lot).  The potential exists to modify the LDA
algorithm to specifically handle queries.  Exactly what the implications of
this would be are not yet clear to me (what does running LDA get you if you
really just want to do a query, for instance?).

\bibliographystyle{plain}
\bibliography{../../../bib/lda/bib}

\end{document}