### 2011-09-09 masters_thesis/si_paper/paper.tex
Ok, final revisions for real this time

\documentclass[smallcondensed]{svjour3}

\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{psfrag}
\usepackage{natbib}

% Better references, I think
\renewcommand{\sec}[1]{Section~\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\alg}[1]{Algorithm~\ref{alg:#1}}

% Algorithmic changes
\renewcommand{\algorithmicforall}{\textbf{for each}}

%% PSO Stuff
\DeclareMathOperator*{\argmin}{arg\;min}
\DeclareMathOperator*{\argmax}{arg\;max}
\DeclareMathOperator*{\arginf}{arg\;inf}
\DeclareMathOperator*{\argsup}{arg\;sup}
\providecommand{\pers}{\ensuremath{P}}
\providecommand{\neigh}{\ensuremath{N}}
\providecommand{\leftind}{\ensuremath{L}}
\providecommand{\rightind}{\ensuremath{R}}
\providecommand{\nURand}{\ensuremath{\Vec{U}^\neigh}}
\providecommand{\pURand}{\ensuremath{\Vec{U}^\pers}}
\providecommand{\ppos}{\ensuremath{\Vec{x}}}
\providecommand{\fppos}{\ensuremath{\Vec{fx}}}
\providecommand{\pvel}{\ensuremath{\Vec{v}}}
\providecommand{\nbest}{\ensuremath{\Vec{b}^\neigh}}
\providecommand{\fnbest}{\ensuremath{\Vec{fb}^\neigh}}
\providecommand{\pbest}{\ensuremath{\Vec{b}^\pers}}
\providecommand{\fpbest}{\ensuremath{\Vec{fb}^\pers}}
\providecommand{\constriction}{\ensuremath{\chi}}
\providecommand{\ncoeff}{\ensuremath{\phi^\neigh}}
\providecommand{\pcoeff}{\ensuremath{\phi^\pers}}
\providecommand{\obs}{\ensuremath{\Vec{\xi}}}
\providecommand{\ofunc}{\ensuremath{f}}
\providecommand{\swarm}{\ensuremath{swarm}}

%SpecExPSO Stuff
\providecommand{\indic}{\ensuremath{I}}
\providecommand{\specvel}{\ensuremath{\Vec{V}}}
\providecommand{\specpos}{\ensuremath{\Vec{X}}}
\providecommand{\leftn}{\ensuremath{\Vec{x}^\leftind}}
\providecommand{\rightn}{\ensuremath{\Vec{x}^\rightind}}
\providecommand{\caseset}{\ensuremath{\mathcal{C}}}
\providecommand{\casegen}{\ensuremath{c}}
\providecommand{\casedef}{\ensuremath{(\pbest,\nbest)}}
\providecommand{\casexn}{\ensuremath{(S,-)}}
\providecommand{\casexx}{\ensuremath{(S,S)}}
\providecommand{\casexl}{\ensuremath{(S,\leftind)}}
\providecommand{\casexr}{\ensuremath{(S,\rightind)}}
\providecommand{\casepn}{\ensuremath{(-,-)}}
\providecommand{\casepl}{\ensuremath{(-,\leftind)}}
\providecommand{\casepr}{\ensuremath{(-,\rightind)}}
\providecommand{\casepN}{\ensuremath{(-,N)}}
\providecommand{\casexN}{\ensuremath{(S,N)}}
\providecommand{\noeval}[1]{\ensuremath{#1^{-e}}}
\providecommand{\nonbest}[1]{\ensuremath{#1^{-n}}}
\providecommand{\p}{\ensuremath{p}}
\providecommand{\pset}{\ensuremath{\mathbf{p}}}
\providecommand{\s}{\ensuremath{s}}
\providecommand{\sset}{\ensuremath{\mathbf{s}}}
\providecommand{\nsset}{\ensuremath{\mathbf{ns}}}
\providecommand{\n}{\ensuremath{n}}
\providecommand{\nset}{\ensuremath{\mathbf{n}}}
\providecommand{\nnset}{\ensuremath{\mathbf{nn}}}

%Other math
\providecommand{\prob}{\ensuremath{\mathrm{Pr}}}

\title{A Speculative Approach to Parallelization in Particle Swarm
Optimization}

\date{Received: date / Accepted: date}

\author{Matthew Gardner, Andrew McNabb, and Kevin Seppi}

\institute{M. Gardner \at
Brigham Young University.
3361 TMCB, Provo, UT 84604
Tel.: 801-422-8717
Fax: 801-422-0169
\email{mjg82@byu.edu}
\and
A. McNabb \at
Brigham Young University.
3361 TMCB, Provo, UT 84604
\email{a@cs.byu.edu}
\and
K. Seppi \at
Brigham Young University.
3361 TMCB, Provo, UT 84604
\email{k@cs.byu.edu}
}

\begin{document}
\maketitle

\begin{abstract}

Particle swarm optimization (PSO) has previously been parallelized primarily by
distributing the computation corresponding to particles across multiple
processors.  In these approaches, the only benefit of additional processors is
an increased swarm size.  However, in many cases this is not efficient when
scaled to very large swarm sizes (on very large clusters).  Current methods
cannot answer well the question: ``How can 1000 processors be fully utilized
when 50 or 100 particles is the most efficient swarm size?''  In this paper we
attempt to answer that question with a speculative approach to the
parallelization of PSO that we refer to as SEPSO.

In our approach, we refactor PSO such that the computation needed for iteration
$t+1$ can be done concurrently with the computation needed for iteration $t$.
Thus we can perform two iterations of PSO at once.  Even with some amount of
wasted computation, we show that this approach to parallelization in PSO often
outperforms the standard parallelization of simply adding particles to the
swarm.  SEPSO produces results that are exactly equivalent to PSO; that is,
SEPSO is a new method of parallelization and not a new PSO algorithm or
variant.

However, given this new parallelization model we can relax the requirement of
exactly reproducing PSO in an attempt to produce better results.  We present
several such relaxations, including keeping the best speculative position
evaluated instead of the one corresponding to the standard behavior of PSO, and
speculating several iterations ahead instead of just one.  We show that these
methods dramatically improve the performance of parallel PSO in many cases,
giving speed ups of up to six compared to previous parallelization techniques.

\keywords{Parallel algorithms \and Optimization methods \and Particle swarm
optimization \and Speculative Decomposition}

\end{abstract}

\section{Introduction}
\label{sec:intro}

Particle swarm optimization (PSO) has been found to be a highly robust and
effective algorithm for solving many types of optimization
problems~\citep{poli-2008-pso-applications,
poli-2007-particle-swarm-optimization}.  For much of the algorithm's history,
PSO was run serially on a single machine.  However, the world's computing power
is increasingly coming from large clusters of processors.  In order to
efficiently utilize these resources for computationally intensive problems, PSO
needs to run in parallel.

Within the last few years, researchers have begun to recognize the need to
develop parallel implementations of PSO, publishing many papers on the subject.
The methods they have used include various synchronous algorithms
\citep{parsopoulos-2004-parallel-vector-evaluated-pso} and asynchronous
algorithms \citep{mostaghim-2006-multi-objective-pso-on-grids}.  Parallelizing
the evaluation of the objective function can also be done in some cases using
standard techniques \citep{grama-2003-intro-to-parallel-computing}, though that
is not an adaption of the PSO algorithm itself and thus is not the focus of
this paper.

These previous parallel techniques distribute the computation needed by the
particles in the swarm over the available processors.  If more processors are
available, these techniques increase the number of particles in the swarm,
either by adding individual particles or by adding entire new sub-swarms.  In
almost all cases, adding additional particles produces better results in the
same amount of time~\citep{mcnabb-2009-large-particle-swarms}.  In
\fig{iters-sphere} we see an example of this on the well-known benchmark
function Sphere (20 dimensions, reporting the average of twenty runs).  In
terms of the number of iterations performed (which is equivalent to wall-clock
time if all particles are evaluated in parallel), every time the swarm size
increases, the performance improves.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{iters_sphere}
  \caption{Function Sphere with various swarm sizes, comparing performance with
  the number of iterations of the algorithm performed.}
  \label{fig:iters-sphere}
\end{figure}

However, it can be seen from the graph that once the swarm is sufficiently
large, there comes a point of diminishing returns with respect to adding
particles.  The increase in performance seen when moving from 50 to 100
particles is roughly equivalent to the increase seen when moving from 1000 to
4000.  In \fig{evals-sphere} we show the value obtained after 50,000 function
evaluations (not iterations) as a function of swarm size, again for the
function Sphere.  Increasing the swarm size from 5 to 10 has a significant
effect on the value obtained.  However, increasing the swarm size from 16 to 30
makes the algorithm less efficient; that is, it reduces the progress the
algorithm makes per evaluation.  Other functions show similar trends, though
often the optimal swarm size is slightly larger.  For this reason, previous
work has recommended the use of a swarm size of 50 for
PSO~\citep{bratton-2007-defining-a-standard-for-pso}.  Thus, in at least some
cases, adding particles indefinitely will not yield an efficient
implementation.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{evals_sphere}
  \caption{Function Sphere with various swarm sizes, showing performance after
  a set number of function evaluations.  Error bars show mean and 10th and 90th
  percentiles.}
  \label{fig:evals-sphere}
\end{figure}

Our purpose is to explore the question of what to do with a thousand processors
when 50 or 100 particles is the most efficient swarm size, and simply adding
particles results in only incremental improvement.  We thus consider PSO
parallelization strategies for clusters of hundreds or thousands of processors
and functions for which a single evaluation will take long enough to merit a
parallelization of one particle per processor---at least hundreds of
milliseconds, but perhaps several minutes or longer.

In order to solve the problem of diminishing returns, we apply the concept of
speculative decomposition~\citep{grama-2003-intro-to-parallel-computing} to
particle swarm optimization, using extra processors to perform two iterations
of PSO at the same time.  Speculative decomposition is analogous to speculative
execution (also known as branch prediction), a technique commonly used in
processors.  Modern processors, when faced with a branch on which they must
wait (e.g., a memory cache miss), guess which way the branch will go and start
executing, ensuring that any changes can be undone.  If the processor guesses
right, execution is much farther ahead than if it had idly waited on the memory
reference.  If it guesses wrong, execution restarts where it would have been
anyway.  Thus the processor speculates about future paths of execution in an
attempt to decrease overall processing time.

In this paper we show that the results of standard PSO can be reproduced
\emph{exactly}\footnote{In fact it is only because the results are exactly the
same that we are confident of our implementation.  With the careful use of
random seeds we were able to detect errors in our speculative implementation
when particle positions were off in the tenth digit.}, two iterations at a
time, using a speculative approach adapted from speculative execution. We show
that the standard PSO equations can be factored such that a set of speculative
positions can be found which will always include the position computed in the
next iteration.  By computing the value of the objective function for each of
the speculative positions at the same time the algorithm evaluates the
objective function for the current position, it is possible to know the
objective function values for both the current and the next iteration at the
same time.  We demonstrate this principle by implementation and show that it
produces exactly the same results as standard PSO, but two iterations at a
time.  The resulting implementation runs efficiently on large clusters where
the number of processors is much larger than a typical or reasonable number of
particles, producing better results in less wall-clock time.

We refer to this parallelization technique as ``speculative evaluation in
particle swarm optimization'', or SEPSO.  It is important to note here that
SEPSO is not a variant of PSO.  We simply propose a new way to think about the
parallelization of PSO that we show takes fewer iterations to reach the same
level of fitness than previous parallelizations in many instances.

Furthermore, we show that if we relax the requirements of the algorithm, no
longer demanding that it strictly reproduce the exact behavior of standard PSO,
we can introduce new speculative techniques that often out-perform both
standard parallelizations of PSO and SEPSO.  These relaxations make better use
of the information obtained from the extra exploration made by the speculative
function evaluations.  We also explore the idea that, like branch prediction in
processors, we need not speculatively evaluate \emph{all} possible future
positions, we can accelerate the algorithm even if we are just \emph{likely} to
have guessed right.  By pruning the speculation to just paths that are
statistically likely to reproduce the paths that are equivalent to PSO we can
increase the swarm size without increasing the number of speculative
evaluations.  We also consider several recovery strategies for cases where the
pruned set of speculative evaluations does not contain the evaluation that
standard PSO would have done.  A further improvement we explore is speculating
several iterations ahead instead of just one, which is made possible by pruning
the number of speculative evaluations.

This paper is organized as follows. \sec{pso} describes the particle swarm
optimization algorithm, and \sec{related} gives a brief overview of previous
parallelization techniques for this algorithm.  \sec{sepso} shows
mathematically how speculative evaluation can be done in parallel PSO to
perform two iterations at once, leaving implementation concerns to the
appendices.  In \sec{relax}, we discuss various methods of improving the
performance of speculative evaluation in PSO, all of which break the
requirement of strictly reproducing the behavior of the original algorithm.
\sec{setup} describes the experiments we ran, and \sec{results} presents the
results of those experiments.  In \sec{conclusion} and \sec{future} we conclude
and discuss future work.

\section{Particle Swarm Optimization}
\label{sec:pso}

Particle swarm optimization was proposed in 1995 by James Kennedy and Russell
Eberhart~\citep{kennedy-1995-particle-swarm-optimization}.  The algorithm is
used to intelligently search a multi-dimensional space in a way that is
inspired by the swarming and flocking behavior of birds and other animals. It
is a social algorithm that depends on interaction between particles to quickly
and consistently approximate the optimal solution to a given objective
function.

The motion of particles through the search space has three components: an
inertial component that gives particles momentum as they move, a cognitive
component where particles remember the best solution they have found and are
attracted back to that place, and a social component by which particles are
attracted to the best solution that any of their neighbors have found.  We use
the constricted PSO algorithm, which was recommended as the standard for PSO
research~\citep{bratton-2007-defining-a-standard-for-pso}.

At each iteration of constricted PSO~\citep{clerc-2002-constricted-pso}, the
position $\ppos_t$ and velocity $\pvel_t$ of each particle are updated as
follows:
\begin{align}
\label{eq:velupdate}
	\pvel_{t+1} &=
		\constriction \bigl[ \pvel_t
			+ \pcoeff\pURand_{t}\otimes(\pbest_{t} - \ppos_{t}) +
			\ncoeff\nURand_{t}\otimes(\nbest_{t} - \ppos_{t})
		\bigr] \\
\label{eq:posupdate}
	\ppos_{t+1} &= \ppos_{t} + \pvel_{t+1}
\end{align}
where \( \pURand_{t} \) and \( \nURand_{t} \) are vectors of independent random
numbers drawn from a standard uniform distribution, the \( \otimes \) operator
is an element-wise vector multiplication, $\pbest$ (called personal best) is
the best position the current particle has seen, and $\nbest$ (called
neighborhood best) is the best position the neighbors of the current particle
have seen~\citep{bratton-2007-defining-a-standard-for-pso}.  The parameters \(
\ncoeff \), \( \pcoeff \), and \( \constriction \) are given prescribed values
required to ensure convergence (2.05, 2.05, and .73,
respectively)~\citep{clerc-2002-constricted-pso}.

Changing the way neighbors are defined, usually called the ``topology,'' has a
significant effect on the performance of the algorithm.  In the Ring topology,
each particle has one neighbor to either side of it; in the Complete
topology\footnote{The Complete topology has often been unfortunately named Star
in the literature, which in graph theory refers to a completely different
topology.  Other names have also been used, including ``global topology'' and
gbest.  We use the graph theory term ``Complete'' in this paper.}, every
particle is a neighbor to every other
particle~\citep{bratton-2007-defining-a-standard-for-pso}.  In all topologies a
particle is also a neighbor to itself in that its own position and value are
considered when updating the particle's neighborhood best, $\nbest$.  Thus with
$p$ particles, using the Ring topology each particle with index $i$ has three
neighbors: $i-1$, $i$ (itself), and $i+1$.  With the Complete topology, each
particle has $p$ neighbors.

In this paper we use these topologies as well as a parallel adaptation of the
Complete topology, called Random, that has been shown to approximate the
behavior of Complete with far less
communication~\citep{mcnabb-2009-large-particle-swarms}.  In the Random
topology, each particle randomly picks two other particles to share information
with at each iteration, along with itself.  Thus in both the Ring and the
Random topologies, all particles have three neighbors.

The ideal topology and swarm size for PSO depend on the objective function.
Researchers have devised various benchmark functions and have found that the
ideal topology for one function may perform very poorly for another function.
The No Free Lunch Theorems for Optimization show that this is true in
general---if an algorithm performs well on average for one class of functions
then it must do poorly on average for other
problems~\citep{wolpert-1997-nfl-for-optimization}.  A good starting point
seems to be a swarm of 50 particles using a Ring
topology~\citep{bratton-2007-defining-a-standard-for-pso}.

\section{Related Work}
\label{sec:related}

The idea of speculative decomposition in the parallelization literature is not
new \citep{grama-2003-intro-to-parallel-computing}.  In the field of function
optimization, simulated annealing has previously been parallelized using this
technique~\citep{witte-1991-parallel-simulated-annealing-speculative}, though
we are not aware of other evolutionary or swarm intelligence based algorithms
having been parallelized with speculative decomposition.

There have been several parallelizations of PSO presented in the literature.
The improvements described in these papers come in two major areas: innovations
in implementation details and innovations in the use of topology and swarm size
to scale PSO to many processors.

\subsection{Innovative Implementations}

There are many ways to parallelize the basic PSO algorithm.  The most
fundamental decision to make in parallel PSO is which parallel architecture to
use.  Several architectures have been proposed, including Master-Slave, fully
distributed (sometimes called ``diffusion''), and reformulating PSO into
Google's MapReduce framework~\citep{belal-2004-parallel-models-for-pso,
mcnabb-2007-parallel-pso-using-mapreduce}.  For inexpensive objective
functions, graphics processors (GPUs) have also been used to parallelize PSO
using either texture rendering~\citep{li-2007-fine-grained-parallel-pso-gpu} or
a general purpose GPU API~\citep{zhou-2009-gpu-based-parallel-pso}.

An important part of the parallel architecture is the type of interprocessor
communication used.  PSO has previously been parallelized with both synchronous
and asynchronous communication.  Synchronous parallel implementations of PSO
reproduce the standard serial algorithm exactly.  This approach was first
described analytically by \citet{belal-2004-parallel-models-for-pso} and first
implemented by \citet{schutte-2004-parallel-global-optimization-with-pso}.  In
a typical master-slave algorithm, the master assigns tasks to slave processors,
and in parallel PSO, each task consists primarily of a function evaluation.
Updating the particle's position and value may also be included in the
task~\citep{belal-2004-parallel-models-for-pso}, or this work may be performed
in serial by the
master~\citep{schutte-2004-parallel-global-optimization-with-pso}.  Before
proceeding to the next iteration, particles communicate, and each particle
updates its neighborhood best.  Whether this communication step happens
sequentially on the master or in parallel, each particle must receive
communication from its neighbors before proceeding.  The benefits of the
synchronous PSO include its simplicity, repeatability, and comparability with
standard PSO, which may be essential in research applications.

Asynchronous parallel particle swarm
optimization~\citep{venter-2005-parallel-pso-asynchronous-evaluations,
koh-2006-parallel-asynchronous-pso} is a modification to the standard algorithm
which removes the synchronization point at the end of each iteration.  Instead,
particles iterate independently and communicate asynchronously.  In a typical
master-slave implementation of asynchronous parallel PSO, the master updates
each particle's personal best, neighborhood best, velocity, and position
immediately after receiving the function value from the slave processor.  Since
this update occurs while other particles are still being evaluated, it may use
information from the previous iteration for some
neighbors.\footnote{Asynchronous parallel PSO has been compared to the
``asynchronous updates'' variant of serial
PSO~\citep{koh-2006-parallel-asynchronous-pso}.  However, serial PSO with
asynchronous updates differs from standard PSO in that particles use newer
information, but asynchronous parallel PSO differs from standard PSO in that
particles use older information.} In a partially asynchronous implementation,
particles might wait for some but not all neighbors to complete before
proceeding~%
\citep{scriven-2008-asynchronous-pso-in-unreliable-distributed-environments}.
In some master-slave implementations, particles never get more than one
iteration ahead of
others~\citep{venter-2005-parallel-pso-asynchronous-evaluations,
koh-2006-parallel-asynchronous-pso}.  However, in a fully distributed
implementation, particles might never wait for information, and one particle
could complete many more iterations than another
particle~\citep{scriven-2008-distributed-pso-using-peer-to-peer}.  The main
effect of asynchronous evaluation is that processors spend less time
idle---this trait is particularly valuable when processors are heterogeneous or
function evaluation times are
varied~\citep{venter-2005-parallel-pso-asynchronous-evaluations,
koh-2006-parallel-asynchronous-pso}.  Asynchronous parallel PSO behaves
differently than the standard algorithm and may even produce different results
between runs.  Most reports conclude that asynchronous communication produces
similar numerical results to the standard algorithm, but the question has not
yet been thoroughly
addressed~\citep{venter-2005-parallel-pso-asynchronous-evaluations,
koh-2006-parallel-asynchronous-pso}.

\subsection{Scaling PSO to many processors}

The other area of research in parallelizing PSO deals not with the
implementation details of architecture and synchronicity, but with what should
be done with the PSO equations when many hundreds or thousands of processors
are available.  The main issues that have been addressed are how many particles
to use for a particular number of processors and what communication topology
should be employed.

The number of particles per processor has typically been decided by how long it
takes to evaluate the function being optimized.  When the function takes longer
than a few seconds to evaluate, previous techniques have assigned the number of
particles in the swarm to be the number of processors available~%
\citep{jin-2005-pso-antenna-designs,mcnabb-2009-large-particle-swarms},
advocating using as many processors as possible to get the best performance.
When the function takes far less time to evaluate than it takes to send a
message across a network (e.g., through the TCP/IP stack), parallel
implementations assign several or many particles to a single
processor~\citep{chu-2006-intelligent-parallel-pso,
chang-2005-parallel-pso-with-communication-strategies}.  Often the processor
only sends information about the best particle it evaluated to other
processors~\citep{belal-2004-parallel-models-for-pso}.

Another popular method is simply to run PSO independently on each of the
processors available, taking the best result when all of the runs complete.  It
should be noted that this is equivalent to the previously stated method of
assigning many particles to each processor, only with no communication between
processors instead of little communication.  Both of these methods can be
described as changes in the communication topology of the original PSO
algorithm~\citep{mcnabb-2009-large-particle-swarms}.

Thus previous work in parallelizing PSO, apart from creating innovative
implementations, has consisted entirely of increasing the swarm size and
adapting the topology to be better suited to parallel computation.

With regard to increasing the swarm size in PSO, some recent work has suggested
that increasing the swarm size throughout the course of the optimization
process provides better results than having a set swarm
size~\citep{hsieh-2009-efficient-population-utilization-for-pso,
montes-de-oca-2011-incremental-social-learning-pso}.  However, these results
focused on serial computation and are based on total number of function
evaluations, which, when running in parallel on expensive functions, is less
important than total number of iterations.  Other work focusing on
parallelization has shown that when extra processors are available they should
be used, as performance increases with swarm size when measuring in terms of
number of iterations~\citep{mcnabb-2009-large-particle-swarms,
jin-2005-pso-antenna-designs}.  If the swarm size were varied throughout the
course of the optimization process, some processors would be sitting idle at
most iterations.

The contribution of our work lies in the realm of efficiently utilizing
parallel resources.  In our work we use a synchronous, MapReduce implementation
of parallel PSO.  While we use a specific implementation, we describe how
speculative evaluation can be performed in any of the synchronous architectures
mentioned in the previous section.  The adaptation of our methods to
asynchronous PSO parallelization methods should be straightforward, though it
is left to future work.

\section{Speculative Evaluation in PSO}
\label{sec:sepso}

PSO can be trivially parallelized by assigning each particle's computation to
an individual processor.  But as we have seen in \fig{evals-sphere}, for some
functions, and for large numbers of processors, just adding particles reaches a
point of diminishing returns.  That is, beyond some point adding processors
with previous techniques does not help the algorithm reach any given level of
fitness significantly faster.  To fix this, instead of adding particles we
employ extra processors in a speculative approach that allows us to perform two
iterations at a time.

Our speculative methods require refactoring the PSO equations such that all
possible positions for each particle at iteration $t+1$ can be evaluated in
parallel along with the position of each particle at iteration $t$.  With some
careful bookkeeping, we can then piece together the results of iteration $t+1$
for each particle, thus using extra processors to evaluate two iterations of
the algorithm in the time it takes to evaluate the function once.  As we will
show in Sections~\ref{sec:proof} and~\ref{sec:topology}, a wise choice of
topology limits the necessary speculative evaluations to seven per particle.

To see the value of this refactoring, suppose that $1000$ processors are
available, and that the evaluation of the objective function takes one hour.
If we only want a swarm of $100$ particles, $900$ of the processors would be
sitting idle for an hour at every iteration, and it would take two hours to run
two iterations.  If instead we perform speculative evaluation, sending each of
the $7$ possible speculative positions of a particle to be computed at the same
time as its current position, we would use $800$ of the $1000$ processors and
perform two iterations in one hour.

In order to do two iterations at once, we must use 8 times as many processors
as there are particles in the swarm.  If these processors were not performing
speculative evaluation, they might instead be used for function evaluations
needed to support a larger swarm.  This raises the question of whether a swarm
of 100 particles doing twice as many iterations outperforms a swarm of 800
particles.  We show in \sec{results} that in many, though not all, instances, a
smaller swarm performing more iterations does in fact outperform a larger
swarm.

\sec{proof} shows in detail how the PSO equations can be refactored to allow
for speculative evaluation, proving that SEPSO exactly reproduces the behavior
of the PSO algorithm.  The section also introduces some notation used later in
the paper.  \sec{topology} gives a brief discussion of how the topology used
affects the amount of speculative computation needed.

\subsection{Refactoring the PSO Equations}
\label{sec:proof}

To perform two iterations at a time we must first refactor PSO such that the
determination of the value of the objective function is separate from the rest
of the computation.  For simplicity, this discussion will describe the case
where PSO is performing function minimization using the Ring topology.  In this
example, each particle has two neighbors, the ``right neighbor'' and ``left
neighbor,'' whose positions are represented as $\rightn$ and $\leftn$
respectively.  Though we will only describe the case of the Ring topology here,
the math is straightforward for other topologies.  Our discussion of the
implementation in \sec{implementation} is independent of specific topologies,
and we use several different topologies in our experiments.

The refactoring hinges on the idea that there are only a few possible new
positions, or updates, for $\pbest$ and $\nbest$ (assuming the random
coefficients $\pURand_{t}$ and $\nURand_{t}$ have been drawn).  For the Ring
topology there are 7 possible update cases, identified in
Table~\ref{tab:evals}.  We label each case with an identifier referring to the
source of the update: a minus sign ($-$) represents no update, $L$ represents
an update to $\nbest$ coming from the left neighbor, $R$ represents an update
to $\nbest$ coming from the right neighbor, and $S$ represents an update to
either $\pbest$ or $\nbest$ coming from the particle itself.  As an example,
$\casexn$ refers to the case that the particle finds a new personal best, but
neither it nor its neighbors find a position that updated its neighborhood
best.  In the equations that follow, we refer to an update case as $\casegen$,
and to the set of cases collectively as $\caseset$.

\begin{table}
  \caption{All possible updates for a particle with two neighbors}
  \label{tab:evals}
  \centering
  \begin{tabular}{lcc}
	Identifier&Source of $\pbest$ update&Source of $\nbest$ update\\
	\hline
	\hline
	$\casepn$&No update&No update\\
	\hline
	$\casepl$&No update&Left Neighbor\\
	\hline
	$\casepr$&No update&Right Neighbor\\
	\hline
	$\casexn$&Self&No update\\
	\hline
	$\casexl$&Self&Left Neighbor\\
	\hline
	$\casexr$&Self&Right Neighbor\\
	\hline
	$\casexx$&Self&Self\\
	\hline
  \end{tabular}
\end{table}

In order to incorporate the determination of which case occurs into the
position and velocity update equations, we introduce an indicator function
$\indic_{t+1}^{\casegen}$ for each case $\casegen \in \caseset$.  When
$\casegen$ corresponds to the case actually taken by PSO,
$\indic_{t+1}^{\casegen}$ evaluates to 1; otherwise it evaluates to 0.  We can
then sum over all of the cases, and the indicator function will make all of the
terms drop to zero except for the case that actually occurs.  For example, the
indicator function for the specific case $\casexn$ (which, as is shown in
Table~\ref{tab:evals}, means that the particle's personal best was updated, but
its neighborhood best was not) can be written as follows:

\begin{align}
  \nonumber
	\indic_{t+1}^{\casexn} & (\ofunc ( \ppos_{t} ) ,\ofunc(\leftn_{t}),
	\ofunc(\rightn_{t}) ,\ofunc(\pbest_{t-1}) ,\ofunc(\nbest_{t-1}))= \\
  \label{eq:deficasexn}
	&\begin{cases}
	   1 & \text{if} \ \ofunc(\ppos_{t}) < \ofunc(\pbest_{t-1}) \\
	   &\quad \text{and} \quad \ofunc(\nbest_{t-1}) < \ofunc(\ppos_{t}) \\
	   &\quad \text{and} \quad \ofunc(\nbest_{t-1}) < \ofunc(\leftn_{t}) \\
	   &\quad \text{and} \quad \ofunc(\nbest_{t-1}) < \ofunc(\rightn_{t}) \\
	   0 & \text{otherwise}
	\end{cases}
\end{align}

For each case $\casegen \in \caseset$, there is also a corresponding
velocity update function $\specvel_{t+1}^{\casegen}$.  When the case is
known, the specific values of $\pbest_t$ and $\nbest_t$ may be substituted
directly into~\eqref{eq:velupdate}.  For example, in case $\casexn$,
$\pbest_{t}=\ppos_{t}$, as \pbest was updated by the particle's current
position, and $\nbest_{t}=\nbest_{t-1}$, as $\nbest$ was not updated at
iteration $t$:
\begin{align}
\nonumber
	\specvel_{t+1}^{\casexn} & (\pvel_t, \ppos_{t}, \leftn_{t}, \rightn_{t},
	\pbest_{t-1}, \nbest_{t-1}, \pURand_{t}, \nURand_{t}) \\
\label{eq:defvcasexn}
		&= \constriction \bigl[ \pvel_{t} +
			\pcoeff\pURand_{t}\otimes(\ppos_{t} - \ppos_{t})
			+ \ncoeff\nURand_{t}\otimes(\nbest_{t-1} -
			\ppos_{t}) \bigr]
\end{align}

In the same way we can create notation for the position update function by
substituting into~\eqref{eq:posupdate}.  For compactness, we will drop the
parameters to $\specvel_{t+1}^{\casegen}$ since they can be inferred from the
subscripts.
\begin{align}
\label{eq:defpcasegen}
	\specpos_{t+1}^{\casegen} & (\ppos_{t}, \pvel_{t}, \leftn_{t},
	\rightn_{t} ,\pbest_{t-1} ,\nbest_{t-1}, \pURand_{t}, \nURand_{t})
	= \ppos_{t} + \specvel_{t+1}^{\casegen}
\end{align}

With this notation we can re-write the original PSO velocity
equation~\eqref{eq:velupdate}, introducing our sum over cases with the
indicator functions.  Again, we represent the indicator functions and velocity
functions without the parameters for compactness.  The equation becomes:
\begin{align}
\nonumber
	\pvel_{t+1} &=
		\constriction \bigl[ \pvel_t
			+ \pcoeff\pURand_{t}\otimes(\pbest_{t} - \ppos_{t})
			+ \ncoeff\nURand_{t}\otimes(\nbest_{t} -
			\ppos_{t}) \bigr] \\
\nonumber
	&= \sum_{c \in \caseset} \indic_{t+1}^{c} \ \constriction \bigl[ \pvel_t
			+ \pcoeff\pURand_{t}\otimes(\pbest_{t} - \ppos_{t})
			+ \ncoeff\nURand_{t}\otimes(\nbest_{t} -
			\ppos_{t}) \bigr]  \\
\label{eq:vel2update}
	&= \sum_{c \in \caseset} \ \indic_{t+1}^{c} \ \specvel_{t+1}^{c}
\end{align}

Similarly, the position update equation~\eqref{eq:posupdate} becomes:
\begin{align}
\label{eq:pos2update}
	\ppos_{t+1} &= \ppos_{t} + \pvel_{t+1}
	= \sum_{c \in \caseset} \ \indic_{t+1}^{c} \ \specpos_{t+1}^{c} 
\end{align}

The value of the objective function at $\ppos_{t+1}$ is given by:
\begin{align}
\label{eq:val2update}
	\ofunc (\ppos_{t+1}) = \sum_{c \in \caseset} \ \indic_{t+1}^{c}
	\ \ofunc(\specpos_{t+1}^{c})
\end{align}

Returning our attention to the computation of $\ppos_{t+1}$ in
\eqref{eq:pos2update} and writing it with the parameters which were omitted
above, we obtain:
\begin{align}
\nonumber
  \ppos_{t+1} = \sum_{c \in \caseset}
	&\indic_{t+1}^{c}(\ofunc ( \ppos_{t} ) ,\ofunc(\leftn_{t}),
	\ofunc(\rightn_{t}) ,\ofunc(\pbest_{t-1}) ,\ofunc(\nbest_{t-1})) \\
\label{eq:val2updatelong}
	& \specpos_{t+1}^{c}(\ppos_{t},\pvel_{t},\leftn_{t},\rightn_{t},
	\pbest_{t-1},\nbest_{t-1},\pURand_{t}, \nURand_{t})
\end{align}

In this form the important point to notice is that there are only $7$ values
(for this Ring topology) in the set $\{\specpos_{t+1}^{\casegen}: \casegen \in
\caseset\}$ and that none of them depend upon $f(\ppos_t)$ or any other
objective function evaluation at iteration $t$.  Note also that while there are
random numbers in the equation, they are assumed fixed once drawn for any
particular particle at a specific iteration.  Thus PSO has been refactored such
that the algorithm can begin computing all $7$ of the objective function
evaluations potentially needed in iteration $t+1$ \emph{before} $f(\ppos_t)$ is
computed.  Once the evaluation of $f(\ppos_{t})$ is completed for all particles
only one of the indicator functions $\indic_{t+1}^{\casegen}$ will be set to 1;
hence only one of the positions $\specpos_{t+1}^\casegen$ will be kept.

Although this speculative approach computes $\ofunc(\specpos_{t+1}^{\casegen})$
for all $\casegen \in \caseset$, even those for which $\indic_{t+1}^{\casegen}
= 0$, these extra computations will be ignored, and might just as well never
have been computed.  We call the set of computations
$\{\ofunc(\specpos_{t+1}^{c}) : \casegen \in \caseset\}$ ``speculative
children'' because only one of them is needed.

\subsection{Topology in Speculative Evaluation}
\label{sec:topology}

The number of speculative evaluations needed per particle depends on the number
of neighbors each particle has.  The number of update cases in a topology where
each particle has $n$ neighbors is $2(n+1)$; there are two possibilities for
updates to $\pbest$ (updated by the particle itself and not updated), and $n+1$
possibilities for updates to $\nbest$ (updated by each neighboring particle and
not updated).  When the particle is also a neighbor to itself, as is always the
case in commonly used topologies, one of the cases can be eliminated, as a
particle cannot be the source of an update to its neighborhood best without
also updating its personal best.  Thus we have $2(n+1)-1$, or $2n+1$,
speculative evaluations per particle.  In a swarm with $p$ particles and $n$
neighbors per particle, $(2n+1)p$ speculative evaluations are needed.

Because the number of speculative evaluations depends on the number of
neighbors a particle has, the choice of topology is an important one.  The use
of the Complete topology, where every particle is a neighbor to every other
particle, would require $O(p^2)$ speculative evaluations per iteration.
Clearly it is much more desirable to have a sparse topology, where $O(np)$ is
much smaller than $O(p^2)$.  However, some functions are better optimized with
the Complete topology and the quick spread of information it entails than with
sparse topologies.  Accordingly, we use the Random topology described
in~\citep{mcnabb-2009-large-particle-swarms}, which has been shown to
approximate the Complete topology.  In~\sec{results} we report results for
SEPSO using both the Ring topology and the Random topology on a number of
common benchmark functions.

\section{Speculative Variations on PSO}
\label{sec:relax}

Refactoring the PSO equations led us to find that speculative approaches are
possible in the parallelization of PSO.  SEPSO reproduces standard PSO exactly,
two iterations at a time, at the expense of requiring several times the number
of processors.  In this section we consider other speculative techniques
inspired by SEPSO that relax the requirement of exactly reproducing the
behavior of the original PSO algorithm.

We outline three main improvements to speculative evaluation.  First, in
\sec{pickbest} we describe a method that uses all of the information found in
doing speculative evaluations.  Then \sec{pruning} presents a technique that
reduces the number of speculative evaluations that need to be done for each
particle, allowing speculative evaluation to use larger swarm sizes with the
same number of processors.  Finally, \sec{manyiters} shows a method for
speculating several iterations ahead, instead of just one.

None of these methods fundamentally change the PSO algorithm.  They simply lead
to particles being at different iterations and having different values for
personal and neighborhood best positions than would have occurred in standard
PSO, because they receive different information.  These kinds of relaxations
are fairly typical in the parallelization of
PSO~\citep{koh-2006-parallel-asynchronous-pso}.

\subsection{Pick the Best Child}
\label{sec:pickbest}

In performing speculative evaluation as we have described it, $2n+1$
speculative evaluations are done per particle, while all but one of them are
completely ignored.  It seems reasonable to try to make use of the information
obtained through those evaluations instead of ignoring it.

To make better use of the extra speculative evaluations, instead of choosing
the speculative child that matches the branch that the original PSO would have
taken, we take the child that has the best value.  The methodology is exactly
the same as with SEPSO except for the process of choosing which speculative
child to accept.  The only change needed in \alg{centralized} (see Appendix A)
is in step 7, where the speculative child $\noeval{\s}_{t+1}$ with the best
value is chosen from the set $\noeval{\sset}_{t+1}$ instead of the child with
the matching branch.  We call this technique Pick Best.

This can be thought of as drawing a number of samples from the next iteration
and accepting the best one.  Speculative particles that move in good directions
are kept.  Intuition says that this technique favors exploitation over
exploration, but as we will show in \sec{results}, that is not always the case.

At this point it is also interesting to note a parallel between our methods and
parallel evolution
strategies~\citep{rudolph-1991-distributed-evolution-strategies}.  In evolution
strategies, a parent individual (representing a potential solution to some
objective function) produces a number of offspring by a mutation operator.  One
of the individuals is selected by a selection operator, and that individual
becomes the parent for the next
generation~\citep{beyer-2002-evolution-strategies}.  Our methods are similar,
where our mutation operator is simply the PSO motion equations and the
selection operator is either the indicator function introduced in \sec{sepso},
in the case of our original speculative algorithm, or the standard selection
operator based on fitness, in the case of this Pick Best technique.

\subsection{Pruning the Speculative Evaluations}
\label{sec:pruning}

Because the SEPSO requires so many extra evaluations, a natural step to take is
to eliminate some of them.  If we could reliably predict which branch were
going to be taken, we could limit ourselves to one speculative evaluation per
particle instead of $2n+1$.  With a fixed number of processors, this would
allow us to greatly increase the swarm size relative to that needed in the
original speculative algorithm (e.g., with 120 processors, a standard
parallelization has a swarm of size 120, complete speculative evaluation has a
swarm of size 15, and pruning the evaluations to only one per particle allows a
swarm of size 60).  As not all of the branches are evaluated in any given
iteration, we call this technique pruning.

We look at the statistical behavior of PSO to find probabilities of taking any
particular branch.  While we cannot with certainty predict which branch a
particle will take every time, if we can use statistics to narrow down the
$2n+1$ possible evaluations to a few likely candidates, we can decrease the
amount of computation required to do speculative evaluation and improve our
performance.

\subsubsection{Branch Statistics}
\label{sec:statistics}

In Table~\ref{tab:evals} we presented all possible branches that a particle
with two neighbors could take.  Here we lump all of the neighbors together and
consider the statistics for the five branches shown in
Table~\ref{tab:branches}.  In the identifiers, N represents an update to
$\nbest$ coming from any neighbor.

\begin{table}[ht]
  \caption{Five Branches to Consider for Statistics}
  \label{tab:branches}
  \centering
  \begin{tabular}{lcc}
	Identifier&Source of $\pbest$ update&Source of $\nbest$ update\\
	\hline
	\hline
	$\casepn$&No update&No update\\
	\hline
	$\casexn$&Self&No update\\
	\hline
	$\casexx$&Self&Self\\
	\hline
	$\casepN$&No update&Some Neighbor\\
	\hline
	$\casexN$&Self&Some Neighbor\\
	\hline
  \end{tabular}
\end{table}

\begin{table}[ht]
  \caption{Branch Statistics in PSO}
  \label{tab:stats}
  \centering
  \scriptsize
  \begin{tabular}{c|c|c|c|c|c|c|}
	Topology&Function&\casepn&\casexn&\casexx&\casepN&\casexN\\
	\hline
	\hline
	Ring&Sphere&53.0\%&9.3\%&11.4\%&20.2\%&6.2\%\\
	&Griewank&51.7\%&8.4\%&12.2\%&20.7\%&7.0\%\\
	&Rastrigin&49.5\%&4.8\%&14.6\%&21.3\%&9.9\%\\
	&Rosenbrock&51.3\%&7.4\%&12.9\%&21.1\%&7.3\%\\
	\cline{2-7}
	&Average&{51.3\%}&{7.5\%}&{12.8\%}&{20.8\%}&
	{7.6\%}\\
	\hline
	\hline
	Random&Sphere&66.7\%&11.9\%&2.6\%&15.6\%&3.1\%\\
	&Griewank&69.0\%&10.9\%&2.5\%&14.9\%&2.7\%\\
	&Rastrigin&81.9\%&5.5\%&1.5\%&10.0\%&1.0\%\\
	&Rosenbrock&74.2\%&7.7\%&2.2\%&14.0\%&1.8\%\\
	\cline{2-7}
	&Average&{73.0\%}&{9.0\%}&{2.2\%}&{13.6\%}&
	{2.2\%}\\
	\hline
	\hline
	Complete&Sphere&31.9\%&9.2\%&0.2\%&45.1\%&13.5\%\\
	&Griewank&35.3\%&8.4\%&0.2\%&44.1\%&11.9\%\\
	&Rastrigin&47.7\%&6.7\%&0.2\%&38.2\%&7.0\%\\
	&Rosenbrock&35.3\%&3.4\%&0.3\%&54.4\%&6.6\%\\
	\cline{2-7}
	&Average&{37.6\%}&{6.9\%}&{0.2\%}&{45.5\%}&
	{9.8\%}\\
	\hline
  \end{tabular}
\end{table}

We seek to find the probability of taking any given branch, given whatever
information is needed: $\prob(\caseset_t|\cdot)$.  In finding these
probabilities, we do not attempt to derive any distribution from the PSO
equations, we simply look at empirical distributions.  However, even with
empirical distributions, the problem with this approach is that it is not clear
what information influences the probability of taking a branch.  We look at two
factors that we believe have a significant influence on $\prob(\caseset_t)$:
topology ($T$) and function ($F$).  Thus we are looking at
$\prob(\caseset_t|T,F)$.

We show in Table~\ref{tab:stats} with what percentage a particle takes each of
these branches for three different topologies and four different functions.
All of our statistics are from swarms of 240 particles.  Brief experimentation
showed that other swarm sizes had similar statistics.  We ran 750 iterations on
all combinations of functions and topologies except for the functions Griewank
and Rastrigin with the Complete topology.  We found that those runs frequently
converged past machine precision after 500 iterations, and that led to
erroneously high values for the probability of \casepn.  Instead we ran for
only 450 iterations on those two combinations.  All of our results were
averaged over 20 runs of the algorithm; thus the probabilities presented are
the averages of 3.6 million trials for the branch taken (2.16 million for the
two with only 450 iterations).  Table~\ref{tab:stats} contains the results.
The definitions for all of the functions in the table are found in \sec{setup}.

The probabilities presented in Table~\ref{tab:stats} are interesting in and of
themselves and could probably be used to better understand the characteristics
of various topologies.  It is notable that there is small variation between
functions in any given topology, but the variation across topologies is far
greater.  However, our concern is with speculative evaluation.  We are
interested in predicting the branch that any given particle will take at a
particular iteration.  For our purposes, it appears that given a topology, the
probability of selecting a branch and the function are close to independent, or
$\prob(\caseset_t|T,F) \approx \prob(\caseset_t|T)$.

From Table~\ref{tab:stats} we can see that with the Random topology, we can
pick the first branch, corresponding to stagnation, and be right around 70\% of
the time.  With the Ring topology, we would be right 50\% of the time.
Branches \casepN\ and \casexN\ really correspond to several actual branches, as
all of the neighbors are lumped together.  The 20\% probability of taking
branch \casepN\ with the Ring topology can be split into two branches, as there
are only two neighbors.  It also turns out that the neighbor that last updated
the neighborhood best is the most likely to update it next time, so keeping
track of that information could be fruitful in trying to predict that branch.

The statistics for the Complete topology are less promising, as there are 240
neighbors that branch \casepN\ splits into, instead of two.  Pruning does,
however, allow for the possibility of using the Complete topology in
speculative evaluation while avoiding the explosion in the number of
evaluations it would otherwise entail.

Because there are a few branches with very high probabilities in the topologies
we are interested in, we can have hope that cutting out some of the evaluations
that have low probability will lead to an increase in overall performance.  In
order to implement this kind of pruning, the only change that needs to be made
to \alg{centralized} is in Step 2, where the speculative children are
generated.  Instead of generating all possible speculative children, generate
the subset of the children that is desired.

\subsubsection{Recovering from Pruning Too Much}
\label{sec:wrong}

When not all of the branches are evaluated, there is some probability of not
evaluating the branch that was actually taken by the original particle.  There
are a few possibilities for recovery in this case.  One is to leave the
particle as it is, not accepting any of the speculative evaluations, because
none of them were correct.  This leads to particles being at different
iterations, as some particles guess correctly while others do not.  Thus we
lose exact compatibility with the original PSO, though this particular
relaxation is nothing new; asynchronous adaptations of PSO do the same
thing~\citep{koh-2006-parallel-asynchronous-pso}.  As an aside, it is equivalent
in this case to simply increment the iteration number of particles which fail
to correctly predict their branch.  This keeps the iteration number constant
across all particles, simplifying the work needed to be done in determining
neighbors when dynamic topologies are involved.  We call this technique Social
Promotion.

Another possibility is to pick the best child, as described in \sec{pickbest}.
This ignores the fact that the branch might have been wrong; it does not
matter, because we simply accept the child that had the best value.

\subsection{More than one iteration ahead}
\label{sec:manyiters}

We need not simply produce speculative children for the next iteration.  We can
view all possible speculative evaluations for a particle as an infinite tree
with branching factor $2n+1$.  As we have already seen that doing one full
level of the tree produces too many extra evaluations to be profitable, it is
incredibly unlikely that doing two full iterations would produce decent
results.  But, if the idea of speculating more than one iteration ahead is
combined with wisely pruning the possible evaluations based on branch
statistics, we can use just a few extra evaluations to go two or more
iterations ahead on the most likely branches.

When speculating more than one iteration ahead, the idea of Social Promotion
cannot feasibly be implemented, as we can only determine correct branches for
the first iteration.  Thus in this case we always pick the child that has the
highest value.

The question of which branches to take in this infinite tree is an intriguing
one that we can only begin to explore here.  If the branch corresponding to
stagnation, \casepn, has a 75\% chance of being taken, as in the Random
topology with most functions, we could speculate three iterations ahead on that
branch and still have a 42\% chance of predicting correctly.  However,
intuition would say that perhaps it is better to hope that the particle is
productive instead of stagnant, so a branch where the particle updates its
personal best might be more fruitful to try.  In our experiments we try just
one of the countless possibilities, but one that turns out to work very well.
More work is needed to compare the different branching possibilities on various
functions.

\section{Experimental Setup}
\label{sec:setup}

\subsection{Parallel Architecture}

At each iteration of the algorithms, we use one processor to perform one
function evaluation.  The exact evaluation time at which this architecture
becomes reasonable depends on the amount of communication overhead in the
parallel implementation and the number of particles in the swarm.  For our
implementation we found that time to be around 100 milliseconds for swarms of
50 particles.  When the swarm size increases, the minimum evaluation time at
which this architecture should be used decreases.

A general rule of thumb for when this architecture is useful is when the
combined function evaluation time is greater than the communication overhead
for a single iteration; i.e., $t*p > o+t$, where $t$ is the function evaluation
time, $p$ is the desired number of particles (in our case $p$ is equal to the
number of available processors), and $o$ is the amount of overhead per
iteration.  The amount of overhead per iteration will vary based on the
communication topology and the number of particles, so this is only an
approximation, but it can provide rough estimates of function evaluation times
at which this parallel architecture should be used.

\subsection{Function Evaluations vs. Time Steps}

Results in serial PSO are typically presented in terms of function evaluations.
This is a natural abstraction from implementation details that still allows a
comparison of the implementation-independent aspects of each algorithm.  The
number of function evaluations performed is assumed to be proportional to
wall-clock time, as all evaluations are performed in serial.  Only reporting
function evaluations could hide the fact that one algorithm requires more
overhead than another and thus actually takes more time to perform the same
number of function evaluations; however, function evaluations are still
considered the standard method of reporting, as evaluation times for functions
also vary greatly and could make the additional overhead negligible.

In parallel PSO on long-running functions, the natural way to present results
is in terms of iterations, not function evaluations.  This is because when all
function evaluations at each iteration are performed concurrently, iterations
are the direct equivalent of wall-clock time.  Thus we report iterations in our
results instead of function evaluations.  But because SEPSO actually performs
two iterations of PSO at each ``iteration,'' and Social Promotion and Many
Iterations make the idea of ``iterations of PSO'' somewhat nebulous, we instead
call each ``iteration'' a ``time step.''

Just as serial PSO papers do not report actual running times of their specific
implementations, we do not report running times, favoring the abstraction of
time steps.  However, given the time required for the evaluation of the
objective function and the communication overhead per iteration for a specific
implementation, a running time can be estimated from all of our results.
Simply multiply the number of time steps by the sum of the function evaluation
time and the overhead.  We wish to stress that the ``time steps'' we report are
proportional to wall-clock time, given the architecture we have assumed.

\subsection{Functions Used}

We experimented with five common benchmark problems defined
in~\citep{herrera-2010-test-suite}: Rastrigin, Sphere, Schwefel 2.21, Griewank,
and Bohachevsky.  While benchmark functions take fractions of a second to
evaluate and thus have no need of parallelization, they are useful for
comparing algorithms because they stand as surrogates for the kinds of
functions practitioners are actually interested in, and they allow us to
explore the behavior of optimization algorithms in a simple and standardized
way.  To provide additional evidence that performance on benchmark functions
corresponds to performance on real-world problems, we also tested our
parallelization methods on a typical research problem, that of fitting a model
to a large quantity of data, which we refer to as ``the model fitting
problem.''

Rastrigin is initialized in $[-5.12,5.12]^D$ and is defined as:
\[f_{Rastrigin}(\Vec{x}) = \sum_{i=1}^D\left(z_i^2 - 10\cos\left(2\pi
z_i\right) + 10\right), z=x-c\]

Sphere is initialized in $[-50,50]^D$ and is defined as:
\[f_{Sphere}(\Vec{x}) = \sum_{i=1}^D z_i^2, z=x-c\]

Schwefel 2.21 is initialized in $[-500,500]^D$ and is defined as:
\[f_{Schwefel}(\Vec{x}) = \max_i{|z_i|, 1 \leq i \leq D}, z=x-c\]

Griewank is initialized in $[-600,600]^D$ and is defined as:
\[f_{Griewank}(\Vec{x}) = \frac{1}{4000}\sum_{i=1}^D z_i^2 - \Pi_{i=1}^D
\cos\left(\frac{z_i}{\sqrt{i}} \right) + 1, z=x-c\]

And Bohachevsky is initialized in $[-15,15]^D$ and is defined as:
\[f_{Bohachevsky}(\Vec{x}) = \sum_{i=1}^D(x_i^2+2x_{i+1}^2-.3\cos(3\pi
x_i)-.4\cos(4\pi x_{i+1})+.7)\]

In computing the branch statistics in \sec{statistics}, we also used the
Rosenbrock function.  That function is defined as:
\[f_{Rosenbrock}(\Vec{x}) = \sum_{i=1}^D
\left(100(x_{i+1}-x_i^2)+(x_i-1)^2\right)\]

The $c$ in the function definitions is a shifted center, in order to avoid
origin-seeking bias in the PSO algorithm and its
variants~\citep{monson-2005-origin-seeking-bias}.  We move the center of the
shifted functions (all except Bohachevsky) to be halfway between the center and
the boundary of the initialization region.  For example, if the initialization
region is $[-50,50]^D$, the center is $(25)^D$.  We tested all of these
functions in their 20 dimension, 50 dimension, and 500 dimension varieties.

The model fitting problem is formulated as follows.  We generated 10,000 data
points from a radial basis function network with 10 bases and some added
Gaussian noise.  We then fit a radial basis function network to the data using
PSO.  This amounted to a 30 dimensional function to optimize, with a function
evaluation time on the order of two seconds.

\subsection{Parallelization Techniques Compared}

The parallelization techniques we compare are the standard parallelization
(here labeled Standard simply for ease of reference), our original speculative
approach (recall that we refer to this as SEPSO), and the four relaxations of
SEPSO discussed in \sec{relax}.  In presenting our results, we call the
approach developed in \sec{pickbest} Pick Best.  The methods described in
\sec{pruning} are called Pick Best Pruned and Social Promotion Pruned, and the
method in \sec{manyiters} we call Many Iterations.

Using the same number of processors for each approach (and thus the same number
of function evaluations per time step) requires that our speculative
parallelizations have a smaller swarm size than the standard parallelization.
For the topologies we used with SEPSO and Pick Best, a particle has three
neighbors including itself.  As shown in Table~\ref{tab:evals}, this results in
$7$ speculative evaluations per particle.  With one evaluation needed for the
original, non-speculative particle, we have $8p$ evaluations for every two
iterations, where $p$ is the number of particles in the speculative swarm.  The
extra evaluations required in our speculative approach would instead be used to
evaluate particles in standard parallelizations, so we compare swarms of size
$p$ in speculative evaluation with swarms of size $8p$ in standard approaches.

When performing pruning in Pick Best Pruned and Social Promotion Pruned, there
are a large number of ways to prune speculative evaluations.  We experimented
with several, but present results for only one possible pruning.  The pruning
we present uses only the two branches where the $\nbest$ value was not updated:
\casepn\ and \casexn.  Those branches are convenient in that no messages are
needed from neighbors in order to produce the positions of the speculative
particles---in distributed frameworks using several rounds of communication
(see Appendix A), one of the rounds of communication can be dropped entirely.
Pruning all but these branches also allows the use of arbitrarily dense
topologies, as the number of speculative particles is no longer dependent on
the number of neighbors the particle has.

Because pruning only requires two speculative evaluations per particle (along
with evaluating the original particle), we can use swarms of size
$\frac{1}{3}p$ when pruning to compare to a swarm of size $p$ with the standard
parallelization, instead of $\frac{1}{8}p$ with other techniques.

There are also many ways to speculate several iterations ahead, and with Many
Iterations we again only show results for one of them.  The combination of
branches we tried uses seven speculative evaluations per particle, matching the
swarm size of the original speculative algorithm.  The seven evaluations we
used corresponded to several iterations of branches \casepn\ and \casexn.  Two
of the evaluations were just one iteration ahead, four were two iterations
ahead, and one was three iterations ahead.  The evaluations that were one
iteration ahead were branches \casepn\ and \casexn; those that went two
iterations ahead were formed by taking either branch \casepn\ or \casexn and
then branch \casepn\ or \casexn; and the evaluation going three iterations
ahead followed branch \casepn\ on all three iterations.  As with Pick Best
Pruned and Social Promotion Pruned, this choice of branches allows the use of
arbitrarily dense topologies, as the number of speculative evaluations per
particle is independent of the number of neighbors the particle has.

Finally, to show that our methods are orthogonal to traditional PSO variants,
we show results of parallelizing a variation of the PSO motion equations, Bare
Bones PSO~\citep{kennedy-2003-bare-bones-pso}, with both standard
parallelizations and our speculative approaches.  Our intent is not to
exhaustively compare parallelization strategies on all possible PSO variants,
but to show some evidence that speculative approaches are generally applicable
and successful across PSO variants.  These results are shown in
\sec{bare-results}.

\subsection{Topologies}

For each benchmark function we report results using the topology that is widely
considered best for that function, as reported in the literature
(e.g.,~\cite{bratton-2007-defining-a-standard-for-pso}).  In this paper we
limit ourselves to the Ring topology and the Complete topology, as is common
practice, along with the Random topology (the parallel approximation to the
Complete topology mentioned in \sec{pso}).  We also mentioned in \sec{related}
that some related work can be described as changes in topology, particularly
that of having subswarms of fully connected topologies that occasionally
communicate with each other.  This related work focused mainly on functions
with very fast evaluation times where such techniques drastically reduce
interprocessor communication.  With long function evaluations these topologies
are not as practical, as only one particle is evaluated by each processor at
each iteration.  We experimented with a subswarm topology and found that in
every instance except one it performed worse than either Complete or Ring, and
thus we leave it out of the results except in the one instance where it
improved performance.

Where the Complete topology would normally be preferred, we use a Random
topology in SEPSO and Pick Best, as Complete leads to an explosion in the
number of speculative evaluations (as noted, the other speculative techniques
can still use a Complete topology; we often report results on both topologies
for those methods).  If speculative evaluation were not being performed, it is
possible that the Complete topology would be used.  However, the Complete
topology also requires a very large amount of interprocessor communication in
distributed PSO, so it is still quite possible that Random would be used even
with standard parallelizations~\citep{mcnabb-2009-large-particle-swarms}.  But,
to be fair in our comparisons, we compare to the standard parallelization using
both the Random topology and the Complete topology (labeled PSO Random and PSO
Complete in our results).  Again, the amount of communication overhead is
heavily dependent on implementation details which are not the focus of this
paper.  A practitioner using a particular implementation can compare the
results given here for whichever topology is most practical given the specific
implementation.

\section{Results}
\label{sec:results}

We frequently present tables summarizing our results.  In each table, we bold
the ``best'' method, meaning that it has at least a 90\% success rate and its
average time to completion is faster than all other methods that have at least
a 90\% success rate\footnote{``Success rate'' here means that the algorithm
reached some predefined value which is mentioned in the table or figure
wherever a success rate is reported.}.  We used a double-sided t-test to test
for statistical significance in these results, and all algorithms whose values
are bolded are significantly superior to non-bolded algorithms with a p-value
of less than $10^{-5}$.

We first present results in \sec{20dims} for the 20 dimensional variants of the
benchmark functions we tested.  We give a discussion of each function
individually at 20 dimensions, as there are interesting characteristics of the
algorithms that are worth discussing.  In \sec{50dims} we then give a summary
of results for the 50 dimensional variants of the benchmark functions; we do
not go into as much detail in our discussion as the results are very similar to
those in 20 dimensions.  We finish our discussion of benchmark functions with
their 500 dimensional variants in \sec{500dims}, and we discuss results on the
model fitting problem in \sec{model}.

\subsection{20 Dimensions}
\label{sec:20dims}

\subsubsection{Sphere}
\label{sec:sphere20dims}

First we look at Sphere, the simplest of common benchmark functions.  The
function has a single global optimum and no other local optima.  Sphere is best
optimized in terms of function evaluations with a small swarm using a Complete
topology.  We expect our methods to be perfect for such functions, and our
results show this intuition to be correct.  For this comparison we used 240
processors, so the methods had swarms of size 30 (for SEPSO, Pick Best, and
Many Iterations), 80 (Pick Best Pruned and Social Promotion Pruned), and 240
(Standard).  We compared too many methods here to fit into one graph, so we
show just a few methods in \fig{sphere-20} and in \tabref{sphere-20} we show a
summary of the results for all methods we tested.

In \fig{sphere-20} we see that Many Iterations using a Complete topology
converges incredibly quickly on a very poor value.  We found this behavior to
be quite consistent across functions for this method, so we rarely show results
for Many Iterations Complete.  However, Pick Best Pruned works very well with
Complete on this function.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{sphere20}
  \caption{Function Sphere with 20 dimensions, comparing a pertinent subset of
  possible methods.  Each method performs one evaluation on each of 240
  processors per time step.}
  \label{fig:sphere-20}
\end{figure}

\begin{table}
  \caption{Summary of results for function Sphere with 20 dimensions,
  measuring number of time steps to reach a value of 1e-35}
  \label{tab:sphere-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&0\%&N/A&N/A\\
  \hline
  Standard Complete&100\%&931.0&19.1\\
  \hline
  SEPSO Ring&0\%&N/A&N/A\\
  \hline
  SEPSO Random&85\%&972.8&15.8\\
  \hline
  Pick Best Ring&100\%&917.4&21.0\\
  \hline
  Pick Best Random&100\%&768.1&14.3\\
  \hline
  Pick Best Pruned Ring&100\%&967.8&9.9\\
  \hline
  Pick Best Pruned Random&100\%&693.5&10.4\\
  \hline
  \textbf{Pick Best Pruned Complete}&\textbf{100\%}&\textbf{389.6}&
    \textbf{9.6}\\
  \hline
  Social Promotion Pruned Ring&0\%&N/A&N/A\\
  \hline
  Social Promotion Pruned Random&100\%&971.1&10.6\\
  \hline
  Social Promotion Pruned Complete&100\%&777.8&16.7\\
  \hline
  Many Iterations Ring&100\%&575.2&9.8\\
  \hline
  Many Iterations Random&100\%&442.4&7.1\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

\subsubsection{Schwefel 2.21}

Schwefel 2.21 is a function similar to Sphere (in that they are both unimodal),
but Schwefel 2.21 is benefited more by larger swarms than Sphere is (as it is
harder to optimize).  Thus our speculative algorithms often fail to outperform
Standard with 240 processors because the simple speculative methods only have
30 particles.  As we will show later, when we use 800 processors at 50
dimensions, our methods perform much better.  However, the pruned versions of
our methods have 80 particles and thus are able to outperform Standard even
with only 240 processors.  Because the graph of Schwefel 2.21 looks very
similar to that of Sphere, we simply present all of our results in
\tabref{schwefel-20}.

\begin{table}
  \caption{Summary of results for function Schwefel with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06}
  \label{tab:schwefel-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&0\%&N/A&N/A\\
  \hline
  Standard Complete&100\%&837.2&44.8\\
  \hline
  SEPSO Random&0\%&N/A&N/A\\
  \hline
  Pick Best Random&5\%&938.0&0.0\\
  \hline
  \textbf{Pick Best Pruned Random}&\textbf{100\%}&\textbf{639.9}&\textbf{29.4}\\
  \hline
  \textbf{Pick Best Pruned Complete}&\textbf{100\%}&\textbf{597.7}&
	\textbf{77.9}\\
  \hline
  Social Promotion Pruned Random&75\%&938.0&33.0\\
  \hline
  Social Promotion Pruned Complete&100\%&815.2&43.5\\
  \hline
  Many Iterations Random&100\%&783.5&94.5\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

\subsubsection{Rastrigin}

Rastrigin is a multi-modal function that is best optimized with a large,
Complete swarm.  It has been shown that with Rastrigin, the more particles
there are in the swarm, the lower function value it finds, up to at least 4000
particles~\citep{mcnabb-2009-large-particle-swarms}.  Smaller swarms get caught
in local optima and converge to poorer values.  In fact, none of the algorithms
that we tested were able to solve this problem optimally.  We thus set a
somewhat arbitrary success threshold, simply to be able to compare algorithms.

Because our speculative algorithms require significantly smaller swarm sizes,
we would expect to not perform very well on functions such as Rastrigin.  Our
experiments show our intuition to be correct.  In this experiment we used 240
processors, so SEPSO and Pick Best each had 30 particles, and the Standard
algorithms had 240 particles.  As expected, SEPSO and Pick Best converge
quickly to worse local optima than Standard does.  \fig{rastrigin} shows the
results graphically, and \tabref{rastrigin-20} shows results for all of the
methods we tried.

\begin{table}
  \caption{Summary of results for function Rastrigin with 20 dimensions,
  measuring number of time steps to reach a value of 20}
  \label{tab:rastrigin-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&25\%&743.4&212.0\\
  \hline
  Standard Random&100\%&372.1&86.6\\
  \hline
  \textbf{Standard Complete}&\textbf{95\%}&\textbf{273.1}&\textbf{99.7}\\
  \hline
  SEPSO Ring&5\%&471.0&0.0\\
  \hline
  SEPSO Random&15\%&208.0&17.0\\
  \hline
  Pick Best Ring&35\%&489.0&152.0\\
  \hline
  Pick Best Random&15\%&136.0&40.1\\
  \hline
  Pick Best Pruned Ring&95\%&545.9&208.5\\
  \hline
  Pick Best Pruned Random&10\%&95.0&11.0\\
  \hline
  Pick Best Pruned Complete&5\%&59.0&0.0\\
  \hline
  Social Promotion Pruned Ring&5\%&580.0&0.0\\
  \hline
  Social Promotion Pruned Random&80\%&278.2&136.3\\
  \hline
  Social Promotion Pruned Complete&50\%&200.5&130.2\\
  \hline
  Many Iterations Ring&10\%&117.0&25.0\\
  \hline
  Many Iterations Random&5\%&57.0&0.0\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{rastrigin}
  \caption{Function Rastrigin with 20 dimensions.  Each method performs one
  evaluation on each of 240 processors per time step.}
  \label{fig:rastrigin}
\end{figure}

\subsubsection{Griewank}
\label{sec:griewank-20}

It is recommended in the literature to use the Ring topology when optimizing
the Griewank function, as Complete is prone to premature convergence on a local
optimum~\citep{bratton-2007-defining-a-standard-for-pso}.  The global optimum
of Griewank has a value of 0.  When most trials reach the global optimum but a
few get stuck, the resultant ``average value'' graph has a flat line that is
misleading.  Thus instead we present plots showing the percent of trials that
have passed some threshold at each time step, as is common practice with these
functions~\citep{mendes-2004-population-topologies-in-pso}.  The threshold we
chose for this case was $10^{-6}$, as that value is below any local optima and
the swarm always successfully reaches 0 once is passes that point.

We ran 50 trials of each experiment with Griewank, so that the curves are more
smooth.  We show results in \fig{basic-griewank1} for swarms of size 100 and
800 using the Ring topology.  One can see in the figure that SEPSO reaches the
global optimum on average close to twice as fast as Standard, while Pick Best
is close to three times as fast and Many Iterations is even faster when it is
successful, though it is only successful 84\% of the time.
\tabref{griewank-20} shows all of our results in tabular form.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank6}
  \caption{Function Griewank with 20 dimensions.  Each method performs one
  evaluation on each of 800 processors per time step.  Instead of showing
  average function value, we show the percentage of runs that have found the
  global optimum by each iteration.  All algorithms use the Ring topology.}
  \label{fig:basic-griewank1}
\end{figure}

\begin{table}
  \caption{Summary of results for function Griewank with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06}
  \label{tab:griewank-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&762.2&19.4\\
  \hline
  SEPSO Ring&100\%&426.0&81.0\\
  \hline
  \textbf{Pick Best Ring}&\textbf{100\%}&\textbf{272.0}&\textbf{17.7}\\
  \hline
  \textbf{Pick Best Pruned Ring}&\textbf{100\%}&\textbf{282.9}&\textbf{10.7}\\
  \hline
  Social Promotion Pruned Ring&100\%&482.0&16.6\\
  \hline
  Many Iterations Ring&84\%&183.2&49.4\\
  \hline
  \end{tabular}
\end{table}

We pause here to show some interesting characteristics of the speculative
techniques we have proposed.  With 800 processors our methods perform very well
on this function.  With 240 processors, however, the results are much more
mixed.  Because 240 processors is near the point at which speculative
evaluation becomes useful, it is enlightening to see the behavior of the
various algorithms at this point.

We show results in \fig{basic-griewank2} for swarms of size 30 and 240 using
the Ring topology.  One can see in the figure that when SEPSO is successful, it
finds the optimum much faster than Standard.  However, because the swarm size
is so small, SEPSO gets stuck almost half of the time.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank1}
  \caption{Function Griewank with 20 dimensions.  Each method performs one
  evaluation of the objective function on each of 240 processors per time step.
  All algorithms use the Ring topology.}
  \label{fig:basic-griewank2}
\end{figure}

When we look at the performance of our Pick Best approach, we see that it
greatly improves performance on Griewank.  This is somewhat counter-intuitive,
because Griewank is deceptive and Pick Best seems like a greedy algorithm.  But
in \fig{griewank-pickbest} we see that Pick Best improves accuracy over SEPSO
by 20\%, while at the same time finding the optimum over 100 time steps sooner
on average.  It seems that while Pick Best is locally greedy, there is enough
exploration in the seven speculative evaluations to overcome the inherent
greediness of the approach.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank3}
  \caption{Function Griewank with 20 dimensions, comparing Pick Best with
  results from \fig{basic-griewank2}.  Each method performs one evaluation on
  each of 240 processors per time step.  All algorithms use the Ring topology.}
  \label{fig:griewank-pickbest}
\end{figure}

When we introduce pruning, our intuition about Pick Best turns out to be more
correct.  While adding 50 more particles to the swarm (as pruning allows us to
have 80 particles with 240 processors instead of only 30), Pick Best with
pruning still gets stuck just as often as the original Pick Best.  However,
Social Promotion does well with pruning; it increases the success rate to close
to 100\%, while still finding the optimum on average much faster than the
original PSO.  These results are shown in \fig{griewank-pruned}.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank4}
  \caption{Function Griewank with 20 dimensions, comparing pruning with the
  best results from \fig{griewank-pickbest}.  Each method performs one
  evaluation on each of 240 processors per time step.  All algorithms use the
  Ring topology.}
  \label{fig:griewank-pruned}
\end{figure}

With Griewank, the premature convergence problems inherent in picking the best
child are exacerbated when speculating several iterations ahead.  When Many
Iterations finds the optimum, it finds it quicker than any other method we
tried, on average four times faster than Standard.  However, it also gets stuck
and fails to find the optimum more than any other method.  The results are
shown in \fig{griewank-manyiters}.   \fig{griewank-manyiters} is also
interesting in that it highlights the trade-off between accuracy and speed in
the various approaches at this swarm size.  The faster the approach finds the
optimum, the less likely it is to be successful.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank5}
  \caption{Function Griewank with 20 dimensions, comparing speculating several
  iterations ahead with selections from \fig{griewank-pruned}.  Each method
  performs one evaluation on each of 240 processors per time step.  All
  algorithms use the Ring topology.}
  \label{fig:griewank-manyiters}
\end{figure}

Note here that \fig{griewank-manyiters} is interesting to compare to
\fig{basic-griewank1}.  The ordering of the methods in terms of the number of
time steps to completion is the same in both figures.  What is different is
that at 240 processors, most of our methods fail to find the optimum 100\% of
the time, while at 800 processors, all but Many Iterations succeed 100\% of the
time, and Many Iterations is very close.  At some point between 240 and 800
processors, SEPSO and Pick Best become successful, and at that point it is by
far better to use speculative methods than standard parallelizations.  At some
number above 800 processors, Many Iterations will be become successful and will
be the preferred method.

To further show this point, we present results using 1920 processors on the
20-dimensional Griewank function in \fig{griewank-all}.  With 1920 processors
Many Iterations finds the optimum 100\% of the time.  What is interesting is
that while going from 240 processors to 1920 processors, the standard method of
parallelization only decreased its time to completion by 5.9\%.  That is, the
extra 1680 processors in standard parallelizations provide \emph{no appreciable
benefit}.  In contrast, Many Iterations with 1920 processors has the same swarm
size of 240 particles and uses the additional processors for speculative
evaluation, giving a decrease in time to completion of 78.9\%, or a factor of
almost 5.  Once there are enough particles in the swarm to guarantee success
when optimizing a function, the best use of additional processors is to
speculate as far ahead as possible, speeding up the progress of the algorithm.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewankall}
  \caption{Function Griewank with 20 dimensions, comparing the performance of
  standard and speculative parallelizations as the number of processors
  increases.  All algorithms use the Ring topology.}
  \label{fig:griewank-all}
\end{figure}

\subsubsection{Bohachevsky}

Bohachevsky is a unimodal function best optimized with a Complete swarm.  It is
similar to Griewank in that there is a global optimum with a value of 0, and
the swarm sometimes finds it and sometimes does not.  Thus we present a graph
similar to those of Griewank, because ``average value'' graphs have a
misleading flat line.  We used 480 processors to optimize this function.  In
\fig{bohachevsky} we show a plot with a few pertinent methods, while
\tabref{bohachevsky-20} shows all of the results.  All of our speculative
approaches found the optimum much quicker than Standard with a Random topology.
However, SEPSO was slower than Standard Complete and got stuck 25\% of the
time.  Pick Best, Pick Best Pruned, and Many Iterations all outperformed
Standard Complete, with Many Iterations finding the optimum about twice as
fast.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{bohachevsky}
  \caption{Function Bohachevsky with 20 dimensions.  Each method performs one
  evaluation on each of 480 processors per time step.}
  \label{fig:bohachevsky}
\end{figure}

\begin{table}
  \caption{Summary of results for function Bohachevsky with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06.}
  \label{tab:bohachevsky-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Random&100\%&472.3&6.6\\
  \hline
  Standard Complete&100\%&238.2&11.2\\
  \hline
  SEPSO Random&75\%&261.6&9.2\\
  \hline
  Pick Best Random&100\%&211.8&9.2\\
  \hline
  Pick Best Pruned Random&100\%&189.7&3.4\\
  \hline
  Pick Best Pruned Complete&25\%&105.4&3.9\\
  \hline
  Social Promotion Pruned Random&100\%&260.3&6.5\\
  \hline
  Social Promotion Pruned Complete&90\%&197.9&12.2\\
  \hline
  \textbf{Many Iterations Random}&\textbf{94\%}&\textbf{118.2}&\textbf{2.6}\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

\subsection{50 Dimensions}
\label{sec:50dims}

Given our observation from the 20 dimensional benchmark functions that Many
Iterations most often outperforms other speculative approaches, we only present
results for Many Iterations and Standard for the 50 and 500 dimensional
variants.  We used 800 processors for these experiments because the problems
are more difficult, and because speculative approaches perform best when many
processors are available, as we showed in \sec{griewank-20}.  Many Iterations
Complete showed the same premature convergence that was observed with the 20
dimensional benchmark functions, so we do not include those results in our
tables; the success rate was 0\% for all functions.

Tables~\ref{tab:rastrigin-50} through~\ref{tab:bohachevsky-50} present
summaries of our results for each of the benchmark functions. As with the 20
dimensional variants, our speculative methods did not perform well on
Rastrigin, though for all other benchmarks Many Iterations significantly
outperformed Standard.

A summary of the results is shown in \tabref{50dims-summary}.  For functions
where Many Iterations outperformed Standard, we report the average speed up.
On all functions except Rastrigin, our methods showed an average speed up of
from two to six times compared to previous methods.

\begin{table}
  \caption{Summary of results for function Rastrigin with 50 dimensions,
  measuring number of time steps to reach a value of 100}
  \label{tab:rastrigin-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&5\%&776.0&0.0\\
  \hline
  \textbf{Standard Random}&\textbf{100\%}&\textbf{593.2}&\textbf{89.5}\\
  \hline
  Standard Complete&80\%&343.8&145.8\\
  \hline
  Many Iterations Ring&25\%&223.8&67.1\\
  \hline
  Many Iterations Random&25\%&98.6&11.0\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Sphere with 50 dimensions,
  measuring number of time steps to reach a value of 1e-06}
  \label{tab:sphere-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&30\%&986.8&8.2\\
  \hline
  Standard Complete&100\%&458.4&12.3\\
  \hline
  Many Iterations Ring&100\%&288.7&3.9\\
  \hline
  \textbf{Many Iterations Random}&\textbf{100\%}&\textbf{248.6}&\textbf{5.1}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Schwefel with 50 dimensions,
  measuring number of time steps to reach a value of 80}
  \label{tab:schwefel-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&70\%&871.6&57.1\\
  \hline
  Standard Complete&100\%&599.0&85.6\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{100\%}&\textbf{238.6}&\textbf{34.2}\\
  \hline
  \textbf{Many Iterations Random}&\textbf{100\%}&\textbf{242.9}&\textbf{95.4}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Griewank with 50 dimensions,
  measuring number of time steps to reach a value of 1e-06}
  \label{tab:griewank-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&1898.5&25.7\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{97\%}&\textbf{312.4}&\textbf{6.3}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Bohachevsky with 50 dimensions,
  measuring number of time steps to reach a value of 0.01}
  \label{tab:bohachevsky-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&1434.0&62.1\\
  \hline
  Standard Random&100\%&704.5&19.7\\
  \hline
  Standard Complete&20\%&377.5&14.9\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{95\%}&\textbf{229.8}&\textbf{11.4}\\
  \hline
  Many Iterations Random&35\%&171.6&4.0\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Average speed up, comparing the number of time steps to completion
  for the best speculative topology to the time steps to completion for the
  best standard topology.  All functions have 50 dimensions.}
  \label{tab:50dims-summary}
  \centering
  \begin{tabular}{|l|r|}
	\hline
	Function&Speed Up Factor\\
	\hline
	\hline
	Rastrigin&N/A\\
	\hline
	Sphere&1.84\\
	\hline
	Schwefel&2.51\\
	\hline
	Griewank&6.08\\
	\hline
	Bohachevsky&3.07\\
	\hline
  \end{tabular}
\end{table}

\subsection{500 Dimensions}
\label{sec:500dims}

At 500 dimensions the performance of constricted PSO on benchmark functions
becomes rather dismal.  In almost all cases, Standard fails to make any
significant progress.  Our search for an explanation seems to say that the
space is too large for particles to converge to the same point (due to
overshooting in some dimension), so all updates to the best position found come
from a single particle wandering in the space.  Every time that single particle
finds a new best position, its velocity contracts, so the particle is unable to
make significant progress on its own.  In constricted PSO it is necessary to
have a collection of particles exploring a promising location to keep the
particles' velocities from contracting prematurely, and in 500 dimensions the
space is too large to get the collection of particles to the same location.

The results that we present show that while Standard parallelizations suffer
from this problem, Many Iterations does not.  Though we can provide some
intuition for why this might be the case, it remains as future work to provide
a strong theoretical explanation.  The intuition is that we are speculating
about future positions along paths where each particle's velocity does not
contract.  The particular branches we selected corresponded to not having
received a new value for $\pbest$ nor $\nbest$.  In those branches, the
particle continues in the same direction it was going, without a contraction in
velocity.

Because none of the Standard approaches were successful at optimizing the 500
dimensional benchmark functions, we do not present tables as we did for the 50
dimensional variants.  We do, however, show a few figures demonstrating the
results we have just explained.  \fig{sphere500} shows the function Sphere,
\fig{griewank500} shows Griewank, and \fig{rastrigin500} shows Rastrigin.  Note
that at 500 dimensions speculative approaches outperform standard
parallelizations on Rastrigin, whereas they did not at 20 and 50 dimensions.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{sphere500}
  \caption{Function Sphere with 500 dimensions.  Each method performs one
  evaluation on each of 800 processors per time step.}
  \label{fig:sphere500}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank500}
  \caption{Function Griewank with 500 dimensions.  Each method performs one
  evaluation on each of 800 processors per time step.}
  \label{fig:griewank500}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{rastrigin500}
  \caption{Function Rastrigin with 500 dimensions.  Each method performs one
  evaluation on each of 800 processors per time step.}
  \label{fig:rastrigin500}
\end{figure}

\subsection{Model Fitting}
\label{sec:model}

For the model fitting problem we used 144 processors.  We show results for
three methods: Standard with a Random topology, Standard with a Subswarm
topology, and Many Iterations with a Random topology.  Standard with a Subswarm
topology had eight independent subswarms of 18 particles each, and the
particles in each subswarm were connected with a Random topology.  In the
graphs we call this method Standard Subswarms.

\fig{rbf} shows our results for this function.  The function value reported is
the sum squared error of the model fit.  \fig{rbf-stuck} shows the percent of
runs that reached a value for sum squared error of 55,000 by each time step,
which we designate as successful.  Many Iterations took on average 126 time
steps to reach this value, while Standard Subswarms took on average 298.  The
p-value for a t-test on this statistic is less than $10^{-8}$.  With our
implementation of parallel PSO, each time step took on average 10.06 seconds;
1.83 seconds of that was function evaluation time and the rest was overhead.
With this particular implementation, then, Many Iterations takes on average 21
minutes to reach a successful value, while Standard Subswarms takes 50 minutes.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{rbf}
  \caption{Results for fitting a radial basis function to noisy data.  We use
  144 processors for each method, so each time step corresponds to 144 function
  evaluations.}
  \label{fig:rbf}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{rbf_stuck}
  \caption{Results for fitting a radial basis function to noisy data.  We use
  144 processors for each method, so each time step corresponds to 144 function
  evaluations.}
  \label{fig:rbf-stuck}
\end{figure}

\subsection{Bare Bones PSO}
\label{sec:bare-results}

In order to demonstrate that our methods are applicable across PSO variants, we
ran experiments with Bare Bones PSO.  For this simple experiment we use the
Sphere function with 20 dimensions, taking the best Standard algorithm and the
best speculative algorithm on that function, which in this case is Standard
Complete and Pick Best Pruned Complete, as shown in \sec{sphere20dims}.
\fig{bare-bones} shows the results of running each of those algorithms with the
Bare Bones PSO equations.  We see a speed up of 1.6 when using speculative
parallelizations instead of Standard techniques.  While this is not quite as
large a speed up as the 2.5 seen with constricted PSO, it is still a
significant decrease in time.  The key point of this experiment is that our
methods still show improvements over Standard parallelizations in PSO variants
other than the standard constricted PSO.

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{bare_bones}
  \caption{Results for optimizing the 20-dimensional Sphere function with Bare
  Bones PSO.  We use 240 processors for each method, so each time step
  corresponds to 240 function evaluations.}
  \label{fig:bare-bones}
\end{figure}

\section{Conclusions}
\label{sec:conclusion}

We have described a new technique for using processors in parallel PSO to
improve the performance of the algorithm.  To our knowledge, this is the first
time extra processors have been used to do anything in PSO besides increase the
swarm size.  In an increasingly parallel world, such advancements will prove to
be crucial to the continued effectiveness of PSO.

We have detailed how to perform speculative evaluation in PSO in several
different parallel architectures.  Using this methodology, the behavior of the
original PSO algorithm can either be exactly reproduced, two iterations at a
time, or the behavior can be modified in order to improve performance.  While
exactly reproducing PSO sometimes uses too many extra processors to be useful,
when we allow ourselves some freedom with the algorithm we see great
improvements over previous methods.  We have shown results that conclusively
demonstrate the superiority of our techniques for several functions over the
standard practice of adding particles to the swarm when extra processors are
available, giving speed ups of up to six compared to previous parallelization
techniques.

What we have presented is not a new variant of the PSO equations.  We presented
a new parallelization technique, so we compared parallelization strategies on
the same algorithm, the original PSO.  Our most promising techniques do change
the behavior of the PSO algorithm slightly, as do some other previously
proposed parallelization techniques, such as asynchronous PSO.  However, our
methods are applicable to almost all PSO variants, and so a comparison using
the same variant for each of the parallelization techniques tested is
justified.  We presented results on a separate PSO variant that give evidence
for this claim.

We have given five different possible approaches to speculative evaluation,
each of which has different properties.  These approaches perform differently
on different functions and at different swarm sizes, as would be expected by
the No Free Lunch Theorem for
Optimization~\citep{wolpert-1997-nfl-for-optimization}.  We have given a brief
evaluation of the premature convergence properties of these methods on
deceptive functions when a smaller number of processors is available.  We have
also shown evidence that when many processors are available by far the best
thing to do in most cases is to speculate as far ahead as the extra processors
allow.  This point is shown in a single graph in \fig{griewank-all}.  In that
case, the extra processors in standard parallelizations provided \emph{no
appreciable benefit.}  In contrast, speculative methods allow additional
processors to provide additional speed ups, as the more processors are
available the farther ahead the algorithm can speculate.

Though our methods show great improvements on some functions, they do not work
for all functions at the swarm sizes we were able to experiment with.  As is
commonly known, in PSO there is a trade-off between exploration and
exploitation.  Some functions need only minimal exploration, and some never
seem to have enough.  Increasing the swarm size is a natural way to increase
exploration in a parallel environment.  However, once ``enough'' exploration
has been reached for any particular function, adding additional particles adds
only incremental benefits.  At this point, a better use of the additional
processors, as we have shown, is to perform some amount of speculative
evaluation.

Large parallel clusters are often required to successfully optimize practical
modern problems, and the world's computing resources are increasingly coming
from such clusters.  To properly use PSO with these systems, a balance needs to
be made between using processors to increase the swarm size and using them to
increase the speed of the algorithm.  This work is a first step in that
direction that opens the door to many future improvements on speculative
methods in the parallelization of PSO.

\section{Future Work}
\label{sec:future}

In this work we have focused on PSO itself and not all of its variants.  It
remains as future work to apply speculative approaches to recent and popular
PSO variants, such as the Fully Informed Particle
Swarm~\citep{mendes-2004-fully-informed-particle-swarm}.  While our methods
will not always be immediately applicable to every variant, we are confident
that some kind of speculative approach will be beneficial to the
parallelization of all forms of PSO, especially as the number of processors
used gets into the thousands.

We mentioned related work showing that increasing the swarm size throughout the
course of the algorithm could provide improved performance over a fixed swarm
size in serial PSO~\citep{montes-de-oca-2011-incremental-social-learning-pso}.
If this method were extended to parallel PSO, most processors would be idle in
the first few iterations, while more would be utilized at the end.  During
iterations where there are many un-utilized processors, a natural use of them
would be speculative evaluation, performing two or more of those iterations at
a time.

The issues of the sampling distribution of speculative relaxations, branch
statistics, and why standard PSO fails at 500 dimensions were briefly mentioned
in this paper.  Each of those issues needs further treatment.  The sampling
distribution of our speculative methods could be compared to Poli's description
of standard PSO's sampling
distribution~\citep{poli-2008-sampling-distribution-of-pso}.  The branch
statistics could be used to analyze topologies and discover why certain
topologies work well on some functions but not on others; perhaps PSO
performance is more dependent on the branch statistics of a combination of
topology and function than on the topology itself.  And discovering exactly why
constricted PSO fails at 500 dimensions while Many Iterations does not could
lead to improvements in the standard PSO algorithm, even when not running in
parallel.

We opened the door to speculative parallelization methods in PSO and described
the possible speculative evaluations to perform as an infinite tree from which
branches are selected.  However, we only presented a few of the countless
possibilities for selecting those branches.  Our methods for determining which
speculative evaluations to perform were independent of the particle; all
particles performed the same number and type of evaluations.  Another way to
allocate speculative evaluations is to somehow use the performance of each
particle to determine how many and which extra evaluations it can have.

Any other optimization algorithm that only depends on current sampling
positions when computing the next position to sample can be parallelized with
this technique.  In particular, genetic algorithms produce future generations
by combining individuals from the current generation.  With a large population
size there would be an unwieldy amount of possible future individuals, but the
potential exists to modify the algorithm to use some kind of speculative
evaluation.

\appendix

\section*{Appendix A Implementing Speculative Evaluation}
\label{sec:implementation}

It is not trivial in some parallel architectures to determine which speculative
position was the correct next position of each particle.  In this section we
discuss in detail some important considerations in the implementation of our
methods.  First we discuss the relatively easy case of a centralized parallel
PSO algorithm with a master computer and many slaves.  In such an architecture,
the master keeps track of all necessary information with only trivial message
passing needed.  Then we discuss the more complicated case of a distributed
algorithm, where each particle is on its own and needs to send and receive
messages to and from other particles.  Finally we discuss the further
complications of a dynamic topology such as Random, where a particle's
neighbors change from one iteration to another.

\subsection*{A.1 Terminology}

To aid in describing our methodology, we introduce a few terms.  A particle's
set of \emph{speculative children} is the set of all possible next iteration
states (including the particle's position, $\nbest$ and $\pbest$ positions)
that a particle could have.  We use $\p_t$ to denote a particle at iteration
$t$ and $\s_{t+1}$ to denote one of $\p_t$'s speculative children,
corresponding to one of the rows in Table~\ref{tab:evals}.  $\n_t$ is a
neighbor of particle $\p_t$.  Sets of particles are given by $\pset$, $\sset$,
or $\nset$, whereas single particles are simply $\p$, $\s$, or $\n$.

We separate each iteration of PSO into several steps.  First there is the
motion step, where a particle updates its position and velocity.  Then a
particle's position is evaluated, and the particle updates its current value
and its personal best.  Finally, a particle gets information from its neighbors
and updates its neighborhood best.

A particle at iteration $t-1$ that has been moved to iteration $t$ using
\eqref{eq:velupdate}~and~\eqref{eq:posupdate}, but whose position has
not yet been evaluated, is denoted as $\noeval{\p}_t$.  Once its position has
been evaluated, but it has still not yet received information from its
neighbors, it is denoted as $\nonbest{\p}_t$.  Only when the particle has
updated its neighborhood best is it a complete particle at iteration $t$.  It
is then simply denoted as $\p_t$.

\subsection*{A.2 Centralized Algorithms}

In a centralized, or Master-Slave, parallel PSO algorithm, one machine, the
master, keeps track of all necessary information, and all other machines are
merely used to evaluate the objective function at various positions as directed
by the master~\citep{belal-2004-parallel-models-for-pso}.  To perform
speculative evaluation in such an architecture, the master generates the
positions to evaluate speculatively as in \eqref{eq:defpcasegen}.  After having
the slaves evaluate the objective function at all necessary positions, the
master then decides which position to accept for each particle, as in
\eqref{eq:val2updatelong}.  The outline of the procedure is given in
\alg{centralized}.

\begin{algorithm}
  \caption{Speculative Evaluation in a Centralized PSO}
  \label{alg:centralized}
  \begin{algorithmic}[1]
	\STATE Move all $\p_{t-1}$ to $\noeval{\p}_t$ using
	  \eqref{eq:velupdate}~and~\eqref{eq:posupdate}
	\STATE For each $\noeval{\p}_t$, get its neighbors $\noeval{\nset}_t$ and
	  generate $\noeval{\sset}_{t+1}$ according to
	  \eqref{eq:defpcasegen}.
	\STATE Evaluate all $\noeval{\p}_t$ and $\noeval{\sset}_{t+1}$ in parallel
	\STATE Update personal best for each $\noeval{\p}_t$ and
	  $\noeval{\s}_{t+1}$, creating $\nonbest{\p}_t$ and $\nonbest{\s}_{t+1}$
	\STATE Update neighborhood best for each $\nonbest{\p}_t$, creating
	  $\pset_t$
	\FORALL{$\p_t$}
	\STATE Pick $\nonbest{\s}_{t+1}$ from $\nonbest{\sset}_{t+1}$ that matches
	  the branch taken by $\p_t$ according to
	  \eqref{eq:val2updatelong}.
	\STATE Pass along personal and neighborhood best values obtained by $\p_t$,
	  making $\nonbest{\p}_{t+1}$
	\ENDFOR
	\STATE Update neighborhood best for each $\nonbest{\p}_{t+1}$, creating
	  $\pset_{t+1}$
	\STATE Repeat from Step 1 until finished
  \end{algorithmic}
\end{algorithm}

Given a set of particles at iteration $t-1$ (perhaps which have just been
initialized), the master must move each particle using \eqref{eq:velupdate} and
\eqref{eq:posupdate} to obtain the set $\noeval{\pset}_t$.  For each particle
$\noeval{\p}_t$, the master must then get its set of neighbors
$\noeval{\nset}_t$ and use their positions, along with the position of
$\noeval{\p}_t$, to calculate all possible values of
$\specpos_{t+1}^{\casegen}$, using \eqref{eq:defpcasegen}.  These positions,
along with the original particle's associated information (such as values for
$\pbest$ and $\nbest$), define a set of speculative children,
$\noeval{\sset}_{t+1}$.  The master then has a set of particles
$\noeval{\pset}_t$, and for each particle a set of speculative children
$\noeval{\sset}_{t+1}$, which can all be evaluated at once.

The master then has the slaves evaluate the particles.  Once all particles,
speculative and original, have been evaluated and the values reported to the
master, the master determines which speculative child of each particle was the
correct one.  Mathematically, this corresponds to the evaluation of an
indicator function similar to that found in \eqref{eq:deficasexn}.  In
practice, this is done first by updating each (original) particle's $\pbest$,
if necessary, then by updating the particle's $\nbest$ with information from
the particle's neighbors.  This is simply the original PSO algorithm, and
corresponds to steps 1--5 in \alg{centralized}.  Given the updates to $\pbest$
and $\nbest$, the case from Table~\ref{tab:evals} can be determined, as per
\eqref{eq:val2updatelong}.  The child with the matching case is kept, and all
other speculative children are discarded (step 7 in \alg{centralized}).

The parent $\p_t$ must pass its personal best value to the child, as the child
knows only the position that it guessed, not the function value at that
position.  It is possible that both $\p_t$ and $\s_{t+1}$ update their personal
bests, but $\p_t$'s value is better.  For example, suppose that $\p_{t-1}$ has
a personal best value of 3, and that we are seeking to minimize the function.
$\noeval{\p}_t$ is created, and $\noeval{\s}_{t+1}$ is moved assuming that
$\p_t$ has updated its personal best with its position at time $t$.  Then both
$\noeval{\p}_t$ and $\noeval{\s}_{t+1}$ are evaluated, with values 1 and 2,
respectively.  $\nonbest{\s}_{t+1}$ would think that its current position is
its personal best, as the value it found, 2, is better than its previous
personal best value of 3.  It needs to receive the personal best value from its
parent to know that its personal best position $\pbest$ is actually the
position of $\p_t$, not $\s_{t+1}$.

The parent also needs to pass the value of the neighborhood best that the child
guessed.  The child only knows the position and needs the value in order to
make future comparisons between neighborhood best positions (step 8).

Upon picking the correct branch for each particle and updating the child's
personal best and neighborhood best value (from iteration $t$), the result is
the set $\nonbest{\pset}_{t+1}$, as the particles are now no longer
speculative.  What remains is to update the neighborhood best of those
particles from their neighbors (from iteration $t+1$), as above, to obtain
$\pset_{t+1}$.  That set of particles can subsequently be used to produce the
sets $\noeval{\pset}_{t+2}$ and $\noeval{\sset}_{t+3}$ (steps 1 and 2 in
\alg{centralized}), and the process repeats itself.

\subsection*{A.3 Distributed Algorithms}

\label{sec:distributed}

In a distributed parallel PSO algorithm, individual processors not only perform
evaluations of particles, but also their movement.  The information for each
particle is not held by a central machine that directs the algorithm; instead,
each processor has the information for the particle or particles that it is in
charge of and must perform the steps of the algorithm for those
particles~\citep{mcnabb-2007-parallel-pso-using-mapreduce}.  Messages such as
values and positions for the neighborhood best are sent between processors.
There may still be some machine that collects information from all of the
particles and outputs the result of the algorithm, though that machine's
importance is much less than in centralized algorithms.

To perform speculative evaluation in a distributed PSO algorithm, there must be
some way to have processors evaluate the speculative children of particles,
without giving the speculative particles the same treatment as actual
particles, as the speculative children only live for one iteration.  One way
that can be done is by assigning each particle a set of machines instead of a
single machine, and the particle directs its extra machines to evaluate its
speculative children.  The same information needs to be passed between
particles no matter the framework used.  We describe here the messages each
particle needs to receive to perform speculative evaluation.

A processor that is controlling a single particle $\p_{t-1}$ must first move
the particle to $\noeval{\p}_t$ and produce the particle's speculative children
$\noeval{\sset}_{t+1}$.  This is done in the same way as described above.
However, in order to produce $\noeval{\sset}_{t+1}$, the processor needs
information about the particle's neighbors, so there must be some message
passing to get that information.  Particularly, the information that the
processor needs is the position of each of the particle's neighbors at
iteration $t$.

To get that information, a round of message passing is required.  Each particle
sends its position to its neighbors at iteration $t$, so that all particles can
generate $\noeval{\sset}_{t+1}$.  After each particle evaluates its position
and the positions of its speculative children, it passes information about the
outcome of iteration $t$ to its neighbors, so that neighboring particles can
update their neighborhood bests to move from $\nonbest{\p}_t$ to $\p_t$.  Once
that communication is finished, the particle can select the speculative child
which matched the branch that iteration $t$ actually produced.  Then another
round of information passing follows, for iteration $t+1$, so that
$\nonbest{\p}_{t+1}$ can be updated to $\p_{t+1}$.  Two iterations have then
been completed with only one round of evaluations, and the next iteration can
start again with the first round of message passing.

In distributed frameworks, synchronizing all of the machines for a round of
message passing can be expensive.  The method just described uses three rounds
of message passing for every two iterations (corresponding to steps 2, 5 and 10
in \alg{centralized}).  It is possible to perform speculative evaluation in PSO
with only one round of communication per two iterations.  However, the
methodology is tedious and distracting from the present discussion, so we defer
its description to Appendix B.

\subsection*{A.4 Dynamic Topologies}

Performing speculative evaluation in PSO with a dynamic topology (where
neighbors change from iteration to iteration) raises a sticky issue of its own.
In a static topology, at iteration $t$ a particle already has all of the
information about the positions of its neighbors during iterations $1$ through
$t-1$.  If the neighbor finds a better position at iteration $t$, the particle
updates its neighborhood best, but if it does not, it still has its old
neighborhood best from its neighbors for all previous iterations.

In a dynamic topology, a particle might not have information about the previous
positions of its neighbors at iteration $t$.  That means that its new
neighborhood best could come not only from its neighbors' positions at
iteration $t$, but also from their personal best from iteration $t-1$, as
neighbors' personal bests are what are used to update a particle's neighborhood
best.  That creates a problem for speculative evaluation---there are
potentially more than $2n+1$ possible next positions, increasing the amount of
work that must be done to perform the second iteration at the same time as the
first.

This is easily fixed by updating each new particle $\noeval{\p}_{t+2}$ with the
currently available information about its neighbors $\noeval{\nset}_{t+2}$
before producing its children $\noeval{\sset}_{t+3}$.  If a particle
$\noeval{\p}_{t+2}$ updates its neighborhood best with the personal bests of
$\noeval{\nset}_{t+2}$ before calculating the next possible positions for
$\noeval{\sset}_{t+3}$, there are still only $2n+1$ possible next positions,
and the problem is avoided.

\section*{Appendix B Alternate form of message passing}

Here we describe a method that requires only one round of communication for
each pair of iterations, which happens at step 5 of \alg{centralized}.  Many
more messages are needed, but that is sometimes more desirable than
synchronizing all of the machines three times.

This second method only requires one round of passing information because
information about both iterations $t$ and $t+1$ is passed at the same time.
Each processor reconstructs from the messages it receives all of the
information that it needs about its neighbors.  Messages are passed directly
after evaluating each particle and its children, so all messages are of the
form of $\nonbest{\p}_t$ or $\nonbest{\s}_{t+1}$.  The first iteration needs to
be treated specially, so each particle can produce its initial set of
speculative children---neighbors need only pass their initial position.  This
kind of message passing necessitates the careful use of random seeds, so that
when each processor computes the motion equations for its neighbors it gets the
same results as its neighboring processors.

With the results of evaluating $\noeval{\p}_t$ and $\noeval{\sset}_{t+1}$,
along with all of the required messages from neighboring particles, the goal is
to produce $\p_{t+1}$ and output $\noeval{\p}_{t+2}$ and $\noeval{\sset}_{t+3}$
ready to be evaluated for the next iteration.  We first focus on the messages
needed to produce $\p_{t+1}$.

Upon evaluation, $\noeval{\p}_t$ becomes $\nonbest{\p}_t$, needing only to get
its neighborhood best information from its neighbors.  All of its neighbors,
then, must send it a message, so that from their updated personal best at
iteration $t$ the particle becomes $\p_t$.  The work done with the messages
received thus far is just as in regular PSO, and is graphically depicted in 
\fig{messages1}.

\begin{figure}
  \centering
  \psfrag{pt-n}[C][C]{$\nonbest{\p}_{t}$}
  \psfrag{nt-n}[C][C]{$\nonbest{\nset}_{t}$}
  \psfrag{pt}[C][C]{$\p_{t}$}
  \includegraphics[width=.15\columnwidth]{messages1}
  \caption{The production of $\p_{t}$ from the original particle
  $\nonbest{\p}_{t}$ and the messages $\nonbest{\nset}_{t}$.}
  \label{fig:messages1}
\end{figure}

With $\p_t$ we can select the correct speculative child as described above and
produce $\nonbest{\p}_{t+1}$.  Again we show the use of messages thus far
graphically, in \fig{messages2}.

\begin{figure}
  \centering
  \psfrag{pt-n}[C][C]{$\nonbest{\p}_{t}$}
  \psfrag{nt-n}[C][C]{$\nonbest{\nset}_{t}$}
  \psfrag{pt}[C][C]{$\p_{t}$}
  \psfrag{pt1-n}[C][C]{$\nonbest{\p}_{t+1}$}
  \psfrag{st1-n}[C][C]{$\nonbest{\sset}_{t+1}$}
  \includegraphics[width=.25\columnwidth]{messages2}
  \caption{The production of $\nonbest{\p}_{t+1}$ from the original particle 
  $\nonbest{\p}_{t}$, messages $\nonbest{\nset}_{t}$ and
  $\nonbest{\sset}_{t+1}$, and intermediate particles.}
  \label{fig:messages2}
\end{figure}

We then need the set of neighbors to $\nonbest{\p}_{t+1}$,
$\nonbest{\nset}_{t+1}$, so we can update $\nonbest{\p}_{t+1}$'s neighborhood
best.  To produce each neighbor $\nonbest{\n}_{t+1}$, we need the same
information for the neighboring particle that we needed to produce the original
particle, $\nonbest{\p}_{t+1}$; we need the original neighbor particle, its
speculative children, and its neighbors.  With that information, the set
$\nonbest{\nset}_{t+1}$ can be obtained by following the same process used to
obtain $\nonbest{\p}_{t+1}$.  We graphically show the messages needed to
produce $\nonbest{\nset}_{t+1}$ in \fig{messages3}.  Note that it looks
identical to \fig{messages2}, just with different sets of particles.

\begin{figure}
  \centering
  \psfrag{nt-n}[C][C]{$\nonbest{\nset}_{t}$}
  \psfrag{nnt-n}[C][C]{$\nonbest{\nnset}_{t}$}
  \psfrag{nt}[C][C]{$\nset_{t}$}
  \psfrag{nt1-n}[C][C]{$\nonbest{\nset}_{t+1}$}
  \psfrag{nst1-n}[C][C]{$\nonbest{\nsset}_{t+1}$}
  \includegraphics[width=.3\columnwidth]{messages3}
  \caption{The production of each $\nonbest{\n}_{t+1}$ from the original
  particle $\nonbest{\n}_{t}$, messages $\nonbest{\nnset}_{t}$ and
  $\nonbest{\nsset}_{t+1}$, and intermediate particles.  $\nnset$ is the set of
  neighbors for each particle $\n$, and $\nsset$ is the set of $\n$'s
  speculative children.  Note the similarity between this and \fig{messages2}.}
  \label{fig:messages3}
\end{figure}

With $\nonbest{\nset}_{t+1}$ and $\nonbest{\p}_{t+1}$, we can produce
$\p_{t+1}$.  This is shown in \fig{messages4}.  Note that we just combined
Figures~\ref{fig:messages2}~and~\ref{fig:messages3}, putting them together
to make $\p_{t+1}$, as all the particle needs is its neighborhood best to be
updated.

\begin{figure}
  \centering
  \psfrag{pt-n}[C][C]{$\nonbest{\p}_{t}$}
  \psfrag{nt-n}[C][C]{$\nonbest{\nset}_{t}$}
  \psfrag{pt}[C][C]{$\p_{t}$}
  \psfrag{pt1-n}[C][C]{$\nonbest{\p}_{t+1}$}
  \psfrag{st1-n}[C][C]{$\nonbest{\sset}_{t+1}$}
  \psfrag{nt-n}[C][C]{$\nonbest{\nset}_{t}$}
  \psfrag{nnt-n}[C][C]{$\nonbest{\nnset}_{t}$}
  \psfrag{nt}[C][C]{$\nset_{t}$}
  \psfrag{nt1-n}[C][C]{$\nonbest{\nset}_{t+1}$}
  \psfrag{nst1-n}[C][C]{$\nonbest{\nsset}_{t+1}$}
  \psfrag{pt1}[C][C]{$\p_{t+1}$}
  \includegraphics[width=.4\columnwidth]{messages4}
  \caption{The production of $\p_{t+1}$ from the original particle 
  $\nonbest{\p}_{t}$, messages $\nonbest{\nset}_{t}$, $\nonbest{\sset}_{t+1}$,
  $\nonbest{\nnset}_{t}$, and $\nonbest{\nsset}_{t+1}$, and intermediate
  particles.  Note that this is just a combination of \fig{messages2} and
  \fig{messages3}.}
  \label{fig:messages4}
\end{figure}

In order to get $\p_{t+1}$, then, a particle needs to receive messages from its
neighbors, its neighbors' neighbors, its speculative children, and its
neighbors' speculative children.  The particle $\p_{t+1}$ can be passed to some
central machine to track the progress of the algorithm, and it can be moved to
$\noeval{\p}_{t+2}$ in order to start the next iteration.

The next goal is to produce the set $\noeval{\sset}_{t+3}$.  As described
above, the necessary components to produce $\noeval{\sset}_{t+3}$ are
$\noeval{\p}_{t+2}$ and the neighbors of $\noeval{\p}_{t+2}$,
$\noeval{\nset}_{t+2}$.  We already have $\noeval{\p}_{t+2}$, so what remains
is to produce $\noeval{\nset}_{t+2}$.  It is sufficient to obtain
$\nset_{t+1}$, as each neighbor particle $\n_{t+1}$ can be moved with the
motion equations to $\noeval{\n}_{t+2}$.

We have already described how to use a set of messages to obtain $\p_{t+1}$.
The process is exactly the same to produce each $\n_{t+1}$, requiring the same
messages, only for the neighbor particles instead of the particle itself.
\fig{messages5} shows graphically how $\noeval{\sset}_{t+3}$ is produced.

\begin{figure}
  \centering
  \psfrag{pt1}[C][C]{$\p_{t+1}$}
  \psfrag{nt1}[C][C]{$\nset_{t+1}$}
  \psfrag{nt2-e}[C][C]{$\noeval{\nset}_{t+2}$}
  \psfrag{pt2-e}[C][C]{$\noeval{\p}_{t+2}$}
  \psfrag{st3-e}[C][C]{$\noeval{\sset}_{t+3}$}
  \includegraphics[width=.25\columnwidth]{messages5}
  \caption{The production of $\noeval{\p}_{t+2}$ and $\noeval{\sset}_{t+3}$
  from $\p_{t+1}$ and $\n_{t+1}$, each of which are produced as in
  \fig{messages4}.}
  \label{fig:messages5}
\end{figure}

Having obtained both $\noeval{\p}_{t+2}$ and $\noeval{\sset}_{t+3}$ from the
messages received, the algorithm then moves to the evaluation phase, and the
process repeats itself.  The particles are evaluated, send their messages, and
produce the next set of particles to be evaluated from the messages received.

To perform the entire process, at each message passing round a particle must
receive messages from its neighbors, its neighbors' neighbors, its neighbors'
neighbors' neighbors, its speculative children, its neighbors' speculative
children, and its neighbors' neighbors' speculative children.  With the Ring
topology, that looks like more messages than it really is, as many of the
neighbors' neighbors are duplicates.  With the Random topology, however, the
list of necessary messages could be rather large.

One more issue arises when dealing with dynamic topologies.  With neighbors
changing each iteration, messages that processors pass to their neighbors need
to be sent to the correct neighbors for each iteration.  A particle cannot
simply send messages to its neighbors' neighbors' neighbors---it needs to send
messages to its iteration $t$ neighbors' iteration $t+1$ neighbors, and so on.
For every neighbor outward information is sent, the iteration also needs to be
incremented, as information about neighbors' neighbors is used during iteration
$t+1$, and information about neighbors' neighbors' neighbors is used to
reconstruct information about iteration $t+2$.  Also, this method of message
passing again requires the use of random seeds if the topology is random, so
that each processor computes the same neighbors for a particle as all other
processors.

This may seem like an inordinate amount of work, and with some distributed PSO
frameworks it is.  However, other parallel frameworks necessitate this type of
message passing, so we have described how speculative evaluation can be
performed in those circumstances.

\bibliographystyle{apa-good}
\bibliography{%
../../../bib/aml/bib,%
../../../bib/nfl/bib,%
../../../bib/optimization/bib,%
../../../bib/parallel/bib,%
../../../bib/pso/general/bib,%
../../../bib/pso/parallel/bib,%
../../../bib/pso/fips/bib,%
../../../bib/pso/topology/bib,%
../../../bib/sim-anneal/parallel/bib,%
../../../bib/ga/bib}

\end{document}